### 2025-04-10 README.qmd
rrtool setup.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# WithinChert

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/jbelmiro/WithinChert/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2025-04-10 analysis/paper/paper.qmd
Daily work session. Chunks moved to paper and SOM. All SEM-EDS graphs for paper and SOM created.

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

```{r}
#| label: libraries

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(ggfortify)
library(FactoMineR)
library(factoextra)
library(ggpmisc)
library(magick)

```

```{r}
#| label: xrd-readtables

# Read grouped supporting data
xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

# Read individual xrd files
sp15_xrd <- read_csv('data/xrd_data/SP15_xrd.csv')
sp58_xrd <- read_csv('data/xrd_data/SP58_xrd.csv')
sp62_xrd <- read_csv('data/xrd_data/SP62_xrd.csv')
sp69_xrd <- read_csv('data/xrd_data/SP69_xrd.csv')

```

```{r}
#| label: xrd-graphs

xrd_marks <- xrd_marks |>
  dplyr::mutate(Phase = case_when(
    Phase == "SiO2" ~ "Quartz",
    TRUE ~ Phase))

# Colour codes
## Observed = "lightsteelblue3"
## Quartz = "salmon"
## Dolomite = "steelblue1"
## Calcite = "darkseagreen"
## Halite = "darkgoldenrod1"
## Background = "orchid"
## Difference = "snow4"

# SP15
sp15_marks <- xrd_marks |>
  filter(Sample == "SP15")
sp15_perc <- xrd_perc |>
  filter(Sample == "SP15") |>
  select(-Sample, -ESD)

sp15_plot <- sp15_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp15_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp15_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp15_perc)))

# SP58 Jordana
sp58_marks <- xrd_marks |>
  filter(Sample == "SP58")
sp58_perc <- xrd_perc |>
  filter(Sample == "SP58") |>
  select(-Sample, -ESD)

sp58_plot <- sp58_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp58_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp58_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp58_perc)))

# SP62 Oliveiras
sp62_marks <- xrd_marks |> 
  filter(Sample == "SP62")
sp62_perc <- xrd_perc |> 
  filter(Sample == "SP62") |> 
  select(-Sample, -ESD)

sp62_plot <- sp62_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp62_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp62_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp62_perc)))

# SP69 Mareta
sp69_marks <- xrd_marks |> 
  filter(Sample == "SP69")
sp69_perc <- xrd_perc |> 
  filter(Sample == "SP69") |> 
  select(-Sample, -ESD)

sp69_plot <- sp69_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp69_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp69_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp69_perc)))

```

```{r}
#| label: pxrf-read

pxrf_raw <- read_csv('data/pxrf_data_dec.csv')

```

```{r}
#| label: sem-read

M18_7 <- read_csv('data/sem_data/M18_7.csv')
M18_8 <- read_csv('data/sem_data/M18_8.csv')
M4_3 <- read_csv('data/sem_data/M4_3.csv')

```

```{r}
#| label: sem-graphs

# Prepare the energy ranges
energy_ranges <- list(
  C = c(0.2, 0.3),
  O = c(0.5, 0.6),
  Al = c(1.48, 1.5),
  Si = c(1.7, 1.8),
  P = c(2, 2.1),
  Ca = c(3.68, 3.7),
  Ca_beta = c(4.01, 4.02),
  Ti = c(4.5, 4.6),
  Ti_beta = c(4.9, 5),
  Ti_L = c("0.45", "0.47"),
  Fe = c(6.38, 6.41),
  Fe_beta = c(7.05, 7.07),
  Fe_L = c(0.71, 0.73))

# M18_8
filtered_M18_8 <- subset(M18_8, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_8 <- list()
for (element in names(energy_ranges)) {
  range_M18_8 <- energy_ranges[[element]]
  subset_M18_8 <- subset(filtered_M18_8, Energy >= range_M18_8[1] & Energy <= range_M18_8[2])
  if (nrow(subset_M18_8) > 0) {
    max_row_M18_8 <- subset_M18_8[which.max(subset_M18_8$Counts), ]
    peaks_list_M18_8[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_8$Energy,
      Counts = max_row_M18_8$Counts
    )
  }
}

peaks_M18_8 <- do.call(rbind, peaks_list_M18_8)
rownames(peaks_M18_8) <- NULL
peaks_M18_8 <- peaks_M18_8 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M18_8_graph <- filtered_M18_8 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M18_8, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M18_8, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M18_7
filtered_M18_7 <- subset(M18_7, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_7 <- list()
for (element in names(energy_ranges)) {
  range_M18_7 <- energy_ranges[[element]]
  subset_M18_7 <- subset(filtered_M18_7, Energy >= range_M18_7[1] & Energy <= range_M18_7[2])
  if (nrow(subset_M18_7) > 0) {
    max_row_M18_7 <- subset_M18_7[which.max(subset_M18_7$Counts), ]
    peaks_list_M18_7[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_7$Energy,
      Counts = max_row_M18_7$Counts
    )
  }
}

peaks_M18_7 <- do.call(rbind, peaks_list_M18_7)
rownames(peaks_M18_7) <- NULL
peaks_M18_7 <- peaks_M18_7 |> 
  filter(Counts > 100)

M18_7_graph <- filtered_M18_7 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M18_7, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M18_7, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M4_3
filtered_M4_3 <- subset(M4_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M4_3 <- list()
for (element in names(energy_ranges)) {
  range_M4_3 <- energy_ranges[[element]]
  subset_M4_3 <- subset(filtered_M4_3, Energy >= range_M4_3[1] & Energy <= range_M4_3[2])
  if (nrow(subset_M4_3) > 0) {
    max_row_M4_3 <- subset_M4_3[which.max(subset_M4_3$Counts), ]
    peaks_list_M4_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M4_3$Energy,
      Counts = max_row_M4_3$Counts
    )
  }
}

peaks_M4_3 <- do.call(rbind, peaks_list_M4_3)
rownames(peaks_M4_3) <- NULL
peaks_M4_3 <- peaks_M4_3 |> 
  filter(Counts > 100)

M4_3_graph <- filtered_M4_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M4_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M4_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

```

```{r}
#| label: pxrf-database-setup

## DB setup
pxrf_db <- pxrf_raw |> 
  filter(!Sample_ID %in% c("I19-1751", "H21-2126")) # Removing 2 altered samples
pxrf_analysis <- pxrf_db |> 
  select("Sample_ID", "pXRF_ID", "Formation/Type",
         "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
         "Sr", "Zr") |>  # Removed elements with 0 or <LOD values in all samples or in high percentages in all groups
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2",
    `Formation/Type` == "TYPE 3" ~ "T3",
    `Formation/Type` == "TYPE 4" ~ "T4",
    `Formation/Type` == "TYPE 5" ~ "T5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  ))

```

```{r}
#| label: pxrf-fig-lod
#| fig-cap: A barplot.

# Table to see the occurrence of 0 or <LOD values within each element and by rm type, to choose the elements in which to apply statistics.
# Combine the multiple columns into a single table
lod_tab <- pxrf_db |>
  select(-contains("Err")) |>
  select(-pXRF_ID, -Application, -Method, -ElapsedTime, -`Cal Check`, -Sample_ID) |>
  mutate(across(.cols = MgO:U,
                .fns = ~ case_when(
                  . == 0 ~"Success 0",
                  . > 0 ~ "Success",
                  TRUE ~ as.character(.))
                ))
# Reshape the data into a long format
lod_tab_long <- lod_tab |>
  pivot_longer(cols = MgO:U,  # Assuming these columns contain the 'Success' and 'LOD' values
               names_to = "Element",  # New column for the names of elements
               values_to = "Value")   # New column for the corresponding Success/LOD values

lod_tab_summary <- lod_tab_long |>
  group_by(Element, Value) |>
  summarise(Count = n(), .groups = "drop")  # Count the occurrences of Success and LOD for each Element

# Reshape the data into a wide format using pivot_wider
lod_tab_table <- lod_tab_summary |>
  pivot_wider(names_from = Value, values_from = Count, values_fill = list(Count = 0)) |> 
  filter(Success > 80) |> 
  select(Element, Success, "< LOD")
lod_tab_table_long <- lod_tab_table |> 
  pivot_longer(cols = c(Success, `< LOD`), names_to = "Metric", values_to = "Count") |> 
  mutate(Metric = factor(Metric, levels = c("Success", "< LOD"))) 
lod_tab_table_long |> 
  ggplot(aes(x = Element, y = Count, fill = Metric)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Success and < LOD", y = "Count") +
  theme_minimal() +
  scale_fill_manual(values = c("Success" = "#99D3A7", "< LOD" = "#ED7673"))

# Create the bar plot
ggplot(lod_tab_summary, aes(x = Element, y = Count, fill = Value)) +
  geom_bar(stat = "identity", position = "stack") +  # Use stat = "identity" to use summarized counts
  scale_fill_manual(values = c("Success" = "darkseagreen", "Success 0" = "lightgoldenrod1", "< LOD" = "indianred3")) +
  labs(title = "Summarized Success and LOD Counts by Element",
       y = "Count", 
       fill = " ") +
  theme_minimal() +
 # facet_wrap(~ Element, scales = "free_y") +  # Facet by Element with independent y-axes
  theme(axis.text.x = element_text(hjust = 1),
        axis.title.x = element_blank()) +
  #geom_text(data = lod_tab_summary_filtered, aes(label = Count), position=position_stack(vjust=0.5), colour="white", size = 3) +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, max(lod_tab_summary$Count), by = 20))

```

```{r}
#| label: pxrf-PCA
#| include: false

## a) PCA geological regional cherts
pca_db_algarve <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, S, Cl, K2O, CaO, TiO2, Fe2O3) |>
  filter(`Formation/Type` %in% c("LJW", "MJE", "UJE", "UJW", "TW"))
# Convert "< LOD" values to NA
pca_db_algarve[pca_db_algarve == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_algarve$`Formation/Type` <- as.factor(pca_db_algarve$`Formation/Type`)  # Ensure it's a factor
group_var_algarve <- pca_db_algarve$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_alg <- pca_db_algarve %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_alg <- complete.cases(df_numeric_alg)  # Identify complete rows
df_numeric_alg <- df_numeric_alg[complete_cases_alg, ]  # Remove NA rows
group_var_algarve <- group_var_algarve[complete_cases_alg]  # Filter group_var accordingly
# Scale values
df_scaled_alg <- scale(df_numeric_alg)
# Ensure row names are properly set
rownames(df_scaled_alg) <- make.names(group_var_algarve, unique = TRUE)
# Perform PCA
pca_result_alg <- PCA(df_scaled_alg, graph = FALSE)
# Summary of PCA
summary(pca_result_alg)
fviz_eig(pca_result_alg, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_alg <- fviz_pca_biplot(pca_result_alg, 
                label = "var", 
                habillage = group_var_algarve,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.95) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_alg <- data.frame(pca_result_alg$ind$coord, Group = group_var_algarve)
# PCA plot with groups
scatter_alg <- ggplot(df_pca_alg, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(title = "PCA of pXRF Data", x = "PC1", y = "PC2") +
  theme(legend.position = "right")

## b) PCA of geological cherts: local Lower Jurassic, Central Portugal and Cadiz
pca_db_nl <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
  filter(`Formation/Type` %in% c("LJW", "CPT", "UCA"))
# Convert "< LOD" values to NA
pca_db_nl[pca_db_nl == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_nl$`Formation/Type` <- as.factor(pca_db_nl$`Formation/Type`)  # Ensure it's a factor
group_var_nl <- pca_db_nl$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_nl <- pca_db_nl %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_nl <- complete.cases(df_numeric_nl)  # Identify complete rows
df_numeric_nl <- df_numeric_nl[complete_cases_nl, ]  # Remove NA rows
group_var_nl <- group_var_nl[complete_cases_nl]  # Filter group_var accordingly
# Scale values
df_scaled_nl <- scale(df_numeric_nl)
# Ensure row names are properly set
rownames(df_scaled_nl) <- make.names(group_var_nl, unique = TRUE)
# Perform PCA
pca_result_nl <- PCA(df_scaled_nl, graph = FALSE)
# Summary of PCA
summary(pca_result_nl)
fviz_eig(pca_result_nl, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_nl <- fviz_pca_biplot(pca_result_nl, 
                label = "var", 
                habillage = group_var_nl,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.95) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_nl <- data.frame(pca_result_nl$ind$coord, Group = group_var_nl)
# PCA plot with groups
scatter_nl <- ggplot(df_pca_nl, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(title = "PCA of pXRF Data", x = "PC1", y = "PC2") +
  theme(legend.position = "right")

## c) PCA archaeological T6 compared to local Lower Jurassic and Central Portugal
pca_db_comp <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2",
    `Formation/Type` == "TYPE 3" ~ "T3",
    `Formation/Type` == "TYPE 4" ~ "T4",
    `Formation/Type` == "TYPE 5" ~ "T5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  )) |> 
  filter(`Formation/Type` %in% c("LJW","CPT", "T6"))
# Convert "< LOD" values to NA
pca_db_comp[pca_db_comp == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_comp$`Formation/Type` <- as.factor(pca_db_comp$`Formation/Type`)  # Ensure it's a factor
group_var_comp <- pca_db_comp$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_comp <- pca_db_comp %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_comp <- complete.cases(df_numeric_comp)  # Identify complete rows
df_numeric_comp <- df_numeric_comp[complete_cases_comp, ]  # Remove NA rows
group_var_comp <- group_var_comp[complete_cases_comp]  # Filter group_var accordingly
# Scale values
df_scaled_comp <- scale(df_numeric_comp)
# Ensure row names are properly set
rownames(df_scaled_comp) <- make.names(group_var_comp, unique = TRUE)
# Perform PCA
pca_result_comp <- PCA(df_scaled_comp, graph = FALSE)
# Summary of PCA
summary(pca_result_comp)
fviz_eig(pca_result_comp, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_comp <- fviz_pca_biplot(pca_result_comp, 
                label = "var", 
                habillage = group_var_comp,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.80) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_comp <- data.frame(pca_result_comp$ind$coord, Group = group_var_comp)
# PCA plot with groups
scatter_comp <- ggplot(df_pca_comp, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(title = "PCA of pXRF Data", x = "PC1", y = "PC2") +
  theme(legend.position = "right")

## d) PCA archaeological local cherts T2-5 and non-local T7 compared with local Lower Jurassic and UCA cherts
pca_db_comp2 <- pxrf_raw |> 
  select(`Formation/Type`, P2O5, TiO2, K2O,Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
  mutate(`Formation/Type` = case_when(
    `Formation/Type` == "TYPE 1" ~ "T1",
    `Formation/Type` == "TYPE 2" | `Formation/Type` == "TYPE 2B" | `Formation/Type` == "TYPE 2R" ~ "T2-5",
    `Formation/Type` == "TYPE 3" ~ "T2-5",
    `Formation/Type` == "TYPE 4" ~ "T2-5",
    `Formation/Type` == "TYPE 5" ~ "T2-5",
    `Formation/Type` == "TYPE 6" | `Formation/Type` == "TYPE 6LLB" | `Formation/Type` == "TYPE 6E" ~ "T6",
    `Formation/Type` == "TYPE 7" ~ "T7",
    `Formation/Type` == "TYPE 8" ~ "T8",
    `Formation/Type` == "TYPE 9" ~ "T9",
    `Formation/Type` == "TYPE 10" ~ "T10",
    `Formation/Type` == "TYPE 11" ~ "T11",
    TRUE ~ `Formation/Type`
  )) |> 
  filter(`Formation/Type` %in% c("LJW","UCA", "T7", "T2-5"))
# Convert "< LOD" values to NA
pca_db_comp2[pca_db_comp2 == "< LOD"] <- NA
# Keep Group column separately before PCA
pca_db_comp2$`Formation/Type` <- as.factor(pca_db_comp2$`Formation/Type`)  # Ensure it's a factor
group_var_comp2 <- pca_db_comp2$`Formation/Type`  
# Convert numeric columns correctly
df_numeric_comp2 <- pca_db_comp2 %>%
  select(-`Formation/Type`) %>%  # Remove Group for PCA
  mutate(across(everything(), as.numeric))  # Convert all columns to numeric
# Remove rows with NA (important: also filter group_var!)
complete_cases_comp2 <- complete.cases(df_numeric_comp2)  # Identify complete rows
df_numeric_comp2 <- df_numeric_comp2[complete_cases_comp2, ]  # Remove NA rows
group_var_comp2 <- group_var_comp2[complete_cases_comp2]  # Filter group_var accordingly
# Scale values
df_scaled_comp2 <- scale(df_numeric_comp2)
# Ensure row names are properly set
rownames(df_scaled_comp2) <- make.names(group_var_comp2, unique = TRUE)
# Perform PCA
pca_result_comp2 <- PCA(df_scaled_comp2, graph = FALSE)
# Summary of PCA
summary(pca_result_comp2)
fviz_eig(pca_result_comp2, addlabels = TRUE, ylim = c(0, 100))
# PCA Biplot with Groups
biplot_comp2 <- fviz_pca_biplot(pca_result_comp2, 
                label = "var", 
                habillage = group_var_comp2,  # Use stored group column
                addEllipses = TRUE, 
                ellipse.level = 0.80) +
  ggtitle(NULL)
# Add Group column back for visualization
df_pca_comp2 <- data.frame(pca_result_comp2$ind$coord, Group = group_var_comp2)
# PCA plot with groups
scatter_comp2 <- ggplot(df_pca_comp2, aes(x = Dim.1, y = Dim.2, color = Group, shape = Group)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, aes(fill = Group), alpha = 0.2, geom = "polygon") +
  theme_minimal() +
  labs(x = "PC1", y = "PC2") +
  theme(legend.position = "right")

```

# Introduction

Here is a citation [@Marwick2017]

# Background

# Methods

# Results

All analysed samples are mainly composed of quartz; in all samples, the atomic percentage of oxygen is approximately double of the atomic percentage of silica, showing a 2:1 ratio (table +).

Similarly, the presence of skeletal remains, represented by peaks of calcium, phosphorus and frequently accompanied by small peaks of sulphur, shows no patterns and occurs in samples from all formations and chert types. Although frequently without a discernible shape, oval fossils (although currently unidentified) were observed in LJW geological samples and archaeological samples from local cherts (T2 and T4). This type of large, oval fossil, also composed of calcium, phosphorus and sulphur, seems to be limited to local cherts and to specific areas of nodules, characterised by high amounts of these skeletal remains.

```{r}
#| label: fig-skeletal
#| fig-cap: "Skeletal remain from sample SP16_BLX, from the Lower Jurassic ourcrops of western Algarve. A: Detail of a phosphatized skeletal remain composed by calcium (A1), phosphorus (A2) and a small peak of sulphur (A3). The skeletal remain seems to be oxidized, with small peaks of iron within (A4)."

knitr::include_graphics(here::here("figures/within-chert/M3_3_map1.png"))

```

```{r}
#| label: fig-oval-fossils
#| fig-cap: "A-C: General view and details from an archaeological sample from local chert type (T4). A: general view of a concentration of oval fossils. B: general view of a concentration of oval fossils and other inclusions. C: detail of an irregular inclusion composed of titanium (C1) and an oval phosphatised skeletal remain composed mainly of calcium (C2), phosphorus (C3) and a small peak of sulphur."

knitr::include_graphics(here::here("figures/within-chert/M15_4.png"))

```

By observing the surfaces of samples and the presence of skeletal remains, some differences could be noted between local cherts and non-local archaeological samples. In comparison to the frequent to common skeletal remains in the Jurassic cherts from the Algarve, and archaeological cherts attributed to local sources (T2-5), cherts from non-local sources such as T6-8 show little porosity and rare presence of inclusions. Whenever present, skeletal remains are small and show small peaks (\<3 cpv/ev).

```{r}
#| label: fig-t6-general
#| fig-cap: "A-D: General view and details from an archaeological sample from non-local chert type (T6). A: general view of the sample with uncommon inclusions. B: detail of an inclusion seen in A,  composed mostly of iron and small peaks of calcium and phosphorus, making this inclusion a possible oxidised skeletal remains of a fossil. C: general view of the sample with no visible inclusions. D: detail of a small and irregular iron inclusion."

#knitr::include_graphics(here::here("figures/within-chert/M18_7_M18_8.png"))
M18_7_M18_8_fig <- magick::image_read(here::here("figures/within-chert/M18_7_M18_8.png"))
M18_7_M18_8_gg <- cowplot::ggdraw() + cowplot::draw_image(M18_7_M18_8_fig)

M18_7_M18_8_plots <- plot_grid(M18_7_graph, M18_8_graph, labels = c(1, 2))

plot_grid(M18_7_M18_8_gg, M18_7_M18_8_plots, ncol = 1, rel_heights = c(1, 1))

```

The rarity of inclusions is not limited to skeletal remains, but to other inclusions such as oxides, which in T6-8 are small and rare. Despite this difference in abundance and size, oxides are mainly composed of iron, often associated with other elements, possibly suggesting the existence of iron oxide (Fe2O3), pyrite (FeS2) and ilmenite (FeTiO3). Iron oxides were found both as inclusions scattered in the samples, but also frequently associated with skeletal remains, which may be related to their oxidization. Pyrite was also identified in the shape of spherical pyrite framboids, as well as dispersed pyrite nanocrystals. These are characterised not only by large peaks of iron and associated sulphur, but also of round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids. These were present mainly in geological samples from the Upper Jurassic (both west and eastern formations), but barely identified in archaeological samples.

```{r}
#| label: fig-framboids
#| fig-cap: "A-B: Detail from the geological sample SP59_JOR from the Upper Jurassic outcrops (Jordana formation) of eastern Algarve. A: detailed view of a spherical pyrite framboid composed of iron (A1) and sulphur (A2). B: detailed view of an inclusion composed of iron (B1) and sulphur (B2)."

knitr::include_graphics(here::here("figures/within-chert/M10_5_map1_map2.png"))

```

In a single sample from Lower Jurassic and samples from local T4 (both characterised by the previously mentioned oval fossils) inclusions composed of titanium (without associated iron) were identified, which may be related to the presence of rutile (TiO2). Small peaks of titanium (\<3 cpv/ev) were also identified in 3 out of 4 non-local T6 samples.

Other elements were identified through EDS maps and spectra (e.g., chlorine, potassium or magnesium), although their distribution also showed no patterns between formations and chert types. The exception are calcium and aluminium.

Whenever isolated, calcium can be found in high peaks in cherts from Upper Jurassic formations and is probably related to the presence of calcite inclusions.

```{r}
#| label: fig-calcite
#| fig-cap: "A-B: General view of calcite inclusions (black arrows) from the geological sample SP69_MAR from the Upper Jurassic outcrops of western Algarve."

knitr::include_graphics(here::here("figures/within-chert/M4_3.png"))

```

Aluminium may be related to the speed of sedimentation; high peaks of aluminum may be related to fastest sedimentations speeds and may be a useful component to identify differences between cherts. Almost all chert samples show aluminum peaks, although mostly small (\< 3 cpv/ev). An exception are cherts from the Eastern basin of the Algarve (MJE and UJE), samples from non-local archaeological T7 and T8. The exception to the presence of aluminum is the chalcedony (both geological from TW and archaeological group T1) and all archaeological samples from the non-local group T6, with sources previously attributed to the Cretaceous cherts from Central Portugal.

```{r}
#| label: fig-pca
#| fig-cap: Several pcas

plot_grid(biplot_alg, biplot_nl, biplot_comp, biplot_comp2, labels = c("a", "b", "c", "d"))

```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2025-04-10 analysis/supplementary-materials/appendix1.qmd
Daily work session. Chunks moved to paper and SOM. All SEM-EDS graphs for paper and SOM created.

---
title: "Title Goes Here"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/SOM"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style

---

# Supplementary materials 1

## XRD results

In this subsection we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC. The original data necessary to produce these visualisations can be found and freely accessed in our online compendium, in the associated GitHub project under the path "../data/xrd_data".

```{r}
#| label: somxrd-library

library(readr)
library(dplyr)
library(ggplot2)
library(janitor)
library(tidyr)
library(ggpmisc)

```

```{r}
#| label: somxrd-readtables

# Read grouped supporting data
xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

xrd_marks <- xrd_marks |> # Correct the quartz denomination on base dataset
  dplyr::mutate(Phase = case_when(
    Phase == "SiO2" ~ "Quartz",
    TRUE ~ Phase))

# Read individual xrd files
sp6_xrd <- read_csv('data/xrd_data/SP6_xrd.csv')
sp7_xrd <- read_csv('data/xrd_data/SP7_xrd.csv')
sp10_xrd <- read_csv('data/xrd_data/SP10_xrd.csv')
sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
sp18_xrd <- read_csv('data/xrd_data/SP18_xrd.csv')
sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
sp28_xrd <- read_csv('data/xrd_data/SP28_xrd.csv')
sp30_xrd <- read_csv('data/xrd_data/SP30_xrd.csv')
sp33_xrd <- read_csv('data/xrd_data/SP33_xrd.csv')
sp34_xrd <- read_csv('data/xrd_data/SP34_xrd.csv')
sp40_xrd <- read_csv('data/xrd_data/SP40_xrd.csv')
sp50_xrd <- read_csv('data/xrd_data/SP50_xrd.csv')
sp52_xrd <- read_csv('data/xrd_data/SP52_xrd.csv')
sp57_xrd <- read_csv('data/xrd_data/SP57_xrd.csv')
sp59_xrd <- read_csv('data/xrd_data/SP59_xrd.csv')
sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
sp66_xrd <- read_csv('data/xrd_data/SP66_xrd.csv')

```

```{r}
#| label: fig-xrd-sp6
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP6_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

# SP6 Cabo S. Vicente
sp6_marks <- xrd_marks |>
  filter(Sample == "SP6")
sp6_perc <- xrd_perc |>
  filter(Sample == "SP6") |>
  select(-Sample, -ESD)

sp6_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 5000) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp6_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp6_perc)))

```

```{r}
#| label: fig-xrd-sp7
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Cabo S. Vicente (SP7_CSV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp7_marks <- xrd_marks |>
  filter(Sample == "SP7")
sp7_perc <- xrd_perc |>
  filter(Sample == "SP7") |>
  select(-Sample, -ESD)

sp7_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = `Quartz (SiO2)`) |> 
  rename(Dolomite = `Dolomite (CaMg(CO3)2)`) |> 
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp7_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp7_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp7_perc)))

```

```{r}
#| label: fig-xrd-sp10
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP10_PdA). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp10_marks <- xrd_marks |>
  filter(Sample == "SP10")
sp10_perc <- xrd_perc |>
  filter(Sample == "SP10") |>
  select(-Sample, -ESD)

sp10_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp10_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp10_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp10_perc)))

```

```{r}
#| label: fig-xrd-sp14
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP14_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp14_marks <- xrd_marks |>
  filter(Sample == "SP14")
sp14_perc <- xrd_perc |>
  filter(Sample == "SP14") |>
  select(-Sample, -ESD)

sp14_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp14_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp14_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp14_perc)))

```

```{r}
#| label: fig-xrd-sp18
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP18_PBX). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp18_marks <- xrd_marks |>
  filter(Sample == "SP18")
sp18_perc <- xrd_perc |>
  filter(Sample == "SP18") |>
  select(-Sample, -ESD)

sp18_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp18_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp18_perc)))

```

```{r}
#| label: fig-xrd-sp21
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Belixe Sul, north of Praia do Belixe (SP21_BLS). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp21_marks <- xrd_marks |>
  filter(Sample == "SP21")
sp21_perc <- xrd_perc |>
  filter(Sample == "SP21") |>
  select(-Sample, -ESD)

sp21_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 55000)) +
  geom_point(data = sp21_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp21_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp21_perc)))

```

```{r}
#| label: fig-xrd-sp28
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Aspa, north of Cabo S. Vicente (SP28_ASP). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp28_marks <- xrd_marks |>
  filter(Sample == "SP28")
sp28_perc <- xrd_perc |>
  filter(Sample == "SP28") |>
  select(-Sample, -ESD)

sp28_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp28_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp28_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp28_perc)))

```

```{r}
#| label: fig-xrd-sp30
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP30_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp30_marks <- xrd_marks |>
  filter(Sample == "SP30")
sp30_perc <- xrd_perc |>
  filter(Sample == "SP30") |>
  select(-Sample, -ESD)

sp30_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp30_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp30_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp30_perc)))

```

```{r}
#| label: fig-xrd-sp33
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Foz dos Fornos (SP33_FZF). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp33_marks <- xrd_marks |>
  filter(Sample == "SP33")
sp33_perc <- xrd_perc |>
  filter(Sample == "SP33") |>
  select(-Sample, -ESD)

sp33_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp33_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp33_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp33_perc)))

```

```{r}
#| label: fig-xrd-sp34
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ponta dos Altos (SP34_PdA_a). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp34_marks <- xrd_marks |>
  filter(Sample == "SP34")
sp34_perc <- xrd_perc |>
  filter(Sample == "SP34") |>
  select(-Sample, -ESD)

sp34_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 60000)) +
  geom_point(data = sp34_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp34_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp34_perc)))

```

```{r}
#| label: fig-xrd-sp40
#| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Ferrel (SP40_b). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp40_marks <- xrd_marks |>
  filter(Sample == "SP40")
sp40_perc <- xrd_perc |>
  filter(Sample == "SP40") |>
  select(-Sample, -ESD)

sp40_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  rename(Quartz = SiO2) |> 
  select(-SiO2_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp40_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp40_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp40_perc)))

```

```{r}
#| label: fig-xrd-sp50
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP50_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp50_marks <- xrd_marks |>
  filter(Sample == "SP50")
sp50_perc <- xrd_perc |>
  filter(Sample == "SP50") |>
  select(-Sample, -ESD)

sp50_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp50_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp50_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp50_perc)))

```

```{r}
#| label: fig-xrd-sp52
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Casal da Colina (SP52_CdC). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp52_marks <- xrd_marks |>
  filter(Sample == "SP52")
sp52_perc <- xrd_perc |>
  filter(Sample == "SP52") |>
  select(-Sample, -ESD)

sp52_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Dolomite_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite ,Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp52_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp52_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp52_perc)))

```

```{r}
#| label: fig-xrd-sp57
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formations) chert sample from the outcrop Jordana (SP57_JOR). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp57_marks <- xrd_marks |>
  filter(Sample == "SP57")
sp57_perc <- xrd_perc |>
  filter(Sample == "SP57") |>
  select(-Sample, -ESD)

sp57_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 35000)) +
  geom_point(data = sp57_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp57_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 30000, label = list(cbind(sp57_perc)))

```

```{r}
#| label: fig-xrd-sp59
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP59_JOR). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp59_marks <- xrd_marks |>
  filter(Sample == "SP59")
sp59_perc <- xrd_perc |>
  filter(Sample == "SP59") |>
  select(-Sample, -ESD)

sp59_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 35000)) +
  geom_point(data = sp59_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp59_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 30000, label = list(cbind(sp59_perc)))

```

```{r}
#| label: fig-xrd-sp61
#| fig-cap: Plot of minerals identified through XRD in the Upper Jurassic (Jordana formation) chert sample from the secondary outcrop Peral (SP61_PER). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp61_marks <- xrd_marks |>
  filter(Sample == "SP61")
sp61_perc <- xrd_perc |>
  filter(Sample == "SP61") |>
  select(-Sample, -ESD)

sp61_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 65000)) +
  geom_point(data = sp61_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp61_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp61_perc)))

```

```{r}
#| label: fig-xrd-sp63
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Oliveiras (SP63_OLV). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp63_marks <- xrd_marks |>
  filter(Sample == "SP63")
sp63_perc <- xrd_perc |>
  filter(Sample == "SP63") |>
  select(-Sample, -ESD)

sp63_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp63_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp63_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp63_perc)))

```

```{r}
#| label: fig-xrd-sp66
#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Malhão (SP66_MALH). The rhombus shape signifies main peaks key to identify each mineral. Labels represent the d spacing values of key mineral peaks.

sp66_marks <- xrd_marks |>
  filter(Sample == "SP66")
sp66_perc <- xrd_perc |>
  filter(Sample == "SP66") |>
  select(-Sample, -ESD)

sp66_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000) |>
  select(-Quartz_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
               names_to = "Phase", 
               values_to = "Count") |> 
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 70000)) +
  geom_point(data = sp66_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp66_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 65000, label = list(cbind(sp66_perc)))

```

## SEM-EDS results

```{r}
#| label: sem-read-som

M2_3_map2 <- read_csv('analysis/data/raw_data/sem_data/M2_3_map2.csv')
M3_1 <- read_csv('analysis/data/raw_data/sem_data/M3_1_map1_map2.csv')
M3_3_s1 <- read_csv('analysis/data/raw_data/sem_data/M3_3_s1.csv')
M3_3_s2 <- read_csv('analysis/data/raw_data/sem_data/M3_3_s2.csv')
M4_2_s5_1 <- read_csv('analysis/data/raw_data/sem_data/M4_2_s5_1.csv')
M4_2_s5_4 <- read_csv('analysis/data/raw_data/sem_data/M4_2_s5_4.csv')
M8_2 <- read_csv('analysis/data/raw_data/sem_data/M8_2.csv')
M10_4 <- read_csv('analysis/data/raw_data/sem_data/M10_4.csv')
M11_1 <- read_csv('analysis/data/raw_data/sem_data/M11_1.csv')
M12_1_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M12_1_s2_1.csv')
M12_1_s3_1 <- read_csv('analysis/data/raw_data/sem_data/M12_1_s3_1.csv')
M12_3_s3_2 <- read_csv('analysis/data/raw_data/sem_data/M12_3_s3_2.csv')
M12_5_s1_1 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s1_1.csv')
M12_5_s1_3 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s1_3.csv')
M12_5_s3_3 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s3_3.csv')
M12_5_s6_1 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s6_1.csv')
M12_5_s6_2 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s6_2.csv')
M13_1_3_s2 <- read_csv('analysis/data/raw_data/sem_data/M13_1_3_s2.csv')
M14_3_2_s1_1 <- read_csv('analysis/data/raw_data/sem_data/M14_3_2_s1_1.csv')
M14_3_2_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M14_3_2_s2_1.csv')
M14_4_1 <- read_csv('analysis/data/raw_data/sem_data/M14_4_1.csv')
M15_2_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M15_2_s2_1.csv')
M15_2_s2_2 <- read_csv('analysis/data/raw_data/sem_data/M15_2_s2_2.csv')
M16_1_1 <- read_csv('analysis/data/raw_data/sem_data/M16_1_1.csv')
M18_1_1 <- read_csv('analysis/data/raw_data/sem_data/M18_1_1.csv')
M18_5_1 <- read_csv('analysis/data/raw_data/sem_data/M18_5_1.csv')
M20_2_1 <- read_csv('analysis/data/raw_data/sem_data/M20_2_1.csv')
M22_4.1_1 <- read_csv('analysis/data/raw_data/sem_data/M22_4.1_1.csv')
M22_5.1_1 <- read_csv('analysis/data/raw_data/sem_data/M22_5.1_1.csv')
M22_5.1_3 <- read_csv('analysis/data/raw_data/sem_data/M22_5.1_3.csv')
M23_2_1 <- read_csv('analysis/data/raw_data/sem_data/M23_2_1.csv')
M23_7_1 <- read_csv('analysis/data/raw_data/sem_data/M23_7_1.csv')
M24_1_1 <- read_csv('analysis/data/raw_data/sem_data/M24_1_1.csv')
M24_1_2 <- read_csv('analysis/data/raw_data/sem_data/M24_1_2.csv')

```

```{r}
#| label: sem-graphs-som

# Prepare the energy ranges
energy_ranges <- list(
  C = c(0.2, 0.3),
  O = c(0.5, 0.6),
  F = c(0.67, 0.69),
  Na = c(1.04, 1.06),
  Mg = c(1.25, 1.27),
  Mg_beta = c(1.3, 1.4),
  Al = c(1.48, 1.5),
  Si = c(1.7, 1.8),
  P = c(2, 2.1),
  S = c(2.3, 2.32),
  S_beta = c(2.46, 2.47),
  Cl = c(2.62, 2.64),
  K = c(3.31, 3.33),
  K_beta = c(3.59, 3.61),
  Ca = c(3.68, 3.7),
  Ca_beta = c(4.01, 4.02),
  Ca_L = c(0.34, 0.36),
  Ti = c(4.5, 4.6),
  Ti_beta = c(4.9, 5),
  Ti_L = c(0.45, 0.47),
  Fe = c(6.38, 6.41),
  Fe_beta = c(7, 7.2),
  Fe_L = c(0.69, 0.73))

# M24_1_2
filtered_M24_1_2 <- subset(M24_1_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M24_1_2 <- list()
for (element in names(energy_ranges)) {
  range_M24_1_2 <- energy_ranges[[element]]
  subset_M24_1_2 <- subset(filtered_M24_1_2, Energy >= range_M24_1_2[1] & Energy <= range_M24_1_2[2])
  if (nrow(subset_M24_1_2) > 0) {
    max_row_M24_1_2 <- subset_M24_1_2[which.max(subset_M24_1_2$Counts), ]
    peaks_list_M24_1_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M24_1_2$Energy,
      Counts = max_row_M24_1_2$Counts
    )
  }
}

peaks_M24_1_2 <- do.call(rbind, peaks_list_M24_1_2)
rownames(peaks_M24_1_2) <- NULL
peaks_M24_1_2 <- peaks_M24_1_2 |> 
  filter(Counts > 250) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S_beta", "S", "Cl"))

M24_1_2_graph <- filtered_M24_1_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M24_1_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M24_1_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M24_1_1
filtered_M24_1_1 <- subset(M24_1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M24_1_1 <- list()
for (element in names(energy_ranges)) {
  range_M24_1_1 <- energy_ranges[[element]]
  subset_M24_1_1 <- subset(filtered_M24_1_1, Energy >= range_M24_1_1[1] & Energy <= range_M24_1_1[2])
  if (nrow(subset_M24_1_1) > 0) {
    max_row_M24_1_1 <- subset_M24_1_1[which.max(subset_M24_1_1$Counts), ]
    peaks_list_M24_1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M24_1_1$Energy,
      Counts = max_row_M24_1_1$Counts
    )
  }
}

peaks_M24_1_1 <- do.call(rbind, peaks_list_M24_1_1)
rownames(peaks_M24_1_1) <- NULL
peaks_M24_1_1 <- peaks_M24_1_1 |> 
  filter(Counts > 300) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "P", "S", "Cl"))

M24_1_1_graph <- filtered_M24_1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M24_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M24_1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M23_7_1
filtered_M23_7_1 <- subset(M23_7_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M23_7_1 <- list()
for (element in names(energy_ranges)) {
  range_M23_7_1 <- energy_ranges[[element]]
  subset_M23_7_1 <- subset(filtered_M23_7_1, Energy >= range_M23_7_1[1] & Energy <= range_M23_7_1[2])
  if (nrow(subset_M23_7_1) > 0) {
    max_row_M23_7_1 <- subset_M23_7_1[which.max(subset_M23_7_1$Counts), ]
    peaks_list_M23_7_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M23_7_1$Energy,
      Counts = max_row_M23_7_1$Counts
    )
  }
}

peaks_M23_7_1 <- do.call(rbind, peaks_list_M23_7_1)
rownames(peaks_M23_7_1) <- NULL
peaks_M23_7_1 <- peaks_M23_7_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M23_7_1_graph <- filtered_M23_7_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M23_7_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M23_7_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M23_2_1
filtered_M23_2_1 <- subset(M23_2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M23_2_1 <- list()
for (element in names(energy_ranges)) {
  range_M23_2_1 <- energy_ranges[[element]]
  subset_M23_2_1 <- subset(filtered_M23_2_1, Energy >= range_M23_2_1[1] & Energy <= range_M23_2_1[2])
  if (nrow(subset_M23_2_1) > 0) {
    max_row_M23_2_1 <- subset_M23_2_1[which.max(subset_M23_2_1$Counts), ]
    peaks_list_M23_2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M23_2_1$Energy,
      Counts = max_row_M23_2_1$Counts
    )
  }
}

peaks_M23_2_1 <- do.call(rbind, peaks_list_M23_2_1)
rownames(peaks_M23_2_1) <- NULL
peaks_M23_2_1 <- peaks_M23_2_1 |> 
  filter(Counts > 70) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "K_beta"))

M23_2_1_graph <- filtered_M23_2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M23_2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M23_2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M22_5.1_3
filtered_M22_5.1_3 <- subset(M22_5.1_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M22_5.1_3 <- list()
for (element in names(energy_ranges)) {
  range_M22_5.1_3 <- energy_ranges[[element]]
  subset_M22_5.1_3 <- subset(filtered_M22_5.1_3, Energy >= range_M22_5.1_3[1] & Energy <= range_M22_5.1_3[2])
  if (nrow(subset_M22_5.1_3) > 0) {
    max_row_M22_5.1_3 <- subset_M22_5.1_3[which.max(subset_M22_5.1_3$Counts), ]
    peaks_list_M22_5.1_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M22_5.1_3$Energy,
      Counts = max_row_M22_5.1_3$Counts
    )
  }
}

peaks_M22_5.1_3 <- do.call(rbind, peaks_list_M22_5.1_3)
rownames(peaks_M22_5.1_3) <- NULL
peaks_M22_5.1_3 <- peaks_M22_5.1_3 |> 
  filter(Counts > 55) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M22_5.1_3_graph <- filtered_M22_5.1_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M22_5.1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M22_5.1_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M22_5.1_1
filtered_M22_5.1_1 <- subset(M22_5.1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M22_5.1_1 <- list()
for (element in names(energy_ranges)) {
  range_M22_5.1_1 <- energy_ranges[[element]]
  subset_M22_5.1_1 <- subset(filtered_M22_5.1_1, Energy >= range_M22_5.1_1[1] & Energy <= range_M22_5.1_1[2])
  if (nrow(subset_M22_5.1_1) > 0) {
    max_row_M22_5.1_1 <- subset_M22_5.1_1[which.max(subset_M22_5.1_1$Counts), ]
    peaks_list_M22_5.1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M22_5.1_1$Energy,
      Counts = max_row_M22_5.1_1$Counts
    )
  }
}

peaks_M22_5.1_1 <- do.call(rbind, peaks_list_M22_5.1_1)
rownames(peaks_M22_5.1_1) <- NULL
peaks_M22_5.1_1 <- peaks_M22_5.1_1 |> 
  filter(Counts > 70) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S"))

M22_5.1_1_graph <- filtered_M22_5.1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M22_5.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M22_5.1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M22_4.1_1
filtered_M22_4.1_1 <- subset(M22_4.1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M22_4.1_1 <- list()
for (element in names(energy_ranges)) {
  range_M22_4.1_1 <- energy_ranges[[element]]
  subset_M22_4.1_1 <- subset(filtered_M22_4.1_1, Energy >= range_M22_4.1_1[1] & Energy <= range_M22_4.1_1[2])
  if (nrow(subset_M22_4.1_1) > 0) {
    max_row_M22_4.1_1 <- subset_M22_4.1_1[which.max(subset_M22_4.1_1$Counts), ]
    peaks_list_M22_4.1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M22_4.1_1$Energy,
      Counts = max_row_M22_4.1_1$Counts
    )
  }
}

peaks_M22_4.1_1 <- do.call(rbind, peaks_list_M22_4.1_1)
rownames(peaks_M22_4.1_1) <- NULL
peaks_M22_4.1_1 <- peaks_M22_4.1_1 |> 
  filter(Counts > 50) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M22_4.1_1_graph <- filtered_M22_4.1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M22_4.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M22_4.1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M20_2_1
filtered_M20_2_1 <- subset(M20_2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M20_2_1 <- list()
for (element in names(energy_ranges)) {
  range_M20_2_1 <- energy_ranges[[element]]
  subset_M20_2_1 <- subset(filtered_M20_2_1, Energy >= range_M20_2_1[1] & Energy <= range_M20_2_1[2])
  if (nrow(subset_M20_2_1) > 0) {
    max_row_M20_2_1 <- subset_M20_2_1[which.max(subset_M20_2_1$Counts), ]
    peaks_list_M20_2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M20_2_1$Energy,
      Counts = max_row_M20_2_1$Counts
    )
  }
}

peaks_M20_2_1 <- do.call(rbind, peaks_list_M20_2_1)
rownames(peaks_M20_2_1) <- NULL
peaks_M20_2_1 <- peaks_M20_2_1 |> 
  filter(Counts > 400) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S", "Mg", "S_beta", "Ca_beta", "Ti_beta", "K_beta", "K"))

M20_2_1_graph <- filtered_M20_2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M20_2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M20_2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M18_5_1
filtered_M18_5_1 <- subset(M18_5_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_5_1 <- list()
for (element in names(energy_ranges)) {
  range_M18_5_1 <- energy_ranges[[element]]
  subset_M18_5_1 <- subset(filtered_M18_5_1, Energy >= range_M18_5_1[1] & Energy <= range_M18_5_1[2])
  if (nrow(subset_M18_5_1) > 0) {
    max_row_M18_5_1 <- subset_M18_5_1[which.max(subset_M18_5_1$Counts), ]
    peaks_list_M18_5_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_5_1$Energy,
      Counts = max_row_M18_5_1$Counts
    )
  }
}

peaks_M18_5_1 <- do.call(rbind, peaks_list_M18_5_1)
rownames(peaks_M18_5_1) <- NULL
peaks_M18_5_1 <- peaks_M18_5_1 |> 
  filter(Counts > 150) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M18_5_1_graph <- filtered_M18_5_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M18_5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M18_5_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M18_1_1
filtered_M18_1_1 <- subset(M18_1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M18_1_1 <- list()
for (element in names(energy_ranges)) {
  range_M18_1_1 <- energy_ranges[[element]]
  subset_M18_1_1 <- subset(filtered_M18_1_1, Energy >= range_M18_1_1[1] & Energy <= range_M18_1_1[2])
  if (nrow(subset_M18_1_1) > 0) {
    max_row_M18_1_1 <- subset_M18_1_1[which.max(subset_M18_1_1$Counts), ]
    peaks_list_M18_1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M18_1_1$Energy,
      Counts = max_row_M18_1_1$Counts
    )
  }
}

peaks_M18_1_1 <- do.call(rbind, peaks_list_M18_1_1)
rownames(peaks_M18_1_1) <- NULL
peaks_M18_1_1 <- peaks_M18_1_1 |> 
  filter(Counts > 200) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M18_1_1_graph <- filtered_M18_1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M18_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M18_1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M16_1_1
filtered_M16_1_1 <- subset(M16_1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M16_1_1 <- list()
for (element in names(energy_ranges)) {
  range_M16_1_1 <- energy_ranges[[element]]
  subset_M16_1_1 <- subset(filtered_M16_1_1, Energy >= range_M16_1_1[1] & Energy <= range_M16_1_1[2])
  if (nrow(subset_M16_1_1) > 0) {
    max_row_M16_1_1 <- subset_M16_1_1[which.max(subset_M16_1_1$Counts), ]
    peaks_list_M16_1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M16_1_1$Energy,
      Counts = max_row_M16_1_1$Counts
    )
  }
}

peaks_M16_1_1 <- do.call(rbind, peaks_list_M16_1_1)
rownames(peaks_M16_1_1) <- NULL
peaks_M16_1_1 <- peaks_M16_1_1 |> 
  filter(Counts > 600) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S_beta"))

M16_1_1_graph <- filtered_M16_1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M16_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M16_1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M15_2_s2_2
filtered_M15_2_s2_2 <- subset(M15_2_s2_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M15_2_s2_2 <- list()
for (element in names(energy_ranges)) {
  range_M15_2_s2_2 <- energy_ranges[[element]]
  subset_M15_2_s2_2 <- subset(filtered_M15_2_s2_2, Energy >= range_M15_2_s2_2[1] & Energy <= range_M15_2_s2_2[2])
  if (nrow(subset_M15_2_s2_2) > 0) {
    max_row_M15_2_s2_2 <- subset_M15_2_s2_2[which.max(subset_M15_2_s2_2$Counts), ]
    peaks_list_M15_2_s2_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M15_2_s2_2$Energy,
      Counts = max_row_M15_2_s2_2$Counts
    )
  }
}

peaks_M15_2_s2_2 <- do.call(rbind, peaks_list_M15_2_s2_2)
rownames(peaks_M15_2_s2_2) <- NULL
peaks_M15_2_s2_2 <- peaks_M15_2_s2_2 |> 
  filter(Counts > 300) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "Na", "P", "S", "Cl", "K", "Ca_beta", "S_beta", "Ca", "K_beta", "Ca_L"))

M15_2_s2_2_graph <- filtered_M15_2_s2_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M15_2_s2_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M15_2_s2_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M15_2_s2_1
filtered_M15_2_s2_1 <- subset(M15_2_s2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M15_2_s2_1 <- list()
for (element in names(energy_ranges)) {
  range_M15_2_s2_1 <- energy_ranges[[element]]
  subset_M15_2_s2_1 <- subset(filtered_M15_2_s2_1, Energy >= range_M15_2_s2_1[1] & Energy <= range_M15_2_s2_1[2])
  if (nrow(subset_M15_2_s2_1) > 0) {
    max_row_M15_2_s2_1 <- subset_M15_2_s2_1[which.max(subset_M15_2_s2_1$Counts), ]
    peaks_list_M15_2_s2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M15_2_s2_1$Energy,
      Counts = max_row_M15_2_s2_1$Counts
    )
  }
}

peaks_M15_2_s2_1 <- do.call(rbind, peaks_list_M15_2_s2_1)
rownames(peaks_M15_2_s2_1) <- NULL
peaks_M15_2_s2_1 <- peaks_M15_2_s2_1 |> 
  filter(Counts > 500) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "K_beta"))

M15_2_s2_1_graph <- filtered_M15_2_s2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M15_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M15_2_s2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M14_4_1
filtered_M14_4_1 <- subset(M14_4_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M14_4_1 <- list()
for (element in names(energy_ranges)) {
  range_M14_4_1 <- energy_ranges[[element]]
  subset_M14_4_1 <- subset(filtered_M14_4_1, Energy >= range_M14_4_1[1] & Energy <= range_M14_4_1[2])
  if (nrow(subset_M14_4_1) > 0) {
    max_row_M14_4_1 <- subset_M14_4_1[which.max(subset_M14_4_1$Counts), ]
    peaks_list_M14_4_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M14_4_1$Energy,
      Counts = max_row_M14_4_1$Counts
    )
  }
}

peaks_M14_4_1 <- do.call(rbind, peaks_list_M14_4_1)
rownames(peaks_M14_4_1) <- NULL
peaks_M14_4_1 <- peaks_M14_4_1 |> 
  filter(Counts > 700) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M14_4_1_graph <- filtered_M14_4_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M14_4_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M14_4_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M14_3_2_s2_1
filtered_M14_3_2_s2_1 <- subset(M14_3_2_s2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M14_3_2_s2_1 <- list()
for (element in names(energy_ranges)) {
  range_M14_3_2_s2_1 <- energy_ranges[[element]]
  subset_M14_3_2_s2_1 <- subset(filtered_M14_3_2_s2_1, Energy >= range_M14_3_2_s2_1[1] & Energy <= range_M14_3_2_s2_1[2])
  if (nrow(subset_M14_3_2_s2_1) > 0) {
    max_row_M14_3_2_s2_1 <- subset_M14_3_2_s2_1[which.max(subset_M14_3_2_s2_1$Counts), ]
    peaks_list_M14_3_2_s2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M14_3_2_s2_1$Energy,
      Counts = max_row_M14_3_2_s2_1$Counts
    )
  }
}

peaks_M14_3_2_s2_1 <- do.call(rbind, peaks_list_M14_3_2_s2_1)
rownames(peaks_M14_3_2_s2_1) <- NULL
peaks_M14_3_2_s2_1 <- peaks_M14_3_2_s2_1 |> 
  filter(Counts > 60) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M14_3_2_s2_1_graph <- filtered_M14_3_2_s2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M14_3_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M14_3_2_s2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M14_3_2_s1_1
filtered_M14_3_2_s1_1 <- subset(M14_3_2_s1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M14_3_2_s1_1 <- list()
for (element in names(energy_ranges)) {
  range_M14_3_2_s1_1 <- energy_ranges[[element]]
  subset_M14_3_2_s1_1 <- subset(filtered_M14_3_2_s1_1, Energy >= range_M14_3_2_s1_1[1] & Energy <= range_M14_3_2_s1_1[2])
  if (nrow(subset_M14_3_2_s1_1) > 0) {
    max_row_M14_3_2_s1_1 <- subset_M14_3_2_s1_1[which.max(subset_M14_3_2_s1_1$Counts), ]
    peaks_list_M14_3_2_s1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M14_3_2_s1_1$Energy,
      Counts = max_row_M14_3_2_s1_1$Counts
    )
  }
}

peaks_M14_3_2_s1_1 <- do.call(rbind, peaks_list_M14_3_2_s1_1)
rownames(peaks_M14_3_2_s1_1) <- NULL
peaks_M14_3_2_s1_1 <- peaks_M14_3_2_s1_1 |> 
  filter(Counts > 500) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M14_3_2_s1_1_graph <- filtered_M14_3_2_s1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M14_3_2_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M14_3_2_s1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M13_1_3_s2
filtered_M13_1_3_s2 <- subset(M13_1_3_s2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M13_1_3_s2 <- list()
for (element in names(energy_ranges)) {
  range_M13_1_3_s2 <- energy_ranges[[element]]
  subset_M13_1_3_s2 <- subset(filtered_M13_1_3_s2, Energy >= range_M13_1_3_s2[1] & Energy <= range_M13_1_3_s2[2])
  if (nrow(subset_M13_1_3_s2) > 0) {
    max_row_M13_1_3_s2 <- subset_M13_1_3_s2[which.max(subset_M13_1_3_s2$Counts), ]
    peaks_list_M13_1_3_s2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M13_1_3_s2$Energy,
      Counts = max_row_M13_1_3_s2$Counts
    )
  }
}

peaks_M13_1_3_s2 <- do.call(rbind, peaks_list_M13_1_3_s2)
rownames(peaks_M13_1_3_s2) <- NULL
peaks_M13_1_3_s2 <- peaks_M13_1_3_s2 |> 
  filter(Counts > 300) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M13_1_3_s2_graph <- filtered_M13_1_3_s2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M13_1_3_s2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M13_1_3_s2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_5_s6_2
filtered_M12_5_s6_2 <- subset(M12_5_s6_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s6_2 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s6_2 <- energy_ranges[[element]]
  subset_M12_5_s6_2 <- subset(filtered_M12_5_s6_2, Energy >= range_M12_5_s6_2[1] & Energy <= range_M12_5_s6_2[2])
  if (nrow(subset_M12_5_s6_2) > 0) {
    max_row_M12_5_s6_2 <- subset_M12_5_s6_2[which.max(subset_M12_5_s6_2$Counts), ]
    peaks_list_M12_5_s6_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s6_2$Energy,
      Counts = max_row_M12_5_s6_2$Counts
    )
  }
}

peaks_M12_5_s6_2 <- do.call(rbind, peaks_list_M12_5_s6_2)
rownames(peaks_M12_5_s6_2) <- NULL
peaks_M12_5_s6_2 <- peaks_M12_5_s6_2 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_5_s6_2_graph <- filtered_M12_5_s6_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s6_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_5_s6_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_5_s6_1
filtered_M12_5_s6_1 <- subset(M12_5_s6_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s6_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s6_1 <- energy_ranges[[element]]
  subset_M12_5_s6_1 <- subset(filtered_M12_5_s6_1, Energy >= range_M12_5_s6_1[1] & Energy <= range_M12_5_s6_1[2])
  if (nrow(subset_M12_5_s6_1) > 0) {
    max_row_M12_5_s6_1 <- subset_M12_5_s6_1[which.max(subset_M12_5_s6_1$Counts), ]
    peaks_list_M12_5_s6_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s6_1$Energy,
      Counts = max_row_M12_5_s6_1$Counts
    )
  }
}

peaks_M12_5_s6_1 <- do.call(rbind, peaks_list_M12_5_s6_1)
rownames(peaks_M12_5_s6_1) <- NULL
peaks_M12_5_s6_1 <- peaks_M12_5_s6_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_5_s6_1_graph <- filtered_M12_5_s6_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s6_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_5_s6_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_5_s3_3
filtered_M12_5_s3_3 <- subset(M12_5_s3_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s3_3 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s3_3 <- energy_ranges[[element]]
  subset_M12_5_s3_3 <- subset(filtered_M12_5_s3_3, Energy >= range_M12_5_s3_3[1] & Energy <= range_M12_5_s3_3[2])
  if (nrow(subset_M12_5_s3_3) > 0) {
    max_row_M12_5_s3_3 <- subset_M12_5_s3_3[which.max(subset_M12_5_s3_3$Counts), ]
    peaks_list_M12_5_s3_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s3_3$Energy,
      Counts = max_row_M12_5_s3_3$Counts
    )
  }
}

peaks_M12_5_s3_3 <- do.call(rbind, peaks_list_M12_5_s3_3)
rownames(peaks_M12_5_s3_3) <- NULL
peaks_M12_5_s3_3 <- peaks_M12_5_s3_3 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_5_s3_3_graph <- filtered_M12_5_s3_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s3_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_5_s3_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_5_s1_3
filtered_M12_5_s1_3 <- subset(M12_5_s1_3, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s1_3 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s1_3 <- energy_ranges[[element]]
  subset_M12_5_s1_3 <- subset(filtered_M12_5_s1_3, Energy >= range_M12_5_s1_3[1] & Energy <= range_M12_5_s1_3[2])
  if (nrow(subset_M12_5_s1_3) > 0) {
    max_row_M12_5_s1_3 <- subset_M12_5_s1_3[which.max(subset_M12_5_s1_3$Counts), ]
    peaks_list_M12_5_s1_3[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s1_3$Energy,
      Counts = max_row_M12_5_s1_3$Counts
    )
  }
}

peaks_M12_5_s1_3 <- do.call(rbind, peaks_list_M12_5_s1_3)
rownames(peaks_M12_5_s1_3) <- NULL
peaks_M12_5_s1_3 <- peaks_M12_5_s1_3 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_5_s1_3_graph <- filtered_M12_5_s1_3 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_5_s1_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_5_s1_1
filtered_M12_5_s1_1 <- subset(M12_5_s1_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_5_s1_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_5_s1_1 <- energy_ranges[[element]]
  subset_M12_5_s1_1 <- subset(filtered_M12_5_s1_1, Energy >= range_M12_5_s1_1[1] & Energy <= range_M12_5_s1_1[2])
  if (nrow(subset_M12_5_s1_1) > 0) {
    max_row_M12_5_s1_1 <- subset_M12_5_s1_1[which.max(subset_M12_5_s1_1$Counts), ]
    peaks_list_M12_5_s1_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_5_s1_1$Energy,
      Counts = max_row_M12_5_s1_1$Counts
    )
  }
}

peaks_M12_5_s1_1 <- do.call(rbind, peaks_list_M12_5_s1_1)
rownames(peaks_M12_5_s1_1) <- NULL
peaks_M12_5_s1_1 <- peaks_M12_5_s1_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L","Mg_beta", "F"))

M12_5_s1_1_graph <- filtered_M12_5_s1_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_5_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_5_s1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_3_s3_2
filtered_M12_3_s3_2 <- subset(M12_3_s3_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_3_s3_2 <- list()
for (element in names(energy_ranges)) {
  range_M12_3_s3_2 <- energy_ranges[[element]]
  subset_M12_3_s3_2 <- subset(filtered_M12_3_s3_2, Energy >= range_M12_3_s3_2[1] & Energy <= range_M12_3_s3_2[2])
  if (nrow(subset_M12_3_s3_2) > 0) {
    max_row_M12_3_s3_2 <- subset_M12_3_s3_2[which.max(subset_M12_3_s3_2$Counts), ]
    peaks_list_M12_3_s3_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_3_s3_2$Energy,
      Counts = max_row_M12_3_s3_2$Counts
    )
  }
}

peaks_M12_3_s3_2 <- do.call(rbind, peaks_list_M12_3_s3_2)
rownames(peaks_M12_3_s3_2) <- NULL
peaks_M12_3_s3_2 <- peaks_M12_3_s3_2 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_3_s3_2_graph <- filtered_M12_3_s3_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_3_s3_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_3_s3_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_1_s3_1
filtered_M12_1_s3_1 <- subset(M12_1_s3_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_1_s3_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_1_s3_1 <- energy_ranges[[element]]
  subset_M12_1_s3_1 <- subset(filtered_M12_1_s3_1, Energy >= range_M12_1_s3_1[1] & Energy <= range_M12_1_s3_1[2])
  if (nrow(subset_M12_1_s3_1) > 0) {
    max_row_M12_1_s3_1 <- subset_M12_1_s3_1[which.max(subset_M12_1_s3_1$Counts), ]
    peaks_list_M12_1_s3_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_1_s3_1$Energy,
      Counts = max_row_M12_1_s3_1$Counts
    )
  }
}

peaks_M12_1_s3_1 <- do.call(rbind, peaks_list_M12_1_s3_1)
rownames(peaks_M12_1_s3_1) <- NULL
peaks_M12_1_s3_1 <- peaks_M12_1_s3_1 |> 
  filter(Counts > 90) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_1_s3_1_graph <- filtered_M12_1_s3_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_1_s3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_1_s3_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M12_1_s2_1
filtered_M12_1_s2_1 <- subset(M12_1_s2_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M12_1_s2_1 <- list()
for (element in names(energy_ranges)) {
  range_M12_1_s2_1 <- energy_ranges[[element]]
  subset_M12_1_s2_1 <- subset(filtered_M12_1_s2_1, Energy >= range_M12_1_s2_1[1] & Energy <= range_M12_1_s2_1[2])
  if (nrow(subset_M12_1_s2_1) > 0) {
    max_row_M12_1_s2_1 <- subset_M12_1_s2_1[which.max(subset_M12_1_s2_1$Counts), ]
    peaks_list_M12_1_s2_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M12_1_s2_1$Energy,
      Counts = max_row_M12_1_s2_1$Counts
    )
  }
}

peaks_M12_1_s2_1 <- do.call(rbind, peaks_list_M12_1_s2_1)
rownames(peaks_M12_1_s2_1) <- NULL
peaks_M12_1_s2_1 <- peaks_M12_1_s2_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M12_1_s2_1_graph <- filtered_M12_1_s2_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M12_1_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M12_1_s2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M2_3_map2
filtered_M2_3_map2 <- subset(M2_3_map2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M2_3_map2 <- list()
for (element in names(energy_ranges)) {
  range_M2_3_map2 <- energy_ranges[[element]]
  subset_M2_3_map2 <- subset(filtered_M2_3_map2, Energy >= range_M2_3_map2[1] & Energy <= range_M2_3_map2[2])
  if (nrow(subset_M2_3_map2) > 0) {
    max_row_M2_3_map2 <- subset_M2_3_map2[which.max(subset_M2_3_map2$Counts), ]
    peaks_list_M2_3_map2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M2_3_map2$Energy,
      Counts = max_row_M2_3_map2$Counts
    )
  }
}

peaks_M2_3_map2 <- do.call(rbind, peaks_list_M2_3_map2)
rownames(peaks_M2_3_map2) <- NULL
peaks_M2_3_map2 <- peaks_M2_3_map2 |> 
  filter(Counts > 100)

M2_3_map2_graph <- filtered_M2_3_map2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M2_3_map2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M2_3_map2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M3_1
filtered_M3_1 <- subset(M3_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M3_1 <- list()
for (element in names(energy_ranges)) {
  range_M3_1 <- energy_ranges[[element]]
  subset_M3_1 <- subset(filtered_M3_1, Energy >= range_M3_1[1] & Energy <= range_M3_1[2])
  if (nrow(subset_M3_1) > 0) {
    max_row_M3_1 <- subset_M3_1[which.max(subset_M3_1$Counts), ]
    peaks_list_M3_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M3_1$Energy,
      Counts = max_row_M3_1$Counts
    )
  }
}

peaks_M3_1 <- do.call(rbind, peaks_list_M3_1)
rownames(peaks_M3_1) <- NULL
peaks_M3_1 <- peaks_M3_1 |> 
  filter(Counts > 100) |> 
  filter(Element != "Ti_L")

M3_1_graph <- filtered_M3_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M3_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M3_3_s1
filtered_M3_3_s1 <- subset(M3_3_s1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M3_3_s1 <- list()
for (element in names(energy_ranges)) {
  range_M3_3_s1 <- energy_ranges[[element]]
  subset_M3_3_s1 <- subset(filtered_M3_3_s1, Energy >= range_M3_3_s1[1] & Energy <= range_M3_3_s1[2])
  if (nrow(subset_M3_3_s1) > 0) {
    max_row_M3_3_s1 <- subset_M3_3_s1[which.max(subset_M3_3_s1$Counts), ]
    peaks_list_M3_3_s1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M3_3_s1$Energy,
      Counts = max_row_M3_3_s1$Counts
    )
  }
}

peaks_M3_3_s1 <- do.call(rbind, peaks_list_M3_3_s1)
rownames(peaks_M3_3_s1) <- NULL
peaks_M3_3_s1 <- peaks_M3_3_s1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta"))

M3_3_s1_graph <- filtered_M3_3_s1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M3_3_s1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M3_3_s1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M3_3_s2
filtered_M3_3_s2 <- subset(M3_3_s2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M3_3_s2 <- list()
for (element in names(energy_ranges)) {
  range_M3_3_s2 <- energy_ranges[[element]]
  subset_M3_3_s2 <- subset(filtered_M3_3_s2, Energy >= range_M3_3_s2[1] & Energy <= range_M3_3_s2[2])
  if (nrow(subset_M3_3_s2) > 0) {
    max_row_M3_3_s2 <- subset_M3_3_s2[which.max(subset_M3_3_s2$Counts), ]
    peaks_list_M3_3_s2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M3_3_s2$Energy,
      Counts = max_row_M3_3_s2$Counts
    )
  }
}

peaks_M3_3_s2 <- do.call(rbind, peaks_list_M3_3_s2)
rownames(peaks_M3_3_s2) <- NULL
peaks_M3_3_s2 <- peaks_M3_3_s2 |> 
  filter(Counts > 90) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M3_3_s2_graph <- filtered_M3_3_s2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M3_3_s2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M3_3_s2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M4_2_s5_1
filtered_M4_2_s5_1 <- subset(M4_2_s5_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M4_2_s5_1 <- list()
for (element in names(energy_ranges)) {
  range_M4_2_s5_1 <- energy_ranges[[element]]
  subset_M4_2_s5_1 <- subset(filtered_M4_2_s5_1, Energy >= range_M4_2_s5_1[1] & Energy <= range_M4_2_s5_1[2])
  if (nrow(subset_M4_2_s5_1) > 0) {
    max_row_M4_2_s5_1 <- subset_M4_2_s5_1[which.max(subset_M4_2_s5_1$Counts), ]
    peaks_list_M4_2_s5_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M4_2_s5_1$Energy,
      Counts = max_row_M4_2_s5_1$Counts
    )
  }
}

peaks_M4_2_s5_1 <- do.call(rbind, peaks_list_M4_2_s5_1)
rownames(peaks_M4_2_s5_1) <- NULL
peaks_M4_2_s5_1 <- peaks_M4_2_s5_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta"))

M4_2_s5_1_graph <- filtered_M4_2_s5_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M4_2_s5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M4_2_s5_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M4_2_s5_4
filtered_M4_2_s5_4 <- subset(M4_2_s5_4, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M4_2_s5_4 <- list()
for (element in names(energy_ranges)) {
  range_M4_2_s5_4 <- energy_ranges[[element]]
  subset_M4_2_s5_4 <- subset(filtered_M4_2_s5_4, Energy >= range_M4_2_s5_4[1] & Energy <= range_M4_2_s5_4[2])
  if (nrow(subset_M4_2_s5_4) > 0) {
    max_row_M4_2_s5_4 <- subset_M4_2_s5_4[which.max(subset_M4_2_s5_4$Counts), ]
    peaks_list_M4_2_s5_4[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M4_2_s5_4$Energy,
      Counts = max_row_M4_2_s5_4$Counts
    )
  }
}

peaks_M4_2_s5_4 <- do.call(rbind, peaks_list_M4_2_s5_4)
rownames(peaks_M4_2_s5_4) <- NULL
peaks_M4_2_s5_4 <- peaks_M4_2_s5_4 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M4_2_s5_4_graph <- filtered_M4_2_s5_4 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M4_2_s5_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M4_2_s5_4, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M8_2
filtered_M8_2 <- subset(M8_2, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M8_2 <- list()
for (element in names(energy_ranges)) {
  range_M8_2 <- energy_ranges[[element]]
  subset_M8_2 <- subset(filtered_M8_2, Energy >= range_M8_2[1] & Energy <= range_M8_2[2])
  if (nrow(subset_M8_2) > 0) {
    max_row_M8_2 <- subset_M8_2[which.max(subset_M8_2$Counts), ]
    peaks_list_M8_2[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M8_2$Energy,
      Counts = max_row_M8_2$Counts
    )
  }
}

peaks_M8_2 <- do.call(rbind, peaks_list_M8_2)
rownames(peaks_M8_2) <- NULL
peaks_M8_2 <- peaks_M8_2 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M8_2_graph <- filtered_M8_2 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M8_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M8_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M10_4
filtered_M10_4 <- subset(M10_4, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M10_4 <- list()
for (element in names(energy_ranges)) {
  range_M10_4 <- energy_ranges[[element]]
  subset_M10_4 <- subset(filtered_M10_4, Energy >= range_M10_4[1] & Energy <= range_M10_4[2])
  if (nrow(subset_M10_4) > 0) {
    max_row_M10_4 <- subset_M10_4[which.max(subset_M10_4$Counts), ]
    peaks_list_M10_4[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M10_4$Energy,
      Counts = max_row_M10_4$Counts
    )
  }
}

peaks_M10_4 <- do.call(rbind, peaks_list_M10_4)
rownames(peaks_M10_4) <- NULL
peaks_M10_4 <- peaks_M10_4 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L","Mg_beta", "F", "Ca_L", "Fe_L", "Na", "Mg", "P", "S", "Cl", "Ca", "Fe_beta"))

M10_4_graph <- filtered_M10_4 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M10_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M10_4, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

# M11_1
filtered_M11_1 <- subset(M11_1, Energy >= 0.2 & Energy < 9)

# Extract the exact peak energy and counts to place labels
peaks_list_M11_1 <- list()
for (element in names(energy_ranges)) {
  range_M11_1 <- energy_ranges[[element]]
  subset_M11_1 <- subset(filtered_M11_1, Energy >= range_M11_1[1] & Energy <= range_M11_1[2])
  if (nrow(subset_M11_1) > 0) {
    max_row_M11_1 <- subset_M11_1[which.max(subset_M11_1$Counts), ]
    peaks_list_M11_1[[element]] <- data.frame(
      Element = element,
      Energy = max_row_M11_1$Energy,
      Counts = max_row_M11_1$Counts
    )
  }
}

peaks_M11_1 <- do.call(rbind, peaks_list_M11_1)
rownames(peaks_M11_1) <- NULL
peaks_M11_1 <- peaks_M11_1 |> 
  filter(Counts > 100) |> 
  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))

M11_1_graph <- filtered_M11_1 |> 
  select(Energy, Counts) |>
  ggplot(aes(x = Energy, y = Counts)) +
  geom_area(fill = "snow3", alpha = 0.7) +
  geom_line(linewidth = 0.5, alpha = 0.5, linetype = "solid", colour = "#8B8989") +
  labs(x = "Energy (keV)", y = "Intensity (Counts)") +
  theme_minimal() +
  geom_point(data = peaks_M11_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
  geom_label(data = peaks_M11_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
  theme(legend.position="none")

```

### Geological samples

#### Lower Jurassic of western Algarve

The analysed samples from the Lower Jurassic outcrops are homogeneous regarding their structure, inclusions and mineral compositions. The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group (Carixian, Lower Jurassic formations of Western Algarve, including the inland outcrop of Ferrel).

They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and small peaks of aluminium, with a single sample also showing a high aluminium peak. Small peaks of chlorine and potassium can also be identified in less than 50% of the samples.

Phosphatised skeletal remains are very frequent, occurring in more than 80% of the samples. These are all composed of calcium and phosphorus, and frequently show an associated smaller peak of sulphur (\~50% of the samples) or sulphur and fluorine (\~70% of the samples). These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the phosphatised skeletal remains. These skeletal remains are frequently irregular and indeterminate; however, a small number of samples show phosphatised skeletal remains with homogeneous oval shapes, which are currently indeterminate.

Finally, iron oxides are present in all the samples, with an iron (Fe) composition. Other types of iron oxides are also present in lesser frequencies, as equally large peaks of iron and titanium (40% of the samples), iron and fluoride (20%), and titanium with a small peak of iron (\<10%). A single sample shows a high titanium peak without iron and another titanium with a smaller peak of phosphorus. Most of these iron or titanium inclusions have irregular shapes and varying sizes. Albeit rare (present in 3 samples), pyrite can be identified in the form of spherical framboids or dispersed pyrite nanocrystals.

### 2025-04-17 analysis/paper/paper.qmd
Final SEM spectrum graphs and figures to include in main paper finished.

@@
 
-M18_7 <- read_csv('data/sem_data/M18_7.csv')
-M18_8 <- read_csv('data/sem_data/M18_8.csv')
-M4_3 <- read_csv('data/sem_data/M4_3.csv')
+M18_7 <- read_csv('analysis/data/raw_data/sem_data/M18_7.csv')
+M18_8 <- read_csv('analysis/data/raw_data/sem_data/M18_8.csv')
+M4_3 <- read_csv('analysis/data/raw_data/sem_data/M4_3.csv')
 
@@
 energy_ranges <- list(
-  C = c(0.2, 0.3),
-  O = c(0.5, 0.6),
-  Al = c(1.48, 1.5),
-  Si = c(1.7, 1.8),
-  P = c(2, 2.1),
-  Ca = c(3.68, 3.7),
-  Ca_beta = c(4.01, 4.02),
-  Ti = c(4.5, 4.6),
-  Ti_beta = c(4.9, 5),
-  Ti_L = c("0.45", "0.47"),
-  Fe = c(6.38, 6.41),
-  Fe_beta = c(7.05, 7.07),
-  Fe_L = c(0.71, 0.73))
+  C = c(0.2, 0.3), # "#584B9F" dark blue
+  O = c(0.5, 0.6), # "#0086B3" blue
+  F = c(0.67, 0.69), # "#3EADA9"  greenish blue
+  Na = c(1.04, 1.06), # "#4FC46A"  green"
+  Mg = c(1.25, 1.27), # "#AEC491" yellowish green
+  Mg_b = c(1.3, 1.4), # "#AEC491" yellowish green
+  Al = c(1.48, 1.5), # "#F2C86C" yellow
+  Si = c(1.7, 1.8), # "#F5A838" orange
+  P = c(2, 2.1), # "#EE850E" dark orange
+  S = c(2.3, 2.32), # "#D54F34" red
+  S_b = c(2.46, 2.47), # "#D54F34" red
+  Cl = c(2.62, 2.64), # "#A51122" wine
+  K = c(3.31, 3.33), # "#A71B4B" pinkish red
+  K_b = c(3.59, 3.61), # "#A71B4B" pinkish red
+  Ca = c(3.68, 3.7), # "#D35C79" pink
+  Ca_b = c(4.01, 4.02), # "#D35C79" pink
+  Ca_L = c(0.34, 0.36), # "#D35C79" pink
+  Ti = c(4.5, 4.6), # "#BD3977" light purple
+  Ti_b = c(4.9, 5), # "#BD3977" light purple
+  Ti_L = c(0.45, 0.47), # "#BD3977" light purple
+  Fe = c(6.38, 6.41), # "#762181" purple
+  Fe_b = c(7, 7.2), # "#762181" purple
+  Fe_L = c(0.69, 0.73)) # "#762181" purple
 
@@
   filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+peaks_M18_8_labels <- peaks_M18_8 |> 
+  filter(Element != "Fe_L")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 5000)) +
   geom_point(data = peaks_M18_8, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M18_8, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  #geom_label_repel(data = peaks_M18_8_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, fill = "white", label.size = 0.2, direction = "y", nudge_y = 0.1, max.overlaps = Inf, show.legend = FALSE) +
+  geom_text(data = peaks_M18_8_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M18_8_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 peaks_M18_7 <- peaks_M18_7 |> 
-  filter(Counts > 100)
+  filter(Counts > 100) |> 
+  filter(!Element %in% c("F", "Ti_L"))
+peaks_M18_7_labels <- peaks_M18_7 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 4000)) +
   geom_point(data = peaks_M18_7, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M18_7, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  #geom_label_repel(data = peaks_M18_8, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, fill = "white", label.size = 0.2, direction = "y", nudge_y = 0.1, max.overlaps = Inf, show.legend = FALSE) +
+  geom_text(data = peaks_M18_7_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M18_7_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 peaks_M4_3 <- peaks_M4_3 |> 
-  filter(Counts > 100)
+  filter(Counts > 100) |> 
+  filter(!Element %in% c("Mg_b", "K_b", "Ti_L"))
+peaks_M4_3_labels <- peaks_M4_3 |> 
+  filter(Element != "Ca_b")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 5000)) +
   geom_point(data = peaks_M4_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M4_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  #geom_text_repel(data = peaks_M4_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, direction = "y", nudge_y = 0.5, max.overlaps = Inf, force = 1, show.legend = FALSE) +
+  geom_text(data = peaks_M4_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M4_3_graph.png", width = 10 ,bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-#knitr::include_graphics(here::here("figures/within-chert/M18_7_M18_8.png"))
-M18_7_M18_8_fig <- magick::image_read(here::here("figures/within-chert/M18_7_M18_8.png"))
-M18_7_M18_8_gg <- cowplot::ggdraw() + cowplot::draw_image(M18_7_M18_8_fig)
-
-M18_7_M18_8_plots <- plot_grid(M18_7_graph, M18_8_graph, labels = c(1, 2))
-
-plot_grid(M18_7_M18_8_gg, M18_7_M18_8_plots, ncol = 1, rel_heights = c(1, 1))
+knitr::include_graphics(here::here("figures/within-chert/M18_7_M18_8_graphs.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/M4_3.png"))
+knitr::include_graphics(here::here("figures/within-chert/M4_3_graphs.png"))
 

### 2025-04-17 analysis/supplementary-materials/appendix1.qmd
SOM SEM color codes added

@@
 energy_ranges <- list(
-  C = c(0.2, 0.3),
-  O = c(0.5, 0.6),
-  F = c(0.67, 0.69),
-  Na = c(1.04, 1.06),
-  Mg = c(1.25, 1.27),
-  Mg_beta = c(1.3, 1.4),
-  Al = c(1.48, 1.5),
-  Si = c(1.7, 1.8),
-  P = c(2, 2.1),
-  S = c(2.3, 2.32),
-  S_beta = c(2.46, 2.47),
-  Cl = c(2.62, 2.64),
-  K = c(3.31, 3.33),
-  K_beta = c(3.59, 3.61),
-  Ca = c(3.68, 3.7),
-  Ca_beta = c(4.01, 4.02),
-  Ca_L = c(0.34, 0.36),
-  Ti = c(4.5, 4.6),
-  Ti_beta = c(4.9, 5),
-  Ti_L = c(0.45, 0.47),
-  Fe = c(6.38, 6.41),
-  Fe_beta = c(7, 7.2),
-  Fe_L = c(0.69, 0.73))
+  C = c(0.2, 0.3), # "#584B9F" dark blue
+  O = c(0.5, 0.6), # "#0086B3" blue
+  F = c(0.67, 0.69), # "#3EADA9"  greenish blue
+  Na = c(1.04, 1.06), # "#4FC46A"  green"
+  Mg = c(1.25, 1.27), # "#AEC491" yellowish green
+  Mg_b = c(1.3, 1.4), # "#AEC491" yellowish green
+  Al = c(1.48, 1.5), # "#F2C86C" yellow
+  Si = c(1.7, 1.8), # "#F5A838" orange
+  P = c(2, 2.1), # "#EE850E" dark orange
+  S = c(2.3, 2.32), # "#D54F34" red
+  S_b = c(2.46, 2.47), # "#D54F34" red
+  Cl = c(2.62, 2.64), # "#A51122" wine
+  K = c(3.31, 3.33), # "#A71B4B" pinkish red
+  K_b = c(3.59, 3.61), # "#A71B4B" pinkish red
+  Ca = c(3.68, 3.7), # "#D35C79" pink
+  Ca_b = c(4.01, 4.02), # "#D35C79" pink
+  Ca_L = c(0.34, 0.36), # "#D35C79" pink
+  Ti = c(4.5, 4.6), # "#BD3977" light purple
+  Ti_b = c(4.9, 5), # "#BD3977" light purple
+  Ti_L = c(0.45, 0.47), # "#BD3977" light purple
+  Fe = c(6.38, 6.41), # "#762181" purple
+  Fe_b = c(7, 7.2), # "#762181" purple
+  Fe_L = c(0.69, 0.73)) # "#762181" purple
 

### 2025-04-17 analysis/supplementary-materials/appendix1.qmd
Macbook session. SOM SEM graphs ..01

@@
   filter(Counts > 250) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S_beta", "S", "Cl"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S_b", "S", "Cl"))
+peaks_M24_1_2_labels <- peaks_M24_1_2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 9500)) +
   geom_point(data = peaks_M24_1_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M24_1_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M24_1_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
 
@@
 peaks_M24_1_1 <- peaks_M24_1_1 |> 
-  filter(Counts > 300) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "P", "S", "Cl"))
+  filter(Counts > 250) |> 
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "P", "S", "Cl", "S_b", "K", "K_b"))
+peaks_M24_1_1_labels <- peaks_M24_1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 8500)) +
   geom_point(data = peaks_M24_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M24_1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M24_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M23_7_1_labels <- peaks_M23_7_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 2500)) +
   geom_point(data = peaks_M23_7_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M23_7_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M23_7_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
 
@@
   filter(Counts > 70) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "K_beta"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
+peaks_M23_2_1_labels <- peaks_M23_2_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 2200)) +
   geom_point(data = peaks_M23_2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M23_2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M23_2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
 
@@
   filter(Counts > 55) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M22_5.1_3_labels <- peaks_M22_5.1_3 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 1800)) +
   geom_point(data = peaks_M22_5.1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M22_5.1_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M22_5.1_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
 
@@
   filter(Counts > 70) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S"))
+peaks_M22_5.1_1_labels <- peaks_M22_5.1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 3800)) +
   geom_point(data = peaks_M22_5.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M22_5.1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M22_5.1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
 

### 2025-04-17 analysis/supplementary-materials/appendix1.qmd
SOM SEM graphs finished in R ..02 Starting to add them to final SOM figs in illustrator.

@@
 
+# ggsave("M24_1_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+
 # M24_1_1
@@
   filter(Counts > 250) |> 
-  filter(!Element %in% c("Ti_L", "Mg_b", "F", "P", "S", "Cl", "S_b", "K", "K_b"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "P", "S", "Cl", "S_b", "K", "K_b", "Mg"))
 peaks_M24_1_1_labels <- peaks_M24_1_1 |> 
@@
   theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M24_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
+# ggsave("M23_7_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+
 # M23_2_1
@@
 
+# ggsave("M23_2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+
 # M22_5.1_3
@@
   filter(Counts > 55) |> 
-  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "Cl"))
 peaks_M22_5.1_3_labels <- peaks_M22_5.1_3 |> 
@@
   theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M22_5.1_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
+# ggsave("M22_5.1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+
 # M22_4.1_1
@@
   filter(Counts > 50) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M22_4.1_1_labels <- peaks_M22_4.1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M22_4.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M22_4.1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M22_4.1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M22_4.1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 400) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S", "Mg", "S_beta", "Ca_beta", "Ti_beta", "K_beta", "K"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S", "Mg", "S_b", "Ca_b", "Ti_b", "K_b", "K"))
+peaks_M20_2_1_labels <- peaks_M20_2_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M20_2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M20_2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M20_2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M20_2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 150) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M18_5_1_labels <- peaks_M18_5_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M18_5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M18_5_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M18_5_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M18_5_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 200) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M18_1_1_labels <- peaks_M18_1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M18_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M18_1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M18_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M18_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 600) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "S_beta"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "S_b"))
+peaks_M16_1_1_labels <- peaks_M16_1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M16_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M16_1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M16_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M16_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 300) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "Na", "P", "S", "Cl", "K", "Ca_beta", "S_beta", "Ca", "K_beta", "Ca_L"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "Na", "P", "S", "Cl", "K", "Ca_b", "S_b", "Ca", "K_b", "Ca_L", "Mg"))
+peaks_M15_2_s2_2_labels <- peaks_M15_2_s2_2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M15_2_s2_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M15_2_s2_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M15_2_s2_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M15_2_s2_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 500) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F", "K_beta"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b", "Na", "Mg"))
+peaks_M15_2_s2_1_labels <- peaks_M15_2_s2_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M15_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M15_2_s2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M15_2_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M15_2_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 700) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "P"))
+peaks_M14_4_1_labels <- peaks_M14_4_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M14_4_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M14_4_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M14_4_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M14_4_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 60) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M14_3_2_s2_1_labels <- peaks_M14_3_2_s2_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M14_3_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M14_3_2_s2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M14_3_2_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M14_3_2_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 500) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M14_3_2_s1_1_labels <- peaks_M14_3_2_s1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M14_3_2_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M14_3_2_s1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M14_3_2_s1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M14_3_2_s1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 300) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M13_1_3_s2_labels <- peaks_M13_1_3_s2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M13_1_3_s2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M13_1_3_s2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M13_1_3_s2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M13_1_3_s2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_5_s6_2_labels <- peaks_M12_5_s6_2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_5_s6_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_5_s6_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_5_s6_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_5_s6_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_5_s6_1_labels <- peaks_M12_5_s6_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_5_s6_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_5_s6_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_5_s6_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_5_s6_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_5_s3_3_labels <- peaks_M12_5_s3_3 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_5_s3_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_5_s3_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_5_s3_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_5_s3_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_5_s1_3_labels <- peaks_M12_5_s1_3 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_5_s1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_5_s1_3, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_5_s1_3_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+#ggsave("M12_5_s1_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L","Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L","Mg_b", "F"))
+peaks_M12_5_s1_1_labels <- peaks_M12_5_s1_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_5_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_5_s1_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_5_s1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_5_s1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_3_s3_2_labels <- peaks_M12_3_s3_2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_3_s3_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_3_s3_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_3_s3_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_3_s3_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 90) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_1_s3_1_labels <- peaks_M12_1_s3_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_1_s3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_1_s3_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_1_s3_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_1_s3_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M12_1_s2_1_labels <- peaks_M12_1_s2_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M12_1_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M12_1_s2_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M12_1_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M12_1_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 peaks_M2_3_map2 <- peaks_M2_3_map2 |> 
-  filter(Counts > 100)
+  filter(Counts > 100) |> 
+  filter(Element != "Ca_L")
+peaks_M2_3_map2_labels <- peaks_M2_3_map2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M2_3_map2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M2_3_map2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M2_3_map2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M2_3_map2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(Element != "Ti_L")
+  filter(!Element %in% c("Ti_L", "K_b"))
+peaks_M3_1_labels <- peaks_M3_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M3_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M3_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+ggsave("M3_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "K_b", "Fe_L"))
+peaks_M3_3_s1_labels <- peaks_M3_3_s1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 3500), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M3_3_s1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M3_3_s1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M3_3_s1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M3_3_s1_graph.png", width = 10, height = 5, bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 90) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
+peaks_M3_3_s2_labels <- peaks_M3_3_s2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M3_3_s2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M3_3_s2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M3_3_s2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M3_3_s2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
+peaks_M4_2_s5_1_labels <- peaks_M4_2_s5_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(xlim = c(0.3, 5)) +
   geom_point(data = peaks_M4_2_s5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M4_2_s5_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M4_2_s5_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+#ggsave("M4_2_s5_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M4_2_s5_4_labels <- peaks_M4_2_s5_4 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M4_2_s5_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M4_2_s5_4, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M4_2_s5_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M4_2_s5_4_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F", "K_b"))
+peaks_M8_2_labels <- peaks_M8_2 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 3500), xlim = c(0, 5)) +
   geom_point(data = peaks_M8_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M8_2, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M8_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+ ggsave("M8_2_graph.png",width = 7, bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L","Mg_beta", "F", "Ca_L", "Fe_L", "Na", "Mg", "P", "S", "Cl", "Ca", "Fe_beta"))
+  filter(!Element %in% c("Ti_L","Mg_b", "F", "Ca_L", "Fe_L", "Na", "Mg", "P", "S", "Cl", "Ca", "Fe_b", "K", "K_b", "S_b"))
+peaks_M10_4_labels <- peaks_M10_4 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M10_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M10_4, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M10_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M10_4_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   filter(Counts > 100) |> 
-  filter(!Element %in% c("Ti_L", "Mg_beta", "F"))
+  filter(!Element %in% c("Ti_L", "Mg_b", "F"))
+peaks_M11_1_labels <- peaks_M11_1 |> 
+  filter(!Element %in% c("Fe_L", "Fe_b", "Mg_b", "S_b", "K_b", "Ca_b", "Ca_L", "Ti_b", "Ti_L"))
 
@@
   geom_point(data = peaks_M11_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_label(data = peaks_M11_1, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 3, fill = "white", label.size = 0.2, vjust = -0.5, show.legend = FALSE) +
-  theme(legend.position="none")
+  geom_text(data = peaks_M11_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+
+# ggsave("M11_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 

### 2025-04-19 analysis/paper/paper.qmd
SOM SEM graphs finished, final figs finished.

@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
 
-# ggsave("M18_8_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M18_8_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
 
-# ggsave("M18_7_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M18_7_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
-  coord_cartesian(ylim = c(0, 5000)) +
+  coord_cartesian(ylim = c(0, 5000), xlim = c(0.3, 4.5)) +
   geom_point(data = peaks_M4_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
 
-# ggsave("M4_3_graph.png", width = 10 ,bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+ggsave("M4_3_graph.png", width = 5, height = 3.75, units = "in", ,bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 

### 2025-04-19 analysis/supplementary-materials/appendix1.qmd
SOM SEM graphs finished, final figs finished.

@@
 
+# Resulting graphics were exported and mounted with the SEM images
+
 # Prepare the energy ranges
@@
   theme_minimal() +
-  coord_cartesian(ylim = c(0, 9500)) +
+  coord_cartesian(ylim = c(0, 9600), xlim = c(0.3, 7.6)) +
   geom_point(data = peaks_M24_1_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M24_1_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M24_1_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M24_1_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+#ggsave("M24_1_2_graph.png", bg = "white", width = 5, height = 3.75, units = "in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
-  coord_cartesian(ylim = c(0, 8500)) +
+  coord_cartesian(ylim = c(0, 8600), xlim = c(0, 7.8)) +
   geom_point(data = peaks_M24_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M24_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M24_1_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M23_7_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M23_7_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
-  coord_cartesian(ylim = c(0, 3800)) +
+  coord_cartesian(ylim = c(0, 3800), xlim = c(0.5, 7.2)) +
   geom_point(data = peaks_M22_5.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M22_5.1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M22_5.1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M22_5.1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M22_5.1_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0,900), xlim = c(0, 8)) +
   geom_point(data = peaks_M22_4.1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M22_4.1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M22_4.1_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 4300), xlim = c(0, 7.7)) +
   geom_point(data = peaks_M18_5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M18_5_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M18_5_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 6000), xlim = c(0, 7.7)) +  
   geom_point(data = peaks_M18_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M18_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M18_1_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 26000), xlim = c(0.5, 7.5)) +
   geom_point(data = peaks_M16_1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M16_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M16_1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M16_1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M16_1_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 28000), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M15_2_s2_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M15_2_s2_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M15_2_s2_2_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 50000), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M15_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M15_2_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M15_2_s2_1_graph.png", bg = "white",width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 52000), xlim = c(0, 7.7)) +
   geom_point(data = peaks_M14_4_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M14_4_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M14_4_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 2500), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M14_3_2_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M14_3_2_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M14_3_2_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M14_3_2_s2_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M14_3_2_s2_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 40000), xlim = c(0.5, 7.5)) +
   geom_point(data = peaks_M14_3_2_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M14_3_2_s1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M14_3_2_s1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M14_3_2_s1_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M14_3_2_s1_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 4100), xlim = c(0, 8)) +
   geom_point(data = peaks_M12_5_s6_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M12_5_s6_2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M12_5_s6_2_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 3800), xlim = c(0, 8)) +
   geom_point(data = peaks_M12_5_s6_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M12_5_s6_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+#ggsave("M12_5_s6_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 4300), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M12_5_s3_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M12_5_s3_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M12_5_s3_3_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 6000), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M12_5_s1_3, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-#ggsave("M12_5_s1_3_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M12_5_s1_3_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 5000), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M12_5_s1_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M12_5_s1_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M12_5_s1_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 6200), xlim = c(0, 7)) +
   geom_point(data = peaks_M12_1_s3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M12_1_s3_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M12_1_s3_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M12_1_s3_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M12_1_s3_1_graph.png", bg = "white" , width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 6000), xlim = c(0.5, 7.6)) +
   geom_point(data = peaks_M12_1_s2_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M12_1_s2_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M12_1_s2_1_graph.png", width = 5, height = 3.75, unit ="in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 4200), xlim = c(0, 8)) +
   geom_point(data = peaks_M2_3_map2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M2_3_map2_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M2_3_map2_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 3400), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M3_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
@@
   C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-ggsave("M3_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M3_1_graph.png",width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
-  coord_cartesian(xlim = c(0.3, 5)) +
+  coord_cartesian(xlim = c(0.3, 5), ylim = c(0, 3500)) +
   geom_point(data = peaks_M4_2_s5_1, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M4_2_s5_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-#ggsave("M4_2_s5_1_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M4_2_s5_1_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+# ggsave("M4_2_s5_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 5700), xlim = c(0, 7.5)) +
   geom_point(data = peaks_M4_2_s5_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M4_2_s5_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M4_2_s5_4_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M4_2_s5_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+## ggsave("M4_2_s5_4_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
-  coord_cartesian(ylim = c(0, 3500), xlim = c(0, 5)) +
+  coord_cartesian(ylim = c(0, 3500), xlim = c(0, 4.5)) +
   geom_point(data = peaks_M8_2, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M8_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
- ggsave("M8_2_graph.png",width = 7, bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M8_2_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M8_2_graph.png",width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
   theme_minimal() +
+  coord_cartesian(ylim = c(0, 12500), xlim = c(0, 7.7)) +
   geom_point(data = peaks_M10_4, aes(x = Energy, y = Counts, colour = Element), size = 3, shape = 18) +
-  geom_text(data = peaks_M10_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 5, vjust = -0.5, show.legend = FALSE) +
-  scale_colour_manual(values = c(
-  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
-  theme(legend.position="none", axis.text = element_text(size = 15), axis.title = element_text(size = 18))
-
-# ggsave("M10_4_graph.png", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+  geom_text(data = peaks_M10_4_labels, aes(x = Energy, y = Counts, label = Element, colour = Element), size = 4, vjust = -0.5, show.legend = FALSE) +
+  scale_colour_manual(values = c(
+  C = "#4461A8", O = "#2BACC2", F = "#3EADA9", Na = "#4FC46A", Mg = "#AEC491", Mg_b = "#AEC491", Al = "#F2C86C", Si = "#F5A838", P = "#EE850E", S = "#D54F34", S_b = "#D54F34", Cl = "#A51122", K = "#A71B4B", K_b = "#A71B4B", Ca = "#D35C79" , Ca_b = "#D35C79", Ca_L = "#D35C79", Ti = "#992766", Ti_b = "#992766", Ti_L = "#992766", Fe = "#762181" , Fe_b = "#762181" , Fe_L = "#762181")) +
+  theme(legend.position="none", axis.text = element_text(size = 12), axis.title = element_text(size = 12))
+
+ggsave("M10_4_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 

### 2025-04-19 analysis/paper/paper.qmd
Small improvements to caps on main paper

@@
 #| label: fig-skeletal
-#| fig-cap: "Skeletal remain from sample SP16_BLX, from the Lower Jurassic ourcrops of western Algarve.
-A: Detail of a phosphatized skeletal remain composed by calcium (A1), phosphorus (A2) and a small peak of sulphur (A3).
-The skeletal remain seems to be oxidized, with small peaks of iron within (A4)."
+#| fig-cap: "A) Skeletal remains from sample SP16_BLX, from the Lower Jurassic outcrops of western Algarve.
+B) Detail of skeletal remains composed of calcium (B1), phosphorus (B2) and a small peak of sulphur (B3).
+The skeletal remains seem to be oxidized, with small peaks of iron within (B4)."
 
@@
 #| label: fig-oval-fossils
-#| fig-cap: "A-C: General view and details from an archaeological sample from local chert type (T4).
-A: general view of a concentration of oval fossils.
-B: general view of a concentration of oval fossils and other inclusions.
-C: detail of an irregular inclusion composed of titanium (C1) and an oval phosphatised skeletal remain composed mainly of calcium (C2), phosphorus (C3) and a small peak of sulphur."
+#| fig-cap: "A-C) General view and details from an archaeological sample from local chert type (T4).
+A) general view of a concentration of oval fossils.
+B) general view of a concentration of oval fossils and other inclusions.
+C) detail of an irregular inclusion composed of titanium (C1) and an oval skeletal remains composed mainly of calcium (C2), phosphorus (C3) and a small peak of sulphur."
 
@@
 #| label: fig-t6-general
-#| fig-cap: "A-D: General view and details from an archaeological sample from non-local chert type (T6).
-A: general view of the sample with uncommon inclusions.
-B: detail of an inclusion seen in A,  composed mostly of iron and small peaks of calcium and phosphorus, making this inclusion a possible oxidised skeletal remains of a fossil.
-C: general view of the sample with no visible inclusions.
-D: detail of a small and irregular iron inclusion."
+#| fig-cap: "A-D) General view and details from an archaeological sample from non-local chert type (T6).
+Dots represent the measured spectrum and shown in the corresponding graph.
+A) general view of the sample with uncommon inclusions.
+B) detail of an inclusion seen in A, composed mostly of iron and small peaks of calcium and phosphorus, making this inclusion a possible oxidised skeletal remains of a fossil (1).
+C) general view of the sample with no visible inclusions.
+D) detail of a small and irregular iron inclusion (2)."
 
@@
 #| label: fig-framboids
-#| fig-cap: "A-B: Detail from the geological sample SP59_JOR from the Upper Jurassic outcrops (Jordana formation) of eastern Algarve.
-A: detailed view of a spherical pyrite framboid composed of iron (A1) and sulphur (A2).
-B: detailed view of an inclusion composed of iron (B1) and sulphur (B2)."
+#| fig-cap: "A-B) Detail from the geological sample SP59_JOR from the Upper Jurassic outcrops (Jordana formation) of eastern Algarve.
+A) detailed view of a spherical pyrite framboid composed of iron (A1) and sulphur (A2).
+B) detailed view of an inclusion composed of iron (B1) and sulphur (B2)."
 
@@
 #| label: fig-calcite
-#| fig-cap: "A-B: General view of calcite inclusions (black arrows) from the geological sample SP69_MAR from the Upper Jurassic outcrops of western Algarve."
+#| fig-cap: "General view of calcite inclusions (light grey) from the geological sample SP69_MAR from the Upper Jurassic outcrops of western Algarve, composed mainly of calcium and silica (1).
+Dot represents the measured spectrum and is shown in the corresponding graph."
 

### 2025-04-20 analysis/supplementary-materials/appendix1.qmd
Desktop work session. Added Lower Jurassic west text, figs and caps to SOM.

@@
 csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
-
 ---
@@
 The analysed samples from the Lower Jurassic outcrops are homogeneous regarding their structure, inclusions and mineral compositions.
-The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group (Carixian, Lower Jurassic formations of Western Algarve, including the inland outcrop of Ferrel).
+The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group (Carixian, Lower Jurassic formations of western Algarve, including the inland outcrop of Ferrel).
 
@@
 
-Phosphatised skeletal remains are very frequent, occurring in more than 80% of the samples.
-These are all composed of calcium and phosphorus, and frequently show an associated smaller peak of sulphur (\~50% of the samples) or sulphur and fluorine (\~70% of the samples).
-These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the phosphatised skeletal remains.
-These skeletal remains are frequently irregular and indeterminate; however, a small number of samples show phosphatised skeletal remains with homogeneous oval shapes, which are currently indeterminate.
-
-Finally, iron oxides are present in all the samples, with an iron (Fe) composition.
-Other types of iron oxides are also present in lesser frequencies, as equally large peaks of iron and titanium (40% of the samples), iron and fluoride (20%), and titanium with a small peak of iron (\<10%).
-A single sample shows a high titanium peak without iron and another titanium with a smaller peak of phosphorus.
+Fossils are very frequent, occurring in more than 80% of the samples.
+These are all composed of calcium and phosphorus, and frequently show an associated smaller peak of sulphur (\~70% of the samples; @fig-m4-2 and @fig-m3-1).
+These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the fossils (@fig-m4-4, @fig-m2-3-mp1 and @fig-m23-3).
+These skeletal remains are frequently irregular and indeterminate; however, a small number of samples show fossils with homogeneous oval shapes, which are currently indeterminate (@fig-m8-3, @fig-m23-2 and @fig-m4-2-m23-2).
+
+Finally, iron oxides are present in all the samples, with an iron (Fe) composition (@fig-m2-1-m4-1, @fig-m23-7 and @fig-m22-4).
+Other types of iron oxides are also present in lesser frequencies, as equally large peaks of iron and titanium (40% of the samples; @fig-m4-4) and titanium with a small peak of iron (\<10%).
+A single sample shows a high titanium peak without iron (@fig-m2-3-mp2) and another titanium with a smaller peak of phosphorus.
 Most of these iron or titanium inclusions have irregular shapes and varying sizes.
-Albeit rare (present in 3 samples), pyrite can be identified in the form of spherical framboids or dispersed pyrite nanocrystals.
+Albeit rare (present in 3 samples), pyrite can possibly be identified in the form of spherical framboids or dispersed pyrite nanocrystals (@fig-m22-5-m23-2).
+
+```{r}
+#| label: fig-m4-2
+#| fig-cap: "A) general view of the sample (x250 magnification) showing several inclusions and porosity.
+B) detail of the inclusions (x900 magnification), showing an irregular  fossil composed mainly of calcium, phosphorus, and with small peaks of sulphur (1).
+Some iron oxides are found in close proximity to the fossil (2).
+Dots represent the location of the measured spectra plotted in graph 1 and 2.
+Geological sample SP34_PdA_b from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M4_2_spectra5_graphs.png"))
+
+```
+
+```{r}
+#| label: fig-m3-1
+#| fig-cap: "A) detail of a fossil (x1k magnification) composed mainly of calcium, phosphorus, and with small peaks of sulphur (1).
+Dot represents the location of the measured spectrum plotted in graph 1.
+Geological sample SP40_FER from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M3_1_map1_map2_graphs.png"))
+
+```
+
+```{r}
+#| label: fig-m4-4
+#| fig-cap: "A) detail of inclusions (x800 magnification) composed of calcium (A1), phosphorus (A2) and sulphur (A3), possibly corresponding to fossils, with a small peak of iron oxides (A4), and an inclusion composed of iron (A4) and titanium (A5).
+Geological sample SP8_PdA from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M4_4_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m2-3-mp1
+#| fig-cap: "A) general view of the sample (x250 magnification) showing several inclusions and porosity.
+B) detail of a fossil (x2.1k magnification) composed of calcium (B1) and phosphorus (B2) with trace presence of sulphur (B3).
+The fossils may be oxidized, due to the presence of iron (B4).
+Geological sample SP14_BLX from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M2_3_map1_gen.png"))
+
+```
+
+```{r}
+#| label: fig-m23-3
+#| fig-cap: "A) detail of an irregular fossil (x1.6k magnification) composed of calcium (A1), phosphorus (A2), a small peak of sulphur, and iron speckles (A3) which correspond to the lighter dots in figure A.
+Geological sample SP32_FZF from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M23_3_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m8-3
+#| fig-cap: "A) general view of the sample (x250 magnification) showing a large oval fossil.
+B) detail of the fossil (x1k magnification) composed by calcium (B1), phosphorus (B2) and a small peak of sulphur (B3).
+The fossil may be oxidized, with small peaks of iron within (B4).
+Geological sample SP21_BLS from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M8_3_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m23-2
+#| fig-cap: "A) detail of an oval fossil (x850 magnification) composed of calcium (A1), phosphorus (A2) and iron speckles (A3) which correspond to the lighter dots in figure A.
+Geological sample SP30_FZF from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M23_2_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m4-2-m23-2
+#| fig-cap: "A) general view of the sample (x250 magnification) showing a large fossil.
+B) detailed view of the fossil (x650 magnification) mainly composed of calcium, phosphorus and small peaks of sulphur.
+Geological sample SP34_PdA_b from the LJW formation.
+C) general of the sample (x250 magnification) showing two large fossils.
+D) detail of the fossil (x1k magnification) composed of calcium, phosphorus and a small peak of sulphur.
+Geological sample SP30_FZF from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M4_2_M23_2.png"))
+
+```
+
+```{r}
+#| label: fig-m2-1-m4-1
+#| fig-cap: "A) detail of inclusions (x420 magnification) mainly composed of iron (A1).
+Geological sample SP15_BLX from the LJW formation.
+B) detail of a small inclusion (x1k magnification) mainly composed of iron (B1), with small peaks of calcium and phosphorus.
+Geological sample SP10_PdA from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M2_1_M4_1.png"))
+
+```
+
+```{r}
+#| label: fig-m23-7
+#| fig-cap: "Detail of small, round inclusions (x1.5 magnification) composed of iron within a quartz matrix composed of silicon, oxygen and a small peak of aluminium (1).
+Dot represents the location of the measured spectrum plotted in graph 1.
+Geological sample SP29_ASP from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M23_7_spectra2_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m22-4
+#| fig-cap: "Detail of an irregular inclusion (x1k magnification) composed mainly of iron (1).
+Dot represents the location of the measured spectrum plotted in graph 1.
+Geological sample SP41_FER from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M22_4.1_spectra1_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m2-3-mp2
+#| fig-cap: "Detail of an inclusion (x3.7k magnification) mainly composed of titanium.
+Dot represents the location of the measured spectrum plotted in graph 1.
+Geological sample SP14_BLX from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M2_3_map2_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m22-5-m23-2
+#| fig-cap: "A) detail of inclusions (x2.3k magnification) composed mainly of iron (A1).
+Possibly originally pyrite nanocrystal aggregates and spherical framboids.
+Geological sample SP40_FER from the LJW formation.
+B) detail of inclusions (x1.4k magnification) composed of iron, silicon and a small peak of aluminium (1).
+The cluster of rounded iron inclusions is possibly originally dispersed pyrite nanocrystals.
+Geological sample SP30_FZF from the LJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M22_5.1_M23_2_graph.png"))
+
+```

### 2025-04-21 analysis/paper/paper.qmd
XRD refinement dataset mutated to present percentages.

@@
 # Read grouped supporting data
-xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
-xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels
+xrd_perc <- read_csv('analysis/data/raw_data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
+xrd_marks <- read_csv('analysis/data/raw_data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels
 
 # Read individual xrd files
-sp15_xrd <- read_csv('data/xrd_data/SP15_xrd.csv')
-sp58_xrd <- read_csv('data/xrd_data/SP58_xrd.csv')
-sp62_xrd <- read_csv('data/xrd_data/SP62_xrd.csv')
-sp69_xrd <- read_csv('data/xrd_data/SP69_xrd.csv')
+sp15_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP15_xrd.csv')
+sp58_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP58_xrd.csv')
+sp62_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP62_xrd.csv')
+sp69_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP69_xrd.csv')
 
@@
 #| label: xrd-graphs
+
+# Turn value from percent table to percentages
+xrd_perc <- xrd_perc %>%
+  mutate(`Content (%)` = round(`Content (%)` * 100, 4))
 

### 2025-04-21 analysis/supplementary-materials/appendix1.qmd
XRD refinement dataset mutated to present percentages.

@@
 # Read grouped supporting data
-xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
+xrd_perc <- read_csv('analysis/data/raw_data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
 xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels
+
+# Turn value from percent table to percentages
+xrd_perc <- xrd_perc %>%
+  mutate(`Content (%)` = round(`Content (%)` * 100, 4))
 
@@
 sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
-sp18_xrd <- read_csv('data/xrd_data/SP18_xrd.csv')
+sp18_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP18_xrd.csv')
 sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
@@
 #| fig-cap: Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP18_PBX).
-The rhombus shape signifies main peaks key to identify each mineral.
+The rhombus shape signifies the main peaks key to identify each mineral.
 Labels represent the d spacing values of key mineral peaks.

### 2025-04-21 analysis/paper/paper.qmd
Main text added to paper

@@
 <!-- The actual document text starts here: -->
+
+<!-- Library, datasets and analysis code chunks section -->
 
@@
 
-Here is a citation [@Marwick2017]
+Understanding the movement of raw materials in prehistory is essential for reconstructing the available technology, human mobility, and trade networks of past hunter-gatherers.
+Among these materials, chert played a crucial role in lithic tool production, making its characterisation and sourcing a key issue in archaeological research, particularly in Upper Palaeolithic (UP) assemblages.
+Determining the origins of chert artefacts  reveals patterns of raw material availability and provides valuable insights into prehistoric procurement strategies, territorial behaviours, and long-distance exchange networks over time.
 
-# Background
+Chert sourcing studies have traditionally relied on macroscopic and petrographic techniques, providing valuable insights into raw material formations and procurement by hunter-gatherer groups.
+While these methods remain essential in lithic sourcing, they also have limitations.
+Their reliance on physical appearance and petrographic homogeneity can make it difficult to distinguish geologically similar cherts, especially in cases of long-distance procurement.
+Moreover, although macroscopic analysis is non-destructive, thin-section petrography requires destructive sampling, raising concerns about the preservation of archaeological materials.
 
-# Methods
+Recent research has increasingly incorporated geochemical and spectroscopic techniques to address these challenges and refine chert sourcing methodologies.
+Luedtke (1978, 1979, 1992) highlights the value of chemical data combined with multivariate statistical analysis, while Brandl (2016) integrates petrographic and geochemical approaches within the Multi-Layered Chert Sourcing (MLA) framework.
+
+In the Iberian Peninsula - a key region for studying human adaptations and cultural transitions - these approaches have also been applied to UP archaeological assemblages.
+These multi-method strategies have successfully characterized cherts from diverse geological formations and identified mobility patterns throughout the Upper Paleolithic, including evidence of long-distance contacts between hunter-gatherer groups.
+
+In recent years, a wide range of analytical techniques - such as X-Ray Diffraction (XRD), X-Ray Fluorescence (XRF), portable XRF, Laser Ablation Inductively Coupled Plasma Mass Spectrometry (LA-ICP-MS), Raman Spectroscopy, and Scanning Electron Microscopy with Energy Dispersive Spectroscopy (SEM-EDS) - has been employed to enhance chert characterisation in this region.
+Each method offers distinct advantages in data resolution, sample preparation, and cost-effectiveness.
+However, no single technique is universally applicable, and a multi-analytical approach is often necessary to effectively address specific research questions.
+
+Southern Portugal, while having benefited from previous geochemical studies of one of the four chert-bearing formations, as well as macroscopic and petrographic studies of chert-bearing formations (Belmiro et al., 2023) and Upper Palaeolithic chert assemblages from the site of Vale Boi (Belmiro et al., 2025), remains largely unexplored through other analytical techniques.
+These may be key to address open research topics and questions by recent studies.
+
+For example, local and regional Jurassic chert sources are abundant in the region and in relative proximity to the site, yet the differentiation between specific chert formations remains unclear, which is critical for understanding regional mobility patterns (Belmiro et al., 2023).
+
+Additionally, evidence suggests that non-local materials were frequently introduced into the archaeological site, indicating complex long-distance procurement dynamics.
+However, these findings were based on macroscopic and petrographic studies, which leaves room for further characterization to strengthen source attributions and test current hypotheses.
+
+Following the results and questions from previous studies by our team (Belmiro et al., 2023; Belmiro et al., 2025), this study employs an integrated analytical approach to geological chert samples from local, regional, and non-local formations, as well as archaeological assemblages from the UP levels of Vale Boi to accomplish three main goals:
+
+- Complete the characterization of chert samples from the Algarve region, including data on their mineral and chemical composition.
+This will deepen our understanding of the characteristics of outcrops across the region and make this information accessible to the scientific community through an online database.
+Additionally, we aim to assess the potential of mineral and chemical analyses to distinguish between macroscopic and petrographically similar chert samples from the Lower Jurassic Carixian (west Algarve basin) and Middle Jurassic Malhão (east Algarve basin) formations.
+This will enhance our understanding of which regional sources were exploited by hunter-gatherer groups at Vale Boi—whether the majority of “pink/yellow” cherts were sourced from local outcrops (within 20 km of Vale Boi) or if some portion was procured from more distant sources.
+
+- Use geochemical methods to further characterize the archaeological samples, as well as geological samples from other regions of interest (e.g., Cretaceous cherts from central Portugal and cherts from the Betic Systems of southern Spain).
+This will enable us to compare the geological and archaeological data, testing, corroborating, and strengthening the petrographic and macroscopic source attributions of archaeological chert types.
+Additionally, comparing local geological and archaeological data may provide valuable insights into the selection of nodules for knapping, especially considering the known process of dolomitization in the Lower Jurassic outcrops.
+
+- Test the potential of various techniques and approaches to enhance the interpretation of archaeological chert raw materials and human mobility patterns.
+Specifically, we will assess whether these methods effectively characterize and differentiate between chert types from various formations, both regional and long-distance, and enable accurate source attribution of archaeological lithic artefacts.
+
+To accomplish these goals, we apply three different methods: X-Ray Diffraction (XRD) to regional Algarve chert samples; Scanning Electron Microscopy coupled with X-Ray Microscopy (SEM-EDS) to regional Algarve chert samples and all identified chert types from the Upper Palaeolithic sequence at the Terrace of Vale Boi; and portable X-Ray Fluorescence (pXRF) to all previously mentioned samples, including additional chert samples from central Portugal (Rio Maior region) and southern Portugal (Betic systems).
+
+By refining the characterization of chert in the region and improving source attribution at the site, this research contributes to broader discussions on lithic procurement, human mobility, and exchange networks in southwestern Iberia during the UP.
+
+# Materials and methods
+
+## X-Ray Diffraction (XRD)
+
+X-ray diffraction (XRD) is a widely used laboratory technique for examining the atomic structure of crystalline materials (Adams, 2005).
+The minerals are crystalline, meaning their atoms are arranged in a three-dimensional repeating pattern.
+The crystal structure is defined by the arrangement of atoms within a single unit cell, as well as the cell’s size and shape.
+
+XRD works by directing a beam of X-rays onto a crystal, causing the rays to diffract and produce a characteristic diffraction pattern.
+The d-spacings—the distances between parallel planes of atoms in the crystal lattice—are derived from the angular positions and intensities of the diffracted beams.
+These d-spacings serve as a unique fingerprint for crystalline compounds, allowing for mineral identification (Marshall & Fairbridge, 1998).
+
+Although powder-XRD is a destructive technique, it has been used alongside petrographic characterisation (thin-section analysis) to gain a comprehensive understanding of the petrographic and mineral composition of lithic raw materials, such as chert, in both geological and archaeological contexts.
+This combined approach enables comparisons with potential archaeological implications, particularly in raw material sourcing studies in the Iberian Peninsula (e.g., Columbu et al., 2023; Gómez De Soler et al., 2023; Ortega et al., 2018).
+
+This study utilised X-ray diffraction (XRD) to determine the mineral composition of 23 geological powdered samples of chert from the Algarve region (Table 1).
+These samples, previously characterised and published by our team (Belmiro et al., 2023), represent at least one specimen from each outcrop, with a focus on primary and sub-primary locations across all chert-bearing formations in the region.
+A detailed description of the sample preparation process is available in our online research repository.
+
+The samples were analysed using a Bruker AXS D8 Discover XRD with the Da Vinci design, equipped with a Cu Kα source operating at 40 kV and 40 mA, and a Lynxeye 1D detector.
+Scans were conducted from 3° to 75° 2θ, with a 0.05° 2θ step, at the HERCULES Laboratory, University of Évora.
+
+Preliminary phase identification was performed using Diffrac.Suite™ (Bruker) and further refined with Profex (version 5.2.9) after converting raw files using PowDll Converter.
+All data processed in Profex was organised and plotted in R Studio, where visual representations of the XRD data were also generated.
+The code and original files used in this analysis are available in our online repository.
+
+Three additional samples (SP21, SP40, and a geological sample from the Triassic outcrop of Western Algarve) had previously been sent for XRD analysis and interpretation at Geociencias Barcelona-CSIC (Table 1).
+The analysis was conducted using a Bruker D8-A25 XRD, equipped with a Cu source (wavelength λ = 1.5405 Å) and an ultra-fast PSD detector (Lynxeye).
+The diffraction spectrum data were interpreted using the Diffrac.Suite software by Bruker by a technician.
+
+## Scanning Electron Microscopy-X-Ray Microscopy (SEM-EDS)
+
+Scanning electron microscopy coupled with energy-dispersive X-ray spectroscopy (SEM-EDS or SEM-EDX) provides various types of data by combining SEM imaging with elemental analysis.
+SEM is an electron microscopy technique that generates high-resolution images by directing a beam of electrons onto a specimen’s surface.
+As the electron beam interacts with the sample, it scatters as backscattered electrons and emits X-rays.
+As different atomic elements release characteristic X-ray at specific energies, it is possible to identify the elements present in the sample (Kirkbride, 2000) using an Energy dispersive detector (EDS).
+
+Although SEM-EDS is a versatile and efficient method, EDS has a relatively high chemical analysis  limit of detection (Kirkbride, 2000).
+
+Despite this limitation, EDS coupled with SEM has been widely used alongside other techniques, such as petrography, XRD, and XRF, to characterise geological and archaeological lithic resources, including chert raw materials.
+This combined approach enhances the understanding of a formation’s characteristics and aids in identifying sources in provenance studies in the Iberian Peninsula (e.g., Abrunhosa et al., 2020; Columbu et al., 2023; Costa et al., 2022; Ortega et al., 2018).
+
+This study employed scanning electron microscopy and energy-dispersive X-ray spectroscopy (SEM-EDS) to analyse the surface structure and elemental composition of the samples.
+The analysis included 39 geological samples from all five chert-bearing formations (Belmiro et al., 2023; Table 2a).
+Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Belmiro et al., 2025; Table 2b).
+
+```{r}
+#| label: tbl-SEM-samples
+#| tbl-cap: "Number of samples (chips) used for the SEM-EDS study, by a) geological formation for the geological samples, and by b) type for the archaeological samples."
+
+```
+
+The samples consisted of lithic chips mounted in resin and sanded until minimal to no scratches were visible under the microscope.
+A detailed description of the sample preparation process and analysis is available in our online repository (SOM_1).
+
+The chip analysis was conducted at the HERCULES Laboratory (Évora) using a variable pressure Hitachi S3700N SEM, coupled with a Quantax EDS microanalysis system.
+Specifically, the Quantax system is equipped with a Bruker AXS 5010 XFlash Silicon Drift Detector, which offers a spectral resolution of 129 eV at FWHM for MnKα.
+The EDS elemental analysis and quantification was carried out with the ESPRIT software by Bruker.
+
+The analysis focused on identifying representative traits of the samples for characterization and comparison.
+Images were captured at 250x and 1000x magnification for each analysed sample.
+To determine the elemental composition, elemental maps and spectra were generated, providing both visual and semi-quantitative data on the selected elements and sample regions.
+At least one map or spectrum was obtained for each distinct type of inclusion or structure identified in the samples.
+
+## Portable X-Ray Fluorescence (pXRF)
+
+X-ray Fluorescence (XRF) works by irradiating the specimen with X-rays, which causes electrons to attain an excited state.
+The specimen emits X-rays (fluorescence) when the atom returns to ground state; this energy is characteristic of the atoms present in the sample.
+As such, XRF spectra contain information about the chemical elements present in the analysed sample (Kirkbride, 2000).
+
+However, laboratory-based XRF has the limitation of being non-portable and requiring sample preparation, which is often incompatible with archaeological studies where assemblages cannot be transported or destroyed.
+In this context, portable XRF (pXRF) instruments offer a practical alternative.
+These devices are easily transportable, non-destructive, and have been successfully used in source discrimination studies, where they provide comparable results to their laboratory-based counterparts (Craig et al., 2007).
+
+pXRF has been used to discriminate between sources of various lithic raw materials, such as obsidian (e.g., Craig et al., 2007; Gemici et al., 2022; Nazaroff et al., 2010), as well as other materials like volcanic and metamorphic rocks (Egeland et al., 2019; Grave et al., 2012).
+
+Studies on chert archaeological assemblages in the Iberian Peninsula have also employed pXRF to distinguish between sources.
+When combined with other methods, pXRF has provided valuable insights (Costa et al., 2022; Gómez De Soler et al., 2023; Pereira et al., 2016, 2021).
+
+For this experiment, a Bruker portable XRF Titan S1 was used in a laboratory benchtop setup, powered by battery (with up to 25% battery charge before being replaced by a fully charged battery).
+The standard databases available in the equipment from Bruker were employed: Geochem application and the Dual Mining method.
+A detailed description of the preparation and analysis process is available in our online repository (SOM_1).
+
+Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 3).
+After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size (Newlander et al., 2015; SOM ++).
+
+For geological samples, fresh, flat surfaces were scanned, avoiding altered faces or cortex.
+When necessary, the samples were prepared by breaking the nodules.
+The samples selected represented all varieties of chert identified in the Algarve region, as well as in the archaeological record of Vale Boi.
+Additional samples were included from other regions, such as central Portugal and southern Spain, to facilitate comparisons and test hypotheses based on macroscopic and petrographic data.
+For archaeological samples, artefacts were chosen from previously identified types (Belmiro et al., 2025), prioritising larger, flatter morphologies with minimal surface alterations.
+
+The analysis and reporting protocol were established based on previous studies, emphasising both the accuracy of the obtained data and the transparency and reproducibility of the results (Johnson et al., 2024; Newlander et al., 2015).
+The analysis focused on descriptive statistics and the generation of ANOVAs to compare different groups.
+
+For further reproducibility and replicability, and working towards the goal of open science (Johnson et al., 2024; Marwick, 2017), all analyses were performed through R Studio.
+The R code used for the analysis, datasets, and visual representations presented in this paper are available through our online repository (++).
+We utilised the rrtools package by Marwick et al.
+(2018) to create a research compendium and produce a reproducible journal article.
+The repository includes the complete set of raw data used in the analysis, along with a custom R project (Wickham, 2015) that contains the necessary code to generate all tables and figures.
+To facilitate broad reuse, the code is provided under the MIT license, the data under CC‐0, and the figures under CC‐BY.
+For more details, see Marwick (2017).
 
@@
 
-All analysed samples are mainly composed of quartz; in all samples, the atomic percentage of oxygen is approximately double of the atomic percentage of silica, showing a 2:1 ratio (table +).
+The results presented in this section focus on the main observed patterns of the analysis between outcrops, formations and chert types.
+The detailed analysis/graphs regarding individual samples for all methods can be found in our online repository (SOM_2).
 
-Similarly, the presence of skeletal remains, represented by peaks of calcium, phosphorus and frequently accompanied by small peaks of sulphur, shows no patterns and occurs in samples from all formations and chert types.
+The XRD results are homogeneous, showing little differences between samples.
+The samples are predominantly composed of quartz.
+Dolomite and calcite were also identified in several samples (>7), and in the samples processed by the Geociencias Barcelona-CSIC, moganite and gypsum were identified.
+Despite this homogeneity, the results show some patterns in the presence of dolomite and calcite.
+
+```{r}
+#| label: tbl-xrd-results
+#| tbl-cap: "Identified crystalline phases of chert geological samples from the Algarve.
+LJW: Lower Jurassic West; MJE: Middle Jurassic East; TW: Triassic West; UPE: Upper Jurassic East; UPW: Upper Jurassic West.
+(*) refers to results obtained from samples analysed and interpreted at the Geociencias Barcelona-CSIC.
+Content percentages of samples analysed at the Hércules laboratory are represented by x: xxx refers to > 90%; xx to 90-30%; x to 10-1%; (x) to <1%."
+
+```
+
+Dolomite was mainly identified in the samples from the Lower Jurassic and Triassic outcrops of the western Algarve.
+Small dolomite peaks are present in more than ~60% of the Lower Jurassic cherts and the single sample from the Triassic.
+In comparison, only 1 of the 11 samples from the Middle and Upper Jurassic outcrops showed the presence of dolomite.
+
+```{r}
+#| label: fig-xrd-15
+#| fig-cap: "Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP15_PBX).
+The rhombus shape signifies the main peaks key to identify each mineral.
+Labels represent the d spacing values of key mineral peaks.
+Quantifications presented in the supporting table were obtained using Rietveld refinement."
+
+```
+
+In opposition, calcite was mainly identified in the Upper Jurassic samples and the Triassic samples.
+The primary and sub-primary cherts from the Jordana formation (Eastern Algarve) and the Upper Jurassic of Praia da Mareta (Western Algarve) all show the presence of calcite peaks.
+In comparison, only 1 sample out of 12 of the Lower Jurassic outcrops and 1 out of 5 of the Middle Jurassic outcrops show the presence of calcite.
+
+```{r}
+#| label: fig-xrd-58
+#| fig-cap: "Plot of minerals identified through XRD in the eastern Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP58_JOR).
+The rhombus shape signifies the main peaks key to identify each mineral.
+Labels represent the d spacing values of key mineral peaks.
+Quantifications presented in the supporting table were obtained using Rietveld refinement."
+
+```
+
+```{r}
+#| label: fig-xrd-69
+#| fig-cap: "Plot of minerals identified through XRD in the western Upper Jurassic (Malm) chert sample from the outcrop Praia da Mareta (SP69_MAR).
+The rhombus shape signifies the main peaks key to identify each mineral.
+Labels represent the d spacing values of key mineral peaks.
+Quantifications presented in the supporting table were obtained using Rietveld refinement."
+
+```
+
+The SEM-EDS results show that all analysed samples are mainly composed of SiO~2~.
+
+Similarly, the presence of skeletal remains, represented by peaks of calcium, phosphorus and frequently accompanied by small peaks of sulphur, occurs in samples from all formations and chert types.
 Although frequently without a discernible shape, oval fossils (although currently unidentified) were observed in LJW geological samples and archaeological samples from local cherts (T2 and T4).
@@
 In comparison to the frequent to common skeletal remains in the Jurassic cherts from the Algarve, and archaeological cherts attributed to local sources (T2-5), cherts from non-local sources such as T6-8 show little porosity and rare presence of inclusions.
-Whenever present, skeletal remains are small and show small peaks (\<3 cpv/ev).
+Whenever present, skeletal remains are small and show small peaks.
 
@@
 The rarity of inclusions is not limited to skeletal remains, but to other inclusions such as oxides, which in T6-8 are small and rare.
-Despite this difference in abundance and size, oxides are mainly composed of iron, often associated with other elements, possibly suggesting the existence of iron oxide (Fe2O3), pyrite (FeS2) and ilmenite (FeTiO3).
-Iron oxides were found both as inclusions scattered in the samples, but also frequently associated with skeletal remains, which may be related to their oxidization.
+Despite this difference in abundance and size, oxides are mainly composed of iron, often associated with other elements, possibly suggesting the existence of iron oxide (Fe~2~O~3~), pyrite (FeS~2~) and ilmenite (FeTiO~3~).
+Iron oxides were found both as inclusions scattered in the samples, but also frequently associated with skeletal remains, which may be related to their oxidisation.
 Pyrite was also identified in the shape of spherical pyrite framboids, as well as dispersed pyrite nanocrystals.
-These are characterised not only by large peaks of iron and associated sulphur, but also of round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids.
+These are characterised not only by large peaks of iron and associated sulphur, but also by round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids.
 These were present mainly in geological samples from the Upper Jurassic (both west and eastern formations), but barely identified in archaeological samples.
@@
 
-In a single sample from Lower Jurassic and samples from local T4 (both characterised by the previously mentioned oval fossils) inclusions composed of titanium (without associated iron) were identified, which may be related to the presence of rutile (TiO2).
-Small peaks of titanium (\<3 cpv/ev) were also identified in 3 out of 4 non-local T6 samples.
+In a single sample from Lower Jurassic and samples from local T4 (both characterised by the previously mentioned oval fossils) inclusions composed of titanium (without associated iron) were identified, which may be related to the presence of rutile (TiO~2~).
+Small peaks of titanium were also identified in 3 out of 4 non-local T6 samples.
 
 Other elements were identified through EDS maps and spectra (e.g., chlorine, potassium or magnesium), although their distribution also showed no patterns between formations and chert types.
-The exception are calcium and aluminium.
+The exceptions are calcium and aluminium.
 
@@
 
-Aluminium may be related to the speed of sedimentation; high peaks of aluminum may be related to fastest sedimentations speeds and may be a useful component to identify differences between cherts.
-Almost all chert samples show aluminum peaks, although mostly small (\< 3 cpv/ev).
-An exception are cherts from the Eastern basin of the Algarve (MJE and UJE), samples from non-local archaeological T7 and T8.
-The exception to the presence of aluminum is the chalcedony (both geological from TW and archaeological group T1) and all archaeological samples from the non-local group T6, with sources previously attributed to the Cretaceous cherts from Central Portugal.
+Aluminium may be related to the speed of sedimentation; low peaks of aluminium may be related to the fastest sedimentations speeds and may be a useful component to identify differences between cherts.
+Almost all chert samples show aluminium peaks, although mostly small.
+An exception is the cherts from the Eastern basin of the Algarve (MJE and UJE), samples from non-local archaeological T7 and T8 which show large peaks.
+The exception to the presence of aluminium is the chert of both geological samples from the Triassic formations and archaeological group T1 and all archaeological samples from the non-local group T6, with sources previously attributed to the Cretaceous cherts from Central Portugal.
+
+The pXRF measured 44 major and minor elements, although, for a large majority, these were mostly composed of values below the detection level (< LOD) or with values equal to 0.
+Only elements with more than 50% (n>80) of successful values were maintained in the analysis to allow the successful interpretation of the data (Fig.
+++).
+This list and a summarised description of the elements and their presence in sedimentary rocks/chert can be found in the supplementary materials (SOM ++).
+
+```{r}
+#| label: fig-pxrf-barplot
+#| fig-cap: "Comparison of success (values successfully measured by the pXRF, excluding values of 0 which were non-existent) and < LOD counts by element.
+Only elements with Success counts above 50% are included."
+
+```
+
+When focusing on the chosen elements, some patterns become clear; these can be seen as differences in the values between different types of chert, as well as similarities between geological and archaeological samples.
+
+In general, silica is the main component of all cherts.
+However, there is some variability and outliers in both silica and other elements.
+From these, LJW cherts (Lower Jurassic West) seem to show higher variability and more outliers.
+This may be related to the higher dolomitization process occurring in LJW cherts.
+Similarly, the lower mean values of silica for UJW (Upper Jurassic West) samples may be related to high values of calcium oxide (Fig.
+++).
+
+```{r}
+#| label: fig-pxrf-boxplots
+#| fig-cap: "Boxplots of raw element percentages (a: SiO2, b: CaO, c: Fe2O3, d: TiO2) by geological formation or archaeological chert type. Only chert groups with n>5 were included in the boxplots."
+
+```
+
+No clear patterns seem to occur between geological groups and archaeological groups, with the exception of the samples from central Portugal (CPT) and non-local archaeological types T6 and T7.
+This is more marked in Fe~2~O~3~ (Fig.
+++ b) and, to a lesser degree, in TiO~2~ (Fig.
+++ c), which show smaller means and values for these groups.
+The similarity between these groups can also be seen in the near absence of Sr; where other groups show smaller values but are well represented, the numbers of CPT, T6 and T8 samples are reduced to <50%, meaning Sr (Table s++) presence is limited in these cherts.
+Similarly, CPT and T6 cherts also show less than 50% samples with Zr (Table s++), making this element also limited in these chert groups.
+
+Finally, archaeological types T9 to T11 seem to show the most differences in element percentages of all analysed samples; due to the small sample, instead of boxplots (with a minimum in the sample of 5), the results are shown in Table s++ in the SOM.
+Albeit with some variation, T9 and T10 to show lower percentages of silica in comparison to other samples.
+As expected, other elements show much higher percentages, especially Fe~2~O~3~, TiO~2~, Al~2O~3 and K~2~O.
+T11 shows the lower percentages of silica (~90%; Table s++), and, similarly to T9-10 high peaks of Fe~2~O~3~, TiO~2~, K~2~O.
+Compared to all the other samples analysed in this study, T11 shows the highest values for Cu and Zr.
+However, these are still small percentages within the total chemical composition (<0.0006).
+
+Multicomponent statistical analysis was applied to the pXRF results, to understand the relation between the several elements and further identify patterns between the geological and archaeological types.
+Several PCAs were applied to the elements, however, only a few elements could consistently be used in this approach due to the high percentages of 0 or < LOD.
+Similarly, small groups (e.g., Triassic and non-local chert types such as T9 and T10) were not included in the PCAs due to the small number of samples.
+
+Fig ++ shows the PCAs for several combinations of elements and comparisons between different groups to understand the various relations.
+The elements used for the PCAs were S, Cl, P~2~O~5~, Fe~2~O~3~, TiO~2~, K~2~O, Cu and CaO.
+In all PCAs the two most representative dimensions are explained by the increased and decreased presence of all elements (Dim1), which may be explained by the increased or decreased values of SiO~2~, and the increased presence of CaO, Cl, S, P~2~O~5~ and Cu in comparison to Fe~2~O~3~, TiO~2~ and K~2~O (Dim2).
 
@@
 #| label: fig-pca
-#| fig-cap: Several pcas
+#| fig-cap: "Visual representations of dimension 1 and 2 of PCAs showing the correlation between elements and different chert groups.
+a) PCA using the geological cherts of the regional southern Portugal formations: Lower Jurassic west (LJW), Triassic west (TW), Upper Jurassic west (UPW), Middle Jurassic east (MJE) and Upper Jurassic east (UJE); b) PCA using geological samples of local Lower Jurassic cherts (LJW) and non-local Cretaceous cherts (central Portugal; CPT) and Upper Jurassic cherts (Betic Systems; UCA); c) PCA using geological samples of local Lower Jurassic cherts (LJW), geological chert samples from Cretaceous central Portugal formations (CPT), and archaeological samples from Vale Boi attributed to CPT sources (T6); d) PCA using geological samples of local Lower Jurassic cherts (LJW), geological chert samples from Upper Jurassic southern Spain formations (UCA), and archaeological samples from Vale Boi attributed to local sources (T2-5) and non-local, southern Spain sources (T7)."
 
@@
 ```
+
+When comparing the regional Algarve cherts (fig.
+++ a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
+By observing the scatterplots, it is noticeable that the Lower Jurassic cherts (LJW) are highly variable when compared to the other cherts, with samples which display higher positive values in Dim1 and Dim2.
+However, when focusing on the concentration of LJW results, they seem to fit within the range shown by Middle Jurassic and Upper Jurassic eastern cherts.
+In comparison, Upper Jurassic cherts from the western basin show higher positive values in Dim2, which is explained by the higher values of CaO.
+
+The differences between the local Lower Jurassic cherts and non-local cherts are more marked.
+Fig ++ b represents the correlation between elements and Lower Jurassic cherts, central Portugal Cretaceous cherts and southern Spain Upper Jurassic cherts.
+Dimension 1 explains ~49% of the variability, while dimension 2 explains 28.5%, with a combined percentage of `r 48.9+28.5`%.
+Once again, it is clear that Jurassic cherts, both from southwestern Portugal and southern Spain, show variability, although they are mostly located in the center/bottom of dimension 2 and center of dimension 1.
+When compared to these values, central Portugal Cretaceous cherts are located in the negative axis of dimension 1 and show little variability.
+
+A similar pattern emerges when applying a PCA using the same elements but comparing them to archaeological, non-local chert T6, previously attributed to central Portugal (Fig.
+++ c).
+In this PCA, dimension 1 explains 45.5% of the variability and dimension 2 29.2%, with a combined percentage of `r 45.5+29.2`%.
+Similarly to geological Cretaceous samples from central Portugal, T6 samples are also concentrated in the negative side of dimension 1, with a clear overlap between the two.
+
+As observed in Fig.
+++ b, Fig.
+++ d shows a concentration of all groups in the centre/bottom of dimension 2 (29.7%) and centre of dimension 1 (44.5%), with the first two dimensions explaining a combined percentage of `r 44.5+29.7`% of the variability.
+The archaeological local cherts T2 to T5 and non-local type T7 (previously attributed to the Upper Jurassic cherts of southern Spain) are congruent with the corresponding geological sources, although with clear decreased variability.
 
@@
 # Acknowledgements
+
+I would like to thank Dr.
+José Mirão for the training in the XRD and SEM-EDS methods and aid in the interpretation of the results, and the Hércules laboratory, for facilitating the equipment necessary for preparation and analyses (++).
+Still from the Hércules team, I thank the technician Sandra Velez for the training, help and bright mood during the preparation of SEM-EDS samples.
+In a similar fashion, I thank Dr.
+Javier Sanchéz-Martinéz for the help in obtaining chips for analysis through meticulous knapping whenever necessary.
+The preparation of bulk samples for XRD would also not have been possible without the help of the CIMA laboratory and their equipment.
+Finally, I would like to thank ICArEHB for providing the necessary equipment, contacts and researchers that allowed me to complete this study and explore the different techniques employed here.
+In particular, I would like to thank Dr.
+Hussein Kanbar for the support during the analysis and key insights for interpretation and visual presentation of geochemical data; and technician Pedro ++ and Dr.
+Vera Aldeias for the continuous support and advice.
 

### 2025-04-21 analysis/paper/paper.qmd
Plots in main paper improved. Removed Cu from the PCA analysis

@@
 library(magick)
+library(ggpubr)
 
@@
 sp58_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP58_xrd.csv')
-sp62_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP62_xrd.csv')
 sp69_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP69_xrd.csv')
@@
   geom_line(linewidth = 0.5) +
-  labs(x = "Diffraction Angle", y = "Intensity") +
+  labs(x = "Diffraction Angle", y = "Intensity (counts)") +
   theme_minimal() +
@@
   annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp58_perc)))
-
-# SP62 Oliveiras
-sp62_marks <- xrd_marks |> 
-  filter(Sample == "SP62")
-sp62_perc <- xrd_perc |> 
-  filter(Sample == "SP62") |> 
-  select(-Sample, -ESD)
-
-sp62_plot <- sp62_xrd |> 
-  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
-  select(-Quartz_nb) |> 
-  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
-               names_to = "Phase", 
-               values_to = "Count") |> 
-  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
-  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
-  geom_line(linewidth = 0.5) +
-  labs(x = "Diffraction Angle", y = "Intensity") +
-  theme_minimal() +
-  coord_cartesian(ylim = c(-5000, 50000)) +
-  geom_point(data = sp62_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
-  geom_label(data = sp62_marks, aes(x = Angle, y = Intensity, label = d), 
-            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
-  scale_colour_manual(values = c("lightsteelblue3", "salmon", "orchid","snow4")) +
-  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp62_perc)))
 
@@
   annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp69_perc)))
-
-```
-
-```{r}
-#| label: pxrf-read
-
-pxrf_raw <- read_csv('data/pxrf_data_dec.csv')
 
@@
 ```{r}
+#| label: pxrf-read
+
+pxrf_raw <- read_csv('analysis/data/raw_data/pxrf_data/pxrf_data_dec.csv')
+
+```
+
+```{r}
 #| label: pxrf-database-setup
@@
 ```{r}
-#| label: pxrf-fig-lod
-#| fig-cap: A barplot.
+#| label: pxrf-lod
 
@@
   mutate(Metric = factor(Metric, levels = c("Success", "< LOD"))) 
-lod_tab_table_long |> 
+succ_barplot <- lod_tab_table_long |> 
   ggplot(aes(x = Element, y = Count, fill = Metric)) +
@@
 
-# Create the bar plot
-ggplot(lod_tab_summary, aes(x = Element, y = Count, fill = Value)) +
-  geom_bar(stat = "identity", position = "stack") +  # Use stat = "identity" to use summarized counts
-  scale_fill_manual(values = c("Success" = "darkseagreen", "Success 0" = "lightgoldenrod1", "< LOD" = "indianred3")) +
-  labs(title = "Summarized Success and LOD Counts by Element",
-       y = "Count", 
-       fill = " ") +
-  theme_minimal() +
- # facet_wrap(~ Element, scales = "free_y") +  # Facet by Element with independent y-axes
-  theme(axis.text.x = element_text(hjust = 1),
-        axis.title.x = element_blank()) +
-  #geom_text(data = lod_tab_summary_filtered, aes(label = Count), position=position_stack(vjust=0.5), colour="white", size = 3) +
-  coord_flip() +
-  scale_y_continuous(breaks = seq(0, max(lod_tab_summary$Count), by = 20))
+```
+
+```{r}
+#| label: boxplots
+
+# Boxplot data preparation
+boxplots_pxrf <- pxrf_analysis |> 
+  select(-Sample_ID, -pXRF_ID) |> 
+  filter(!`Formation/Type` %in% c("TW", "T1", "T9", "T10", "T11")) |> # Removing chalcedony
+  mutate(across(-`Formation/Type`, ~ as.numeric(.)))
+  #pivot_longer(cols = c("MgO", "Al2O3", "SiO2", "P2O5", "S", "Cl", "K2O", "CaO", "TiO2", "Fe2O3", "Cu",
+        # "Sr", "Zr"), names_to = "Variable", values_to = "Value")
+boxplots_pxrf <- boxplots_pxrf |> 
+  mutate(`Formation/Type` = case_when(
+    `Formation/Type` == "T2" | `Formation/Type` == "T3" | `Formation/Type` == "T4" | `Formation/Type` == "T5" ~ "T2-5", TRUE ~ `Formation/Type`
+  )) |> 
+  mutate(`Formation/Type` = factor(`Formation/Type`, levels = c("LJW", "UJW", "MJE", "UJE", "CPT", "UCA", "T1", "T2-5", "T6", "T7", "T8")))
+
+## SiO2
+count_data_sio2 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
+  select(`Formation/Type`, SiO2) |> 
+  group_by(`Formation/Type`) |> 
+  filter(!is.na(SiO2)) |> 
+  summarise(n = n(), .groups = "drop")
+
+si_boxplot <- boxplots_pxrf |> 
+  ggplot(aes(x = `Formation/Type`, y = SiO2, fill = `Formation/Type`)) +
+  geom_boxplot(color = "#5C5C5C") +
+  stat_summary(fun = mean, geom = "point", shape = 18) +
+  #geom_text(data = count_data_sio2, aes(x = `Formation/Type`, y = 70, label = n), size = 3) +
+  coord_cartesian(ylim = c(70, 100)) +
+  theme_minimal() +
+  scale_fill_manual(values = c("LJW" = "#EDA92B", "UJW" = "#F1B6DA", "UJE" = "#8E95C8", "MJE" = "#BA68C8", "CPT" = "#4EB3D3", "UCA" = "#93C54E", "T2-5" = "#FEE391","T6" = "#9CCFD1", "T7" = "#B1C7AB", "T8" = "#D7617A")) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+
+## CaO
+count_data_cao <- boxplots_pxrf |> # Count number of observations per group where Value > 0
+  select(`Formation/Type`, CaO) |> 
+  group_by(`Formation/Type`) |> 
+  filter(!is.na(CaO)) |> 
+  summarise(n = n(), .groups = "drop")
+
+ca_boxplot <- boxplots_pxrf |> 
+  ggplot(aes(x = `Formation/Type`, y = CaO, fill = `Formation/Type`)) +
+  geom_boxplot(color = "#5C5C5C") +
+  stat_summary(fun = mean, geom = "point", shape = 18) +
+  #geom_text(data = count_data_cao, aes(x = `Formation/Type`, y = -0.3, label = n), size = 3) +
+  theme_minimal() +
+  scale_fill_manual(values = c("LJW" = "#EDA92B", "UJW" = "#F1B6DA", "UJE" = "#8E95C8", "MJE" = "#BA68C8", "CPT" = "#4EB3D3", "UCA" = "#93C54E", "T2-5" = "#FEE391","T6" = "#9CCFD1", "T7" = "#B1C7AB", "T8" = "#D7617A")) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+
+## TiO2
+count_data_tio2 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
+  select(`Formation/Type`, TiO2) |> 
+  group_by(`Formation/Type`) |> 
+  filter(!is.na(TiO2)) |> 
+  summarise(n = n(), .groups = "drop")
+
+ti_boxplot <- boxplots_pxrf |> 
+  ggplot(aes(x = `Formation/Type`, y = TiO2, fill = `Formation/Type`)) +
+  geom_boxplot(color = "#5C5C5C") +
+  stat_summary(fun = mean, geom = "point", shape = 18) +
+  #geom_text(data = count_data_tio2, aes(x = `Formation/Type`, y = -0.02, label = n), size = 3) +
+  theme_minimal() +
+  scale_fill_manual(values = c("LJW" = "#EDA92B", "UJW" = "#F1B6DA", "UJE" = "#8E95C8", "MJE" = "#BA68C8", "CPT" = "#4EB3D3", "UCA" = "#93C54E", "T2-5" = "#FEE391","T6" = "#9CCFD1", "T7" = "#B1C7AB", "T8" = "#D7617A")) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+
+## Fe2O3
+count_data_fe2o3 <- boxplots_pxrf |> # Count number of observations per group where Value > 0
+  select(`Formation/Type`, Fe2O3) |> 
+  group_by(`Formation/Type`) |> 
+  filter(!is.na(Fe2O3)) |> 
+  summarise(n = n(), .groups = "drop")
+
+fe_boxplot <- boxplots_pxrf |> 
+  ggplot(aes(x = `Formation/Type`, y = Fe2O3, fill = `Formation/Type`)) +
+  geom_boxplot(color = "#5C5C5C") +
+  stat_summary(fun = mean, geom = "point", shape = 18) +
+  #geom_text(data = count_data_fe2o3, aes(x = `Formation/Type`, y = 0, label = n), size = 3) +
+  coord_cartesian(ylim = c(0, 0.75)) +
+  theme_minimal() +
+  scale_fill_manual(values = c("LJW" = "#EDA92B", "UJW" = "#F1B6DA", "UJE" = "#8E95C8", "MJE" = "#BA68C8", "CPT" = "#4EB3D3", "UCA" = "#93C54E", "T2-5" = "#FEE391","T6" = "#9CCFD1", "T7" = "#B1C7AB", "T8" = "#D7617A")) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
 
@@
                 addEllipses = TRUE, 
-                ellipse.level = 0.95) +
+                ellipse.level = 0.95,
+                palette = c("LJW" = "#EDA92B", "UJW" = "#FF8989", "UJE" = "#5C6BC0", "MJE" = "#BA68C8", "TW" = "#EC407A"),
+                col.var = "#757575") +
+  scale_shape_manual(values=c(16, 17, 8, 18, 15)) +
   ggtitle(NULL)
@@
 pca_db_nl <- pxrf_raw |> 
-  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
+  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, S) |> # Removing CaO which is high NA in CPT
   filter(`Formation/Type` %in% c("LJW", "CPT", "UCA"))
@@
                 addEllipses = TRUE, 
-                ellipse.level = 0.95) +
+                ellipse.level = 0.80,
+                palette = c("LJW" = "#EDA92B", "CPT" = "#4EB3D3", "UCA" = "#93C54E"),
+                col.var = "#757575") +
+  scale_shape_manual(values=c(17, 16, 15)) +
   ggtitle(NULL)
@@
 pca_db_comp <- pxrf_raw |> 
-  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
+  select(`Formation/Type`, P2O5, K2O, TiO2, Fe2O3, S) |> # Removing CaO which is high NA in CPT
   mutate(`Formation/Type` = case_when(
@@
                 addEllipses = TRUE, 
-                ellipse.level = 0.80) +
+                ellipse.level = 0.80,
+                palette = c("LJW" = "#EDA92B", "T6" = "#7FCDBB", "CPT" = "#4EB3D3"),
+                col.var = "#757575") +
+  scale_shape_manual(values=c(17, 16, 15)) +
   ggtitle(NULL)
@@
 pca_db_comp2 <- pxrf_raw |> 
-  select(`Formation/Type`, P2O5, TiO2, K2O,Fe2O3, Cu, S) |> # Removing CaO which is high NA in CPT
+  select(`Formation/Type`, P2O5, TiO2, K2O,Fe2O3, S) |> # Removing CaO which is high NA in CPT
   mutate(`Formation/Type` = case_when(
@@
                 label = "var", 
+                repel = TRUE,
                 habillage = group_var_comp2,  # Use stored group column
                 addEllipses = TRUE, 
-                ellipse.level = 0.80) +
+                ellipse.level = 0.80,
+                palette = c("LJW" = "#EDA92B", "T2-5" = "#EF6C00", "T7" = "#00796B", "UCA" = "#93C54E"),
+                col.var = "#757575") +
+  scale_shape_manual(values=c(16, 17, 18, 15)) +
   ggtitle(NULL)
+
 # Add Group column back for visualization
@@
 The samples are predominantly composed of quartz.
-Dolomite and calcite were also identified in several samples (>7), and in the samples processed by the Geociencias Barcelona-CSIC, moganite and gypsum were identified.
+Dolomite and calcite were also identified in several samples (\>7), and in the samples processed by the Geociencias Barcelona-CSIC, moganite and gypsum were identified.
 Despite this homogeneity, the results show some patterns in the presence of dolomite and calcite.
@@
 Dolomite was mainly identified in the samples from the Lower Jurassic and Triassic outcrops of the western Algarve.
-Small dolomite peaks are present in more than ~60% of the Lower Jurassic cherts and the single sample from the Triassic.
+Small dolomite peaks are present in more than \~60% of the Lower Jurassic cherts and the single sample from the Triassic.
 In comparison, only 1 of the 11 samples from the Middle and Upper Jurassic outcrops showed the presence of dolomite.
@@
 
+plot(sp15_plot)
+
 ```
@@
 ```{r}
-#| label: fig-xrd-58
-#| fig-cap: "Plot of minerals identified through XRD in the eastern Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP58_JOR).
+#| label: fig-xrd-58-69
+#| fig-cap: "Plot of minerals identified through XRD in the a) eastern Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP58_JOR) and b) western Upper Jurassic (Malm) chert sample from the outcrop Praia da Mareta (SP69_MAR).
 The rhombus shape signifies the main peaks key to identify each mineral.
@@
 
-```
-
-```{r}
-#| label: fig-xrd-69
-#| fig-cap: "Plot of minerals identified through XRD in the western Upper Jurassic (Malm) chert sample from the outcrop Praia da Mareta (SP69_MAR).
-The rhombus shape signifies the main peaks key to identify each mineral.
-Labels represent the d spacing values of key mineral peaks.
-Quantifications presented in the supporting table were obtained using Rietveld refinement."
+plot_grid(sp58_plot, sp69_plot, labels = c("a", "b"), rows = 2)
 
@@
 
-The pXRF measured 44 major and minor elements, although, for a large majority, these were mostly composed of values below the detection level (< LOD) or with values equal to 0.
-Only elements with more than 50% (n>80) of successful values were maintained in the analysis to allow the successful interpretation of the data (Fig.
+The pXRF measured 44 major and minor elements, although, for a large majority, these were mostly composed of values below the detection level (\< LOD) or with values equal to 0.
+Only elements with more than 50% (n\>80) of successful values were maintained in the analysis to allow the successful interpretation of the data (Fig.
 ++).
@@
 Only elements with Success counts above 50% are included."
+
+plot(succ_barplot)
 
@@
 
+ggarrange(si_boxplot, ca_boxplot, fe_boxplot, ti_boxplot, labels = c("a","b","c","d"),
+          ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
 ```
@@
 ++ c), which show smaller means and values for these groups.
-The similarity between these groups can also be seen in the near absence of Sr; where other groups show smaller values but are well represented, the numbers of CPT, T6 and T8 samples are reduced to <50%, meaning Sr (Table s++) presence is limited in these cherts.
+The similarity between these groups can also be seen in the near absence of Sr; where other groups show smaller values but are well represented, the numbers of CPT, T6 and T8 samples are reduced to \<50%, meaning Sr (Table s++) presence is limited in these cherts.
 Similarly, CPT and T6 cherts also show less than 50% samples with Zr (Table s++), making this element also limited in these chert groups.
@@
 As expected, other elements show much higher percentages, especially Fe~2~O~3~, TiO~2~, Al~2O~3 and K~2~O.
-T11 shows the lower percentages of silica (~90%; Table s++), and, similarly to T9-10 high peaks of Fe~2~O~3~, TiO~2~, K~2~O.
+T11 shows the lower percentages of silica (\~90%; Table s++), and, similarly to T9-10 high peaks of Fe~2~O~3~, TiO~2~, K~2~O.
 Compared to all the other samples analysed in this study, T11 shows the highest values for Cu and Zr.
-However, these are still small percentages within the total chemical composition (<0.0006).
+However, these are still small percentages within the total chemical composition (\<0.0006).
 
 Multicomponent statistical analysis was applied to the pXRF results, to understand the relation between the several elements and further identify patterns between the geological and archaeological types.
-Several PCAs were applied to the elements, however, only a few elements could consistently be used in this approach due to the high percentages of 0 or < LOD.
+Several PCAs were applied to the elements, however, only a few elements could consistently be used in this approach due to the high percentages of 0 or \< LOD.
 Similarly, small groups (e.g., Triassic and non-local chert types such as T9 and T10) were not included in the PCAs due to the small number of samples.
@@
 Fig ++ shows the PCAs for several combinations of elements and comparisons between different groups to understand the various relations.
-The elements used for the PCAs were S, Cl, P~2~O~5~, Fe~2~O~3~, TiO~2~, K~2~O, Cu and CaO.
-In all PCAs the two most representative dimensions are explained by the increased and decreased presence of all elements (Dim1), which may be explained by the increased or decreased values of SiO~2~, and the increased presence of CaO, Cl, S, P~2~O~5~ and Cu in comparison to Fe~2~O~3~, TiO~2~ and K~2~O (Dim2).
+The elements used for the PCAs were S, Cl, P~2~O~5~, Fe~2~O~3~, TiO~2~, K~2~O and CaO.
+In all PCAs the two most representative dimensions are explained by the increased and decreased presence of all elements (Dim1), which may be explained by the increased or decreased values of SiO~2~, and the increased presence of CaO, Cl, S and P~2~O~5~ in comparison to Fe~2~O~3~, TiO~2~ and K~2~O (Dim2).
 
@@
 ++ a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
-By observing the scatterplots, it is noticeable that the Lower Jurassic cherts (LJW) are highly variable when compared to the other cherts, with samples which display higher positive values in Dim1 and Dim2.
+By observing the scatterplots, it is noticeable that the Lower Jurassic cherts (LJW) are highly variable when compared to the other cherts, with samples which display higher positive values in dimension 1 (Dim1) and dimension 2 (Dim2).
 However, when focusing on the concentration of LJW results, they seem to fit within the range shown by Middle Jurassic and Upper Jurassic eastern cherts.
-In comparison, Upper Jurassic cherts from the western basin show higher positive values in Dim2, which is explained by the higher values of CaO.
+In comparison, Upper Jurassic cherts from the western basin show higher positive values in dimension 2, which is explained by the higher values of CaO.
 
@@
 Fig ++ b represents the correlation between elements and Lower Jurassic cherts, central Portugal Cretaceous cherts and southern Spain Upper Jurassic cherts.
-Dimension 1 explains ~49% of the variability, while dimension 2 explains 28.5%, with a combined percentage of `r 48.9+28.5`%.
+Dimension 1 explains \~57% of the variability, while dimension 2 explains \~29%, with a combined percentage of `r 56.6+28.6`%.
 Once again, it is clear that Jurassic cherts, both from southwestern Portugal and southern Spain, show variability, although they are mostly located in the center/bottom of dimension 2 and center of dimension 1.
@@
 ++ c).
-In this PCA, dimension 1 explains 45.5% of the variability and dimension 2 29.2%, with a combined percentage of `r 45.5+29.2`%.
+In this PCA, dimension 1 explains \~52% of the variability and dimension 2 \~30%, with a combined percentage of `r 52.2+30.1`%.
 Similarly to geological Cretaceous samples from central Portugal, T6 samples are also concentrated in the negative side of dimension 1, with a clear overlap between the two.
@@
 ++ b, Fig.
-++ d shows a concentration of all groups in the centre/bottom of dimension 2 (29.7%) and centre of dimension 1 (44.5%), with the first two dimensions explaining a combined percentage of `r 44.5+29.7`% of the variability.
+++ d shows a concentration of all groups in the centre/bottom of dimension 2 (28.7%) and centre of dimension 1 (58.7%), with the first two dimensions explaining a combined percentage of `r 58.7+28.7`% of the variability.
 The archaeological local cherts T2 to T5 and non-local type T7 (previously attributed to the Upper Jurassic cherts of southern Spain) are congruent with the corresponding geological sources, although with clear decreased variability.

### 2025-04-21 analysis/supplementary-materials/appendix1.qmd
Plots in main paper improved. Removed Cu from the PCA analysis

@@
 sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
+sp62_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP62_xrd.csv')
 sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
@@
   annotate(geom = "table", x = 75, y = 60000, label = list(cbind(sp61_perc)))
+
+```
+
+```{r}
+#| label: fig-xrd-sp62
+#| fig-cap: Plot of minerals identified through XRD in the Middle Jurassic (Malhão formation) chert sample from the outcrop Oliveiras (SP7_CSV).
+The rhombus shape signifies main peaks key to identify each mineral.
+Labels represent the d spacing values of key mineral peaks.
+
+# SP62 Oliveiras
+sp62_marks <- xrd_marks |> 
+  filter(Sample == "SP62")
+sp62_perc <- xrd_perc |> 
+  filter(Sample == "SP62") |> 
+  select(-Sample, -ESD)
+
+sp62_plot <- sp62_xrd |> 
+  dplyr::mutate(Difference = Observed - Calculated - 3000) |> 
+  select(-Quartz_nb) |> 
+  pivot_longer(cols = c(Difference, Background, Observed, Quartz), 
+               names_to = "Phase", 
+               values_to = "Count") |> 
+  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Background", "Difference"))) |>
+  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
+  geom_line(linewidth = 0.5) +
+  labs(x = "Diffraction Angle", y = "Intensity (counts)") +
+  theme_minimal() +
+  coord_cartesian(ylim = c(-5000, 50000)) +
+  geom_point(data = sp62_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
+  geom_label(data = sp62_marks, aes(x = Angle, y = Intensity, label = d), 
+            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
+  scale_colour_manual(values = c("lightsteelblue3", "salmon", "orchid","snow4")) +
+  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp62_perc)))
 

### 2025-04-21 analysis/paper/paper.qmd
Desktop work session. Added tables to main paper and ESM_3.qmd

@@
 ---
-title: "Title Goes Here"
+title: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
 author:
@@
   keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
+
 ---
@@
 Keywords: `r rmarkdown::metadata$keywords`
-
-Highlights: `r rmarkdown::metadata$highlights`
 
@@
 
+```{r}
+#| label: tbl-samples
+#| tbl-cap: "Number of samples (chips) used for the SEM-EDS study, by a) geological formation for the geological samples, and by b) type for the archaeological samples."
+
+# Read table
+tbl_samples <- read_csv('analysis/data/raw_data/geochem_samples_n.csv')
+# Turn NAs into empty strings
+tbl_samples <- tbl_samples %>%
+  mutate(across(everything(), ~ replace(.x, is.na(.x), "")))
+
+knitr::kable(tbl_samples)
+
+```
+
 The samples were analysed using a Bruker AXS D8 Discover XRD with the Da Vinci design, equipped with a Cu Kα source operating at 40 kV and 40 mA, and a Lynxeye 1D detector.
@@
 Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Belmiro et al., 2025; Table 2b).
-
-```{r}
-#| label: tbl-SEM-samples
-#| tbl-cap: "Number of samples (chips) used for the SEM-EDS study, by a) geological formation for the geological samples, and by b) type for the archaeological samples."
-
-```
 
@@
 Content percentages of samples analysed at the Hércules laboratory are represented by x: xxx refers to > 90%; xx to 90-30%; x to 10-1%; (x) to <1%."
+
+# Read table
+xrd_results_tbl <- read_csv('analysis/data/raw_data/xrd_results.csv')
+# Turn NAs into empty strings
+xrd_results_tbl <- xrd_results_tbl %>%
+  mutate(across(everything(), ~ replace(.x, is.na(.x), "")))
+
+knitr::kable(xrd_results_tbl)
 

### 2025-04-21 analysis/supplementary-materials/ESM_3.qmd
Desktop work session. Added tables to main paper and ESM_3.qmd

---
title: "Online resource 3. Chemical elements/compounds and geological implications"
subtitle: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
---

**Silica (Silicon dioxide, SiO~2~)**

Silica is mostly found in nature as quartz and primarily composes cherts in a microcrystalline, cryptocrystalline or microfibrous form (Folk, 1980; Knauth, 1994). It is the key compound in chert diagenetic genesis, and can be expected to be found in high percentages within chert samples (Marshall & Fairbridge, 1998; Murray, 1994). Silica can also be found in limestone and takes a key role in the preservation of some fossils, through their silicification due to organic silica naturally occurring in living organisms (e.g., sponge spicules).

**Iron(II) oxide (Fe~2~O~3~)**

Iron is a component in all mineral classes, and it occurs mainly as reduced ferrous iron (Fe^2+^) or oxidized ferric iron (Fe^3+^) in sedimentary rocks (Marshall & Fairbridge, 1998). Iron oxides are a common constituent of cherts and are often related to metal input during the diagenetic formation process of cherts. Its presence has been used to identify the cherts diagenetic history due to being immobile during chert formation (Murray, 1994). The thermodynamically favoured phase of iron oxides in mostly sedimentary environments is hematite (Fe~2~O~3~). Ferric oxides in sedimentary environments with abundant organic matter may form Pyrite (FeS~2~) due to the interaction with sulphur (Marshall & Fairbridge, 1998).

```{r}
#| label: fig-scheme
#| fig-cap: "Schematic diagram of diagenetic chemical fractionations of chemical elements potentially occurring during chert formation. Adapted from Murray (1994)."

knitr::include_graphics("figures/SOM/murray_mobility.png")

```

**Titanium dioxide (TiO~2~)**

Titanium dioxide is frequently associated to terrigenous contribution. It is common in cherts, as it is considered immobile during diagenetic chert formation (Murray, 1994). Common Ti phases are the rutile (TiO~2~) and ilmenite (FeTiO~3~) (Marshall & Fairbridge, 1998). Enrichment of TiO~2~ in cherts (in comparison to Al) has been shown to be a useful indicator of specific chert diagenetic and weathering formation processes (Sugitani et al., 1996).

**Aluminium oxide (Al2O3)**

Aluminium (Al) is an abundant element in the Earth’s crust. Aluminium oxide is a common component of cherts, frequently associated with terrigenous fraction (Marshall & Fairbridge, 1998). It is considered that Al is immobile during diagenetic chert formation (Murray, 1994), and its correlation with other immobile elements (e.g., Ti) has been used to identify chert diagenetic and weathering processes (Sugitani et al., 1996). Similarly, Al~2~O~3~ concentrations in different types of cherts may vary (Murray, 1994), making them a key element to discern between sources, as shown in previous archaeological geochemical studies (Nazaroff et al., 2013).

**Phosphate (P~2~O~5~)**

Phosphate refers to the compound in which phosphorus (P) can be found in rocks and soils; P oxidizes to phosphate and is never found free in nature (Marshall & Fairbridge, 1998). This main form of P in nature is orthophosphate which is abundant in rocks, but is also abundant in aquatic environments (Rezania et al., 2021). Phosphorus is mobilized and removed during silicification of carbonate sediments, thus making its presence less stable than other elements (Murray, 1994).

**Sulphur (S)**

Sulphur is abundant in marine sediments as it is part of a complex cycle of chemical reactions that change the reactions and products when interacting with other elements, depending on the depth and type of sediment (Marshall & Fairbridge, 1998). Sulfur can also be incorporated into limestones as trace sulfate or be present in organic matter derived from microbial activity. In chert, sulfur is typically found in minor amounts, mainly as disseminated sulfides like pyrite.

**Chlorine (Cl)**

Chlorine (Cl) is typically present in sedimentary carbonate rocks in trace amounts, mainly as dissolved chloride (Cl⁻) in pore waters or as part of fluid inclusions within carbonate minerals. Chlorine is present in different abundances in different types of rocks. In sedimentary rocks, chlorine is more present in dolomites than in limestones (Johns & Huang, 1967). In chert, Cl is generally found in very low concentrations, primarily as adsorbed chloride ions in pore waters or as inclusions from saline fluids during silica diagenesis.

**Potassium oxide (K~2~O)**

Potassium (K) is a trace element in the Earth, but during the weathering of rocks dissolves into the weathering solutions and enters the hydrosphere; it is a major constituent of rock forming minerals (Marshall & Fairbridge, 1998). In sedimentary carbonate rocks, K~2~O is typically present in trace amounts, associated in clay minerals.

**Calcium oxide (CaO)**

Calcium (Ca) is a major element in most earth systems and major constituent of rock forming minerals. Calcium oxide (CaO) is an essential component of sedimentary carbonate rocks, primarily found in the form of calcium carbonate (CaCO~3~) in rocks such as calcite, dolomite and aragonite. The formation of carbonate rocks occurs through biological and chemical precipitation in marine and lacustrine environments, where CaO-rich minerals accumulate over time (Marshall & Fairbridge, 1998). Chert typically contains very low amounts of calcium oxide (CaO) compared to sedimentary carbonate rocks, due to its high mobilization during the diagenetic process (Murray, 1994). However, CaO can still be present in chert as a minor component, mainly due to impurities such as carbonate inclusions, detrital calcite, or interbedded carbonate layers. In some cases, cherts associated with limestone or dolomite may retain traces of CaO from the original carbonate material.

**Copper (Cu)**

Copper (Cu) is typically present in trace amounts in sedimentary carbonate rocks (and in natural waters), often as a result of diagenetic processes, hydrothermal activity, or detrital input. In carbonate formations, Cu can occur as dispersed ions, adsorbed onto clay minerals, or incorporated into secondary minerals, especially in oxidizing environments. In some cases, trace amounts of Cu can be incorporated into biogenic cherts due to the presence of organic matter or associated sulfide minerals (e.g., chalcopyrite, CuFeS~2~) (Marshall & Fairbridge, 1998).

**Strontium (Sr)**

Strontium is commonly present in sedimentary carbonate rocks. It readily substitutes for Ca in the crystal structure of carbonate minerals such as calcite, dolomite and aragonite, with aragonite generally containing higher Sr concentrations due to its greater affinity for Sr incorporation. Sr in carbonate rocks originates from seawater, biological processes (e.g., coral and shell formation), and diagenetic alteration. During diagenesis, Sr can be released from aragonite as it recrystallizes to calcite, leading to Sr-enriched pore waters or secondary mineral precipitation (De Vos and Tarvainen, 2006). In chert, Sr is typically present in trace amounts, often introduced through diagenetic interactions with surrounding carbonate sediments or the incorporation of biogenic materials, and has been used to discern between cherts of different formations (Yamamoto, 1986).

**Zirconium (Zr)**

Zirconium is typically present in very low concentrations in sedimentary carbonate rocks. Zr is commonly associated with detrital heavy minerals such as zircon (ZrSiO~4~), which can be present as fine-grained inclusions in carbonate sediments, especially in siliciclastic-influenced depositional environments. Due to its resistance to weathering and diagenesis, zircon grains in carbonate rocks often originate from the erosion of older igneous and metamorphic rocks (De Vos and Tarvainen, 2006). In chert, Zr is also typically present in trace amounts, and has been used to successfully discern between different chert sources (Nazaroff et al., 2013), due to it being relatively unaffected during chert diagenesis (Murray, 1994).

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

### 2025-04-22 analysis/supplementary-materials/ESM_3.qmd
Macbook work session. ESM 3 updated and finished

@@
 subtitle: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
+author: Joana Belmiro
 date: now
@@
 execute:
-  echo: true
+  echo: false
   warning: false
@@
 csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
 ---
 
-**Silica (Silicon dioxide, SiO~2~)**
+### Aluminium oxide (Al~2~O~3~)
 
-Silica is mostly found in nature as quartz and primarily composes cherts in a microcrystalline, cryptocrystalline or microfibrous form (Folk, 1980; Knauth, 1994).
-It is the key compound in chert diagenetic genesis, and can be expected to be found in high percentages within chert samples (Marshall & Fairbridge, 1998; Murray, 1994).
-Silica can also be found in limestone and takes a key role in the preservation of some fossils, through their silicification due to organic silica naturally occurring in living organisms (e.g., sponge spicules).
-
-**Iron(II) oxide (Fe~2~O~3~)**
-
-Iron is a component in all mineral classes, and it occurs mainly as reduced ferrous iron (Fe^2+^) or oxidized ferric iron (Fe^3+^) in sedimentary rocks (Marshall & Fairbridge, 1998).
-Iron oxides are a common constituent of cherts and are often related to metal input during the diagenetic formation process of cherts.
-Its presence has been used to identify the cherts diagenetic history due to being immobile during chert formation (Murray, 1994).
-The thermodynamically favoured phase of iron oxides in mostly sedimentary environments is hematite (Fe~2~O~3~).
-Ferric oxides in sedimentary environments with abundant organic matter may form Pyrite (FeS~2~) due to the interaction with sulphur (Marshall & Fairbridge, 1998).
+Aluminium (Al) is an abundant element in the Earth’s crust.
+Aluminium oxide is a common component of cherts, frequently associated with terrigenous fraction [@encyclop1998].
+It is considered that Al is immobile during diagenetic chert formation (@fig-scheme), and its correlation with other immobile elements (e.g., Ti) has been used to identify chert diagenetic and weathering processes [@sugitani1996].
+Similarly, Al~2~O~3~ concentrations in different types of cherts may vary [@murray1994], making them a key element to discern between sources, as shown in previous archaeological geochemical studies [@nazaroff2013].
 
@@
 
-knitr::include_graphics("figures/SOM/murray_mobility.png")
+knitr::include_graphics(here::here('analysis/figures/SOM/murray_mobility.png'))
 
@@
 
-**Titanium dioxide (TiO~2~)**
+### Calcium oxide (CaO)
 
-Titanium dioxide is frequently associated to terrigenous contribution.
-It is common in cherts, as it is considered immobile during diagenetic chert formation (Murray, 1994).
-Common Ti phases are the rutile (TiO~2~) and ilmenite (FeTiO~3~) (Marshall & Fairbridge, 1998).
-Enrichment of TiO~2~ in cherts (in comparison to Al) has been shown to be a useful indicator of specific chert diagenetic and weathering formation processes (Sugitani et al., 1996).
+Calcium (Ca) is a major element in most earth systems and major constituent of rock forming minerals.
+Calcium oxide (CaO) is an essential component of sedimentary carbonate rocks, primarily found in the form of calcium carbonate (CaCO~3~) in rocks such as calcite, dolomite and aragonite.
+The formation of carbonate rocks occurs through biological and chemical precipitation in marine and lacustrine environments, where CaO-rich minerals accumulate over time [@encyclop1998].
+Chert typically contains very low amounts of calcium oxide (CaO) compared to sedimentary carbonate rocks, due to its high mobilization during the diagenetic process (@fig-scheme).
+However, CaO can still be present in chert as a minor component, mainly due to impurities such as carbonate inclusions, detrital calcite, or interbedded carbonate layers.
+In some cases, cherts associated with limestone or dolomite may retain traces of CaO from the original carbonate material.
 
-**Aluminium oxide (Al2O3)**
-
-Aluminium (Al) is an abundant element in the Earth’s crust.
-Aluminium oxide is a common component of cherts, frequently associated with terrigenous fraction (Marshall & Fairbridge, 1998).
-It is considered that Al is immobile during diagenetic chert formation (Murray, 1994), and its correlation with other immobile elements (e.g., Ti) has been used to identify chert diagenetic and weathering processes (Sugitani et al., 1996).
-Similarly, Al~2~O~3~ concentrations in different types of cherts may vary (Murray, 1994), making them a key element to discern between sources, as shown in previous archaeological geochemical studies (Nazaroff et al., 2013).
-
-**Phosphate (P~2~O~5~)**
-
-Phosphate refers to the compound in which phosphorus (P) can be found in rocks and soils; P oxidizes to phosphate and is never found free in nature (Marshall & Fairbridge, 1998).
-This main form of P in nature is orthophosphate which is abundant in rocks, but is also abundant in aquatic environments (Rezania et al., 2021).
-Phosphorus is mobilized and removed during silicification of carbonate sediments, thus making its presence less stable than other elements (Murray, 1994).
-
-**Sulphur (S)**
-
-Sulphur is abundant in marine sediments as it is part of a complex cycle of chemical reactions that change the reactions and products when interacting with other elements, depending on the depth and type of sediment (Marshall & Fairbridge, 1998).
-Sulfur can also be incorporated into limestones as trace sulfate or be present in organic matter derived from microbial activity.
-In chert, sulfur is typically found in minor amounts, mainly as disseminated sulfides like pyrite.
-
-**Chlorine (Cl)**
+### Chlorine (Cl)
 
@@
 Chlorine is present in different abundances in different types of rocks.
-In sedimentary rocks, chlorine is more present in dolomites than in limestones (Johns & Huang, 1967).
+In sedimentary rocks, chlorine is more present in dolomite than in limestone [@johns1967].
 In chert, Cl is generally found in very low concentrations, primarily as adsorbed chloride ions in pore waters or as inclusions from saline fluids during silica diagenesis.
 
-**Potassium oxide (K~2~O)**
-
-Potassium (K) is a trace element in the Earth, but during the weathering of rocks dissolves into the weathering solutions and enters the hydrosphere; it is a major constituent of rock forming minerals (Marshall & Fairbridge, 1998).
-In sedimentary carbonate rocks, K~2~O is typically present in trace amounts, associated in clay minerals.
-
-**Calcium oxide (CaO)**
-
-Calcium (Ca) is a major element in most earth systems and major constituent of rock forming minerals.
-Calcium oxide (CaO) is an essential component of sedimentary carbonate rocks, primarily found in the form of calcium carbonate (CaCO~3~) in rocks such as calcite, dolomite and aragonite.
-The formation of carbonate rocks occurs through biological and chemical precipitation in marine and lacustrine environments, where CaO-rich minerals accumulate over time (Marshall & Fairbridge, 1998).
-Chert typically contains very low amounts of calcium oxide (CaO) compared to sedimentary carbonate rocks, due to its high mobilization during the diagenetic process (Murray, 1994).
-However, CaO can still be present in chert as a minor component, mainly due to impurities such as carbonate inclusions, detrital calcite, or interbedded carbonate layers.
-In some cases, cherts associated with limestone or dolomite may retain traces of CaO from the original carbonate material.
-
-**Copper (Cu)**
+### Copper (Cu)
 
@@
 In carbonate formations, Cu can occur as dispersed ions, adsorbed onto clay minerals, or incorporated into secondary minerals, especially in oxidizing environments.
-In some cases, trace amounts of Cu can be incorporated into biogenic cherts due to the presence of organic matter or associated sulfide minerals (e.g., chalcopyrite, CuFeS~2~) (Marshall & Fairbridge, 1998).
+In some cases, trace amounts of Cu can be incorporated into biogenic cherts due to the presence of organic matter or associated sulphide minerals (e.g., chalcopyrite, CuFeS~2~) [@encyclop1998].
 
-**Strontium (Sr)**
+### Iron(II) oxide (Fe~2~O~3~)
+
+Iron is a component in all mineral classes, and it occurs mainly as reduced ferrous iron (Fe^2+^) or oxidized ferric iron (Fe^3+^) in sedimentary rocks [@encyclop1998].
+Iron oxides are a common constituent of cherts and are often related to metal input during the diagenetic formation process of cherts.
+Its presence has been used to identify the cherts diagenetic history due to being immobile during chert formation (@fig-scheme).
+The thermodynamically favoured phase of iron oxides in mostly sedimentary environments is hematite (Fe~2~O~3~).
+Ferric oxides in sedimentary environments with abundant organic matter may form Pyrite (FeS~2~) due to the interaction with sulphur [@encyclop1998].
+
+### Phosphate (P~2~O~5~)
+
+Phosphate refers to the compound in which phosphorus (P) can be found in rocks and soils; P oxidizes to phosphate and is never found free in nature [@encyclop1998].
+This main form of P in nature is orthophosphate which is abundant in rocks, but is also abundant in aquatic environments [@rezania2021].
+Phosphorus is mobilized and removed during silicification of carbonate sediments, thus making its presence less stable than other elements (@fig-scheme).
+
+### Potassium oxide (K~2~O)
+
+Potassium (K) is a trace element in the Earth, but during the weathering of rocks dissolves into the weathering solutions and enters the hydrosphere; it is a major constituent of rock forming minerals [@encyclop1998].
+In sedimentary carbonate rocks, K~2~O is typically present in trace amounts, associated in clay minerals.
+
+### Silica (Silicon dioxide, SiO~2~)
+
+Silica is mostly found in nature as quartz and primarily composes cherts in a microcrystalline, cryptocrystalline or microfibrous form [@folk1980; @knauth1994].
+It is the key compound in chert diagenetic genesis and can be expected to be found in high percentages within chert samples [@encyclop1998; @murray1994].
+Silica can also be found in limestone and takes a key role in the preservation of some fossils, through their silicification due to organic silica naturally occurring in living organisms (e.g., sponge spicules).
+
+### Strontium (Sr)
 
@@
 Sr in carbonate rocks originates from seawater, biological processes (e.g., coral and shell formation), and diagenetic alteration.
-During diagenesis, Sr can be released from aragonite as it recrystallizes to calcite, leading to Sr-enriched pore waters or secondary mineral precipitation (De Vos and Tarvainen, 2006).
-In chert, Sr is typically present in trace amounts, often introduced through diagenetic interactions with surrounding carbonate sediments or the incorporation of biogenic materials, and has been used to discern between cherts of different formations (Yamamoto, 1986).
+During diagenesis, Sr can be released from aragonite as it recrystallizes to calcite, leading to Sr-enriched pore waters or secondary mineral precipitation [@devos2006].
+In chert, Sr is typically present in trace amounts, often introduced through diagenetic interactions with surrounding carbonate sediments or the incorporation of biogenic materials and has been used to discern between cherts of different formations [@yamamoto1986].
 
-**Zirconium (Zr)**
+### Sulphur (S)
+
+Sulphur is abundant in marine sediments as it is part of a complex cycle of chemical reactions that change the reactions and products when interacting with other elements, depending on the depth and type of sediment [@encyclop1998].
+Sulphur can also be incorporated into limestone as trace sulphate or be present in organic matter derived from microbial activity.
+In chert, sulphur is typically found in minor amounts, mainly as disseminated sulphides like pyrite.
+
+### Titanium dioxide (TiO~2~)
+
+Titanium dioxide is frequently associated to terrigenous contribution.
+It is common in cherts, as it is considered immobile during diagenetic chert formation (@fig-scheme).
+Common Ti phases are the rutile (TiO~2~) and ilmenite (FeTiO~3~) [@encyclop1998].
+Enrichment of TiO~2~ in cherts (in comparison to Al) has been shown to be a useful indicator of specific chert diagenetic and weathering formation processes [@sugitani1996].
+
+### Zirconium (Zr)
 
@@
 Zr is commonly associated with detrital heavy minerals such as zircon (ZrSiO~4~), which can be present as fine-grained inclusions in carbonate sediments, especially in siliciclastic-influenced depositional environments.
-Due to its resistance to weathering and diagenesis, zircon grains in carbonate rocks often originate from the erosion of older igneous and metamorphic rocks (De Vos and Tarvainen, 2006).
-In chert, Zr is also typically present in trace amounts, and has been used to successfully discern between different chert sources (Nazaroff et al., 2013), due to it being relatively unaffected during chert diagenesis (Murray, 1994).
+Due to its resistance to weathering and diagenesis, zircon grains in carbonate rocks often originate from the erosion of older igneous and metamorphic rocks [@devos2006].
+In chert, Zr is also typically present in trace amounts, and has been used to successfully discern between different chert sources [@nazaroff2013], due to it being relatively unaffected during chert diagenesis [@murray1994].
 

### 2025-04-22 analysis/supplementary-materials/ESM_1.qmd
SOM detailed methods and sample prep added, finished and rendered

---
title: "Online resource 1. Sample preparation and detailed method description"
subtitle: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
author: Joana Belmiro
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

## X-Ray diffraction (XRD)

To obtain a representative collection of cherts with diverse macroscopic characteristics—particularly those displaying intra-outcrop variability—samples were selected from all four chert-bearing geological formations identified in the Algarve region.

All samples underwent manual preparation following standardized procedures to ensure optimal and reproducible results. The process began with crushing the samples into small chips, during which any cortex or alteration rinds were carefully removed. This step ensured that only unaltered areas of the nodules were used. A steel geological hammer was employed for this purpose, and the work was carried out on a clean plastic surface, which was either cleaned or replaced between samples to avoid cross-contamination.

Following the initial preparation, the chips were ground into a fine, homogeneous powder using a planetary ball mill equipped with agate balls. The equipment and tools were thoroughly cleaned and dried after each sample to prevent contamination. The resulting powders were stored in individual plastic vials, each clearly labelled with the corresponding sample ID.

The initial sample preparation was conducted at the ICArEHB laboratories (Interdisciplinary Centre for Archaeology and the Evolution of Human Behaviour), while bulk sample processing took place at the Laboratory of Environmental Sedimentology at CIMA (Centro de Investigação Marinha e Ambiental), University of Algarve (https://www.cima.ualg.pt/pt/cima/infraestruturas).

To ensure the reproducibility of results across different XRD instruments and researchers, larger quantities of chips were processed, allowing for the preparation of multiple bulk samples from the same nodules and the same internal areas within each nodule.

X-ray diffraction (XRD) analysis was performed using a Bruker AXS D8 Discover system with Da Vinci design, equipped with a Cu Kα source (operating at 40 kV and 40 mA) and a Lynxeye one-dimensional detector. Scans were conducted over a 2θ range of 3° to 75°, with a step size of 0.05°, at the HERCULES Laboratory at the University of Évora.

Following analysis, bulk samples were returned to their original containers. Standard circular XRD mounts were cleaned and dried between uses to prevent contamination.

Preliminary phase identification was performed using Bruker's Diffrac.Suite™ software, available at the HERCULES Laboratory. The data, initially generated in Bruker's proprietary format, were converted to raw files using the free PowDll Converter software to enable further analysis using open-source tools.

Phase identification was completed using Profex (version 5.2.9), with the Bruker Lynxeye Cu source configuration file, utilizing the Search-Match functionality. Mineral quantification was performed with the Rietveld Refinement functionality. All collected data from Profex were organized and visualized using R Studio.

## Scanning Electron Microscopy and Energy Dispersive X-Ray (SEM-EDS)

SEM-EDS analysis was conducted on chert chips mounted in resin molds. Several small chips (1–5 mm in size) were carefully selected from the previously prepared samples.

For geological specimens, chip selection followed the same procedure as that used for XRD sample preparation. Selection criteria also included chip flatness, appropriate thickness, and the macroscopic characteristics of the nodules to ensure a representative sampling of chert variability within each formation. Selected chips were stored individually for further processing, while remaining material was retained for powdering.

For archaeological specimens, chips were chosen based on the morphology of the lithic artifacts, with preference given to edges that allowed for chip removal with minimal damage. Unretouched artifacts from various archaeological layers were prioritized for sampling.

Chips were extracted using either a small metal cutter or by controlled knapping with a non-chert stone or antler. A small marker dot was placed on each artifact to indicate the area from which the chip was removed, and the artifact was labelled with a reference to the corresponding chip ID. This procedure ensured clear identification of modified archaeological artifacts for future reference.

```{r}
#| label: fig-moulds
#| fig-cap: "Resin moulds with chert chips after polishing. All moulds and chips were given an individual ID (e.g., M3_1)."

knitr::include_graphics(here::here('analysis/figures/SOM/sample_scheme.png'))

```

The selected chips were then mounted in plastic moulds filled with transparent epoxy resin (Struers Epofix). Chip arrangement within each mould was carefully planned to optimize visibility and identification during analysis—typically in alternating rows of even and uneven numbers when feasible. Prior to resin casting, the arranged chips were photographed for documentation. The resin mixture (resin and hardener) was then poured into the mould and left to cure for 8–12 hours. This step was conducted at both the HERCULES Laboratory (University of Évora) with technical assistance, and the ICArEHB Geochemistry Laboratory (University of Algarve).

Once the resin had fully hardened, the moulds were removed, and the chips were cut and sanded to produce a uniform height. Cutting was performed using a precision saw at the HERCULES Laboratory. Each chip was labelled on the cut surface and sides with the corresponding mould ID and an orientation arrow referencing the earlier photographs.

The exposed chip faces were then sanded to improve surface quality for SEM-EDS analysis. Sanding was carried out using a water-cooled sanding machine at the ICArEHB Geoarchaeology Laboratory, employing a sequence of progressively finer sandpaper grits (e.g., 240, 800, 1000, 1200, 2400, and 4000). Final polishing was done using a cloth and a liquid diamond polishing compound to achieve a smooth, reflective surface.

After each round of sanding, samples were inspected under a microscope to check for residual surface scratches. If needed, additional sanding was performed until a flawless finish was achieved. Once polishing was complete, the moulds were secured with double-sided tape inside small boxes for storage and transport. This ensured the polished surfaces remained protected from contact with rough materials, minimizing the risk of scratches during handling.

## Portable X-Ray Fluorescence (pXRF)

Portable X-ray fluorescence (pXRF) analysis was carried out using a Bruker Titan S1 portable XRF device, operated in a benchtop configuration at the ICArEHB laboratory. The instrument was powered by a battery—maintained at a minimum 25% charge before being replaced with a fully charged unit. Battery operation was preferred over direct power connection due to a faulty power cable and to avoid potential power fluctuations during the analysis process.

Prior to sample analysis, a validation run was conducted using two standard reference materials provided by Bruker, applying the standard calibration. The Bruker standard database was utilized with the Geochem application and the Dual Mining method.

A total of 166 chert samples were analysed, representing a range of geological and archaeological sources and types. Each sample was scanned for a total of 180 seconds: 90 seconds in the first phase targeting major elements, and 90 seconds in the second phase targeting minor elements. A minimum of one scan was performed per sample, with additional scans conducted on specimens displaying macroscopic variability (e.g., differing colours or degrees of translucency).

Following each scan, the analysed point was marked with an "X" directly on the sample surface. Additionally, the sample bags were labelled with a unique pXRF ID code. In cases where multiple scans were conducted on a single sample, corresponding IDs were also marked adjacent to each "X" to ensure accurate identification.

Prior to storage, the thickness and diameter of the scanned face of each sample were measured to confirm compliance with the minimum requirements for pXRF analysis. This step was guided by findings from previous research [e.g., @newlander2015], which indicate that sample size and thickness can significantly affect data homogeneity and analytical accuracy. Complete measurements for each sample, including thickness and diameter, are provided in Online resource 2.

For geological samples, scans were conducted on fresh, flat surfaces, deliberately avoiding areas with cortex or signs of alteration. When necessary, nodules were broken to expose suitable surfaces. The selected geological specimens were chosen to represent the full spectrum of chert types identified in the Algarve region, as well as comparative samples from Central Portugal and southern Spain. This broad selection enabled comparative analysis and supported hypotheses grounded in both macroscopic and petrographic data.

For archaeological samples, artifacts were selected based on previously established typologies [@belmiro2025], with priority given to pieces that exhibited larger, flatter surfaces and minimal surface alteration. This ensured reliable analytical results while preserving the integrity of the artifacts.

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

### 2025-04-23 analysis/supplementary-materials/appendix1.qmd
Desktop work session. Added tables to main paper and ESM_3.qmd
SOM results continued. Added geological SEM text and chunks.

@@
 
-# Supplementary materials 1
+# Online resource 4
 
@@
 
-#### Lower Jurassic of western Algarve
-
-The analysed samples from the Lower Jurassic outcrops are homogeneous regarding their structure, inclusions and mineral compositions.
+#### Lower Jurassic of western Algarve (LJW)
+
+The analysed samples from the Lower Jurassic outcrops (Carixian) are homogeneous regarding their structure, inclusions and mineral compositions.
 The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group (Carixian, Lower Jurassic formations of western Algarve, including the inland outcrop of Ferrel).
@@
 ```
+
+#### Triassic of western Algarve (TW)
+
+#### Upper Jurassic of western Algarve (UJW)
+
+The analysed samples from the Upper Jurassic outcrops of western Algarve (Malm) are homogeneous regarding their structure, inclusions and mineral compositions.
+
+They are mainly composed of cryptocrystalline quartz (SiO2), and small peaks of aluminium.
+Small peaks of magnesium were also identified, and calcite was identified in a single sample.
+In this sample, calcite represents the only identified inclusions, which seem to show varied shapes, which may be skeletal remains.
+
+Skeletal remains are rare, occurring in a single sample.
+These are composed of calcium, phosphorus, and also show an associated smaller peak of sulphur.
+These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the skeletal remains (@fig-m24-1).
+These skeletal remains are frequently irregular and indeterminate.
+
+Finally, iron oxides are present in 3 of 4 of the samples, with an iron (Fe) composition, but also iron and sulphur or iron and titanium (@fig-m24-4).
+Although characterized by irregular iron inclusions, iron is also found in framboids.
+Of the 4 samples, 3 show the presence of spherical framboids, as well as dispersed nanocrystals (in association with spherical framboids; @fig-m24-1-m24-2).
+
+```{r}
+#| label: fig-m24-1
+#| fig-cap: "A) general view of the sample (x250 magnification).
+B) detail of a fossil (x400 magnification) composed of calcium (B1), phosphorus (B2), sulphur (B3) and iron (B4) possibly the result of the oxidation.
+Geological sample SP69_MAR from the UJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M24_1.png"))
+
+```
+
+```{r}
+#| label: fig-m24-4
+#| fig-cap: "A) general view of the sample (x250 magnification) showing an inclusion.
+Round and oval shapes in lighter grey may be calcispheres identified in thin-section.
+B) detailed view of the inclusion (x950 magnification) composed of titanium (B1) and iron, as well as scattered, irregular iron oxides (B2).
+Geological sample SP69_MAR from the UJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M24_4_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m24-1-m24-2
+#| fig-cap: "A) detailed view of iron inclusions (x1.8k magnification), possibly originally spherical pyrite framboids (black arrow) and dispersed pyrite nanocrystals (white arrow).
+B) detail of an irregular inclusion (x2.1k magnification) composed mainly of iron (1) and inclusions composed of iron, possibly originally dispersed pyrite nanocrystals (2).
+Dot represents the location of the measured spectrum plotted in graph 1 and 2.
+Geological sample SP69_MAR from the UJW formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M24_1_M24_2_graphs.png"))
+
+```
+
+#### Middle Jurassic of eastern Algarve (MJE)
+
+The analysed samples from the Middle Jurassic outcrops (Malhão formation) are somewhat variable in their structure, inclusions and mineral compositions.
+The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group.
+
+They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and small and large peaks of aluminium. Small peaks of chlorine and potassium can also be identified.
+
+Skeletal remains occur in less than 50% of the samples.
+These are all composed of calcium and phosphorus and frequently show an associated smaller peak of sulphur (@fig-m8-2-mp1).
+These skeletal remains are all irregular and indeterminate.
+
+Iron oxides are present in more than 90% of the samples, with an iron (Fe) composition (@fig-m12-5-s1).
+Other types of iron oxides are also present in lesser frequencies, such as equally large peaks of iron and titanium (20% of the samples; @fig-m12-5-s1) and titanium with a small peak of iron (20%; @fig-m7-3).
+EDS results also show a small subset of samples with a small peak of calcium frequently associated with iron inclusions (@fig-m12-3-mp1) and inclusions which are mainly composed of large peaks of potassium, aluminium, and magnesium and small peaks of iron or iron with titanium (@fig-m12-1-s2).
+Most of these inclusions have irregular shapes and varying sizes, although framboids are also present.
+
+Spherical framboids are present in 40% of the samples, and dispersed nanocrystals are present in 10% of the samples (in association with spherical framboids), and composed of iron, aluminium and a small peak of calcium and magnesium (@fig-m12-5-s6; @fig-m8-1; @fig-m12-3-s3).
+
+```{r}
+#| label: fig-m8-2-mp1
+#| fig-cap: "A) general view of the sample (x250 magnification) showing several inclusions and porosity.
+B) detail of a fossil (x1k magnification) composed mainly of calcium (B1), phosphorus (B2) and small peaks of sulphur (B3; 1).
+Dot represents the location of the measured spectrum plotted in graph 1.
+Geological sample SP49_CdC from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M8_2_map1_spectra1_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m12-5-s1
+#| fig-cap: "A) general view of the sample (x450 magnification) showing several inclusions.
+B) detail of small inclusions (x1.3k magnification) composed of titanium, iron and a large peak of aluminium (1) and iron (2).
+C) detail of the large inclusion (x3.5k magnification) composed mainly of iron (3).
+Dot represents the location of the measured spectra plotted in graphs 1-3.
+Geological sample SP50_CdC from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M12_5_s1_s3_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m7-3
+#| fig-cap: "A) detail of the sample (x1k magnification) showing an irregular inclusion composed mainly of titanium (A1) with very small peaks of iron (A3).
+Small peaks of potassium (A2) can be observed around the inclusion.
+Geological sample SP52_GUI from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M7_3_map2.png"))
+
+```
+
+```{r}
+#| label: fig-m12-3-mp1
+#| fig-cap: "A) general view of the sample (x250 magnification) showing frequent irregular inclusions and porosity.
+B) detail of the cluster of irregular inclusions (x700 magnification) composed of iron (B1) and calcium (B2).
+Geological sample SP67_DESCAM from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M12_3_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m12-1-s2
+#| fig-cap: "A, C) general view of the sample (x250 magnification) showing porosity and uncommon inclusions.
+B) detail of irregular inclusions composed mainly of iron (1).
+D) detail of an inclusion characterised by a large peak of aluminium and potassium, and a small peak of iron and titanium (2, unmarked).
+Dot represents the location of the measured spectra plotted in graphs 1-2.
+Geological sample SP67_DESCAM from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M12_1_s2_s3.png"))
+
+```
+
+```{r}
+#| label: fig-m12-5-s6
+#| fig-cap: "A) general view of the sample (x250 magnification) showing several inclusions and porosity.
+B) detail of inclusions (x1.7k magnification) composed mainly of iron (1-2).
+The inclusion from spectrum 1 is possibly a spherical framboid.
+Dot represents the location of the measured spectra plotted in graphs 1-2.
+Geological sample SP50_CdC from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M12_5_spectra6_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m8-1
+#| fig-cap: "A) iron inclusion (x1k magnification) which was possibly originally a spherical pyrite framboid.
+It is composed mainly of iron (A1) and possibly some indeterminate fossils characterised by a small peak of calcium (A2), phosphorus (A3) and sulphur (A4).
+B) detail of the inclusion (x2.3k magnification).
+Geological sample SP50_CdC from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M8_1_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m12-3-s3
+#| fig-cap: "A) detailed view (x1.5k magnification) and B) close-up (x3.5k magnification) of inclusions composed of iron and aluminium, possibly framboids.
+Geological sample SP67_DESCAM from the MJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M12_3_spectra3.png"))
+
+```
+
+#### Upper Jurassic of eastern Algarve (UJE)
+
+The analysed samples from the Upper Jurassic outcrops (Jordana formation) are homogeneous, with little variation in the identified inclusions.
+The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group.
+Here, we also include the secondary deposition outcrop sample of Peral, since it shows similar patterns to the samples in primary deposition.
+
+They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and large peaks of aluminium, with smaller peaks of aluminium being present in a small number of samples (n=3).
+Small peaks of potassium can also be identified.
+
+Skeletal remains were identified in 30% of the samples (n=3).
+These are composed of calcium and phosphorus and two samples show an associated smaller peak of sulphur (@fig-m6-2-mp1).
+These skeletal remains are all irregular and indeterminate.
+
+Iron oxides are present in all samples, presenting mainly an iron (Fe) composition, and less commonly associated with small titanium (n=2; @fig-m6-1-mp1) and sulphur (n=4) peaks (@fig-m10-3-mp1).
+A single sample shows an inclusion composed of titanium with a small peak of iron (@fig-m10-4-s1).
+
+EDS results also show a small subset of samples with a peak of calcium, rarely with sulphur peaks (n=1) and occasionally associated with iron oxides.
+These calcium peaks occur in irregular, large structures, and represent calcite in the chert or replacing fossils (@fig-m6-3-mp1; @fig-m10-2-mp1; @fig-m10-6-mp2).
+
+Spherical pyrite framboids are present in 20% of the samples (n=2) and are composed of iron or iron with sulphur (@fig-m10-1-mp1).
+
+```{r}
+#| label: fig-m6-2-mp1
+#| fig-cap: "A) general view of the sample (x250 magnification) with a fossil, inclusions and porosity.
+B) detail of the fossil (x1k magnification) composed of calcium (B1), phosphorus (B2) and a small peak of sulphur (B3).
+Iron (B4) is identifiable in certain areas of the fossil which are possibly oxidised and scattered in the sample, corresponding to the white dots.
+Geological sample SP57_JOR from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M6_2_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m6-1-mp1
+#| fig-cap: "A) general view of the sample (x250 magnification) with iron inclusions and a large unidentifiable inclusion.
+B) detail of the inclusion (x1k magnification) composed of aluminium (B1), potassium (B2), magnesium (B3), iron (B4) and possibly small peaks of titanium.
+Geological sample SP59_JOR from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M6_1_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m10-3-mp1
+#| fig-cap: "A) general view of the sample (x250 magnification) with uncommon inclusions.
+B) detail of an irregular inclusion (x3.7k magnification) characterised by iron (B1) and possible small peaks of sulphur (B2).
+Geological sample SP58_JOR_002 from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M10_3_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m10-4-s1
+#| fig-cap: "Detail of an inclusion (x1.9k magnification) mainly composed of titanium and a small peak of iron (1).
+Dot represents the location of the measured spectra plotted in graphs 1.
+Geological sample SP58_JOR from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M10_4_spectra1_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m6-3-mp1
+#| fig-cap: "A) general view (x140 magnification) of a large, irregular calcium inclusion.
+B) detail of a section of the inclusion (x400 magnification), composed mainly of calcium (B1) with small peaks of iron (B2) occurring in the white speckles around the calcium.
+Geological sample SP58_JOR_004 from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M6_3_map1_spectra1.png"))
+
+```
+
+```{r}
+#| label: fig-m10-2-mp1
+#| fig-cap: "A) general view (x140 magnification) of a large, irregular calcium inclusion.
+B) detail of a section of the inclusion (x400 magnification), composed mainly of calcium (B1) with small peaks of iron (B2) occurring in the white speckles around the calcium.
+Geological sample SP58_JOR_004 from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M10_2_map1.png"))
+
+```
+
+```{r}
+#| label: fig-m10-6-mp2
+#| fig-cap: "A) general view of the sample (x250 magnification) showing a large area (lighter grey) composed mainly of calcite, seen in the maps as the presence of calcium (A1) and small peaks of sulphur (A2).
+Small iron concentrations (A3) can also be seen in the sample, represented by the bright white splotches in A.
+Geological sample SP59_JOR_001 from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M10_6_map2.png"))
+
+```
+
+```{r}
+#| label: fig-m10-1-mp1
+#| fig-cap: "A) detail of a spherical pyrite framboid (x1.5k magnification) and B) close-up (x2.7k magnification), composed mainly of iron (A1) and a small peak of sulphur (A2).
+Geological sample SP58_JOR_003 from the UJE formation."
+
+knitr::include_graphics(here::here("figures/within-chert/SOM/M10_1_map1.png"))
+
+```
+
+### Archaeological samples
+
+#### Local chert types
+
+#### Non-local chert types

### 2025-04-23 analysis/supplementary-materials/ESM_4.qmd
ESM_4 finished. Rendered version not included due to size.

--- 2025-04-23 analysis/supplementary-materials/appendix1.qmd
@@
 ---
-title: "Title Goes Here"
-author:
-  - Jane Doe:
-      correspondence: "yes"
-      email: janedoe@fosg.org
-      orcid: 0000-0003-1689-0557
-      institute:
-        - fosg
-        - fop
-  - John Q. Doe:
-      institute: fosg
-      orcid: 0000-0003-1689-0558
-  - Peder Ås:
-      institute: fosg
-      orcid: 0000-0003-1689-0559
-  - Juan Pérez:
-      orcid: 0000-0003-1689-0551
-      institute:
-        - name: Acme Corporation
-  - Max Mustermann:
-      orcid: 0000-0003-1689-0552
-institute:
-  - fosg:
-      name: Formatting Open Science Group
-      address: 23 Science Street, Eureka, Mississippi, USA
-  - fop: Federation of Planets
+title: "Online resource 4. Supplementary graphs and detailed descriptions of XRD and SEM-EDS results"
+subtitle: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
+author: Joana Belmiro
 title-block-published: "Last updated"  
@@
     reference-doc: "../templates/template.docx" # Insert path for the DOCX file
+    toc: true
 execute:
-  echo: true
+  echo: false
   warning: false
@@
   comment: "#>"
-  fig-path: "../figures/SOM"
   fig-dpi: 600
@@
 
-# Online resource 4
-
-## XRD results
-
-In this subsection we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC.
-The original data necessary to produce these visualisations can be found and freely accessed in our online compendium, in the associated GitHub project under the path "../data/xrd_data".
+# XRD results
+
+In this section we present the individual spectra graphs for all XRD samples, organised by geological formation, excluding those presented in the main paper and the results analysed and interpreted at the Geociencias Barcelona-CSIC.
+The original data and code necessary to produce these visualisations can be found and freely accessed in our online compendium (https://doi.org/10.17605/OSF.IO/W78V6), in the associated GitHub project under the path "../data/raw_data/xrd_data" and "../supplementary-materials/ESM_4", respectively.
 
@@
 library(ggpmisc)
+library(here)
 
@@
 # Read grouped supporting data
-xrd_perc <- read_csv('analysis/data/raw_data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
-xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels
+xrd_perc <- read_csv(here::here('analysis/data/raw_data/xrd_data/xrd_goals.csv')) # File with component percentage of all samples for annotations
+xrd_marks <- read_csv(here::here('analysis/data/raw_data/xrd_data/xrd_peaks.csv')) # File with peak data and d spacing for plot labels
 
@@
 # Read individual xrd files
-sp6_xrd <- read_csv('data/xrd_data/SP6_xrd.csv')
-sp7_xrd <- read_csv('data/xrd_data/SP7_xrd.csv')
-sp10_xrd <- read_csv('data/xrd_data/SP10_xrd.csv')
-sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
-sp18_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP18_xrd.csv')
-sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
-sp28_xrd <- read_csv('data/xrd_data/SP28_xrd.csv')
-sp30_xrd <- read_csv('data/xrd_data/SP30_xrd.csv')
-sp33_xrd <- read_csv('data/xrd_data/SP33_xrd.csv')
-sp34_xrd <- read_csv('data/xrd_data/SP34_xrd.csv')
-sp40_xrd <- read_csv('data/xrd_data/SP40_xrd.csv')
-sp50_xrd <- read_csv('data/xrd_data/SP50_xrd.csv')
-sp52_xrd <- read_csv('data/xrd_data/SP52_xrd.csv')
-sp57_xrd <- read_csv('data/xrd_data/SP57_xrd.csv')
-sp59_xrd <- read_csv('data/xrd_data/SP59_xrd.csv')
-sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
-sp62_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP62_xrd.csv')
-sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
-sp66_xrd <- read_csv('data/xrd_data/SP66_xrd.csv')
+sp6_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP6_xrd.csv'))
+sp7_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP7_xrd.csv'))
+sp10_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP10_xrd.csv'))
+sp14_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP14_xrd.csv'))
+sp18_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP18_xrd.csv'))
+sp21_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP21_xrd.csv'))
+sp28_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP28_xrd.csv'))
+sp30_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP30_xrd.csv'))
+sp33_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP33_xrd.csv'))
+sp34_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP34_xrd.csv'))
+sp40_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP40_xrd.csv'))
+sp50_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP50_xrd.csv'))
+sp52_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP52_xrd.csv'))
+sp57_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP57_xrd.csv'))
+sp59_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP59_xrd.csv'))
+sp61_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP61_xrd.csv'))
+sp62_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP62_xrd.csv'))
+sp63_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP63_xrd.csv'))
+sp66_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP66_xrd.csv'))
 
@@
 Labels represent the d spacing values of key mineral peaks.
+#| fig-width: 6
+#| fig-asp: 0.6
+#| fig-align: center
 
@@
   geom_line(linewidth = 0.5) +
-  labs(x = "Diffraction Angle", y = "Intensity") +
-  theme_minimal() +
-  coord_cartesian(ylim = c(-5000, 50000)) +
+  labs(x = "Diffraction Angle", y = "Intensity (count)") +
+  theme_minimal() +
+  coord_cartesian(ylim = c(-5000, 60000)) +
   geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
@@
   scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
-  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp6_perc)))
+  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp6_perc)))
 
@@
   geom_line(linewidth = 0.5) +
-  labs(x = "Diffraction Angle", y = "Intensity") +
+  labs(x = "Diffraction Angle", y = "Intensity (count)") +
   theme_minimal() +
@@
   geom_line(linewidth = 0.5) +
-  labs(x = "Diffraction Angle", y = "Intensity") +
-  theme_minimal() +
-  coord_cartesian(ylim = c(-5000, 50000)) +
+  labs(x = "Diffraction Angle", y = "Intensity (count)") +
+  theme_minimal() +
+  coord_cartesian(ylim = c(-5000, 58000)) +
   geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
@@
   scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "darkgoldenrod1" ,"orchid","snow4")) +
-  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp18_perc)))
+  annotate(geom = "table", x = 75, y = 55000, label = list(cbind(sp18_perc)))
 
@@
   geom_line(linewidth = 0.5) +
-  labs(x = "Diffraction Angle", y = "Intensity (counts)") +
+  labs(x = "Diffraction Angle", y = "Intensity (count)") +
   theme_minimal() +
@@
 
-## SEM-EDS results
+# SEM-EDS results
+
+In this section we present the supplementary SEM photos and associated maps/spectra, organised by geological formation and archaeological chert type, excluding those presented in the main paper.
+The original data and code necessary to produce the spectra visualisations can be found and freely accessed in our online compendium (https://doi.org/10.17605/OSF.IO/W78V6), in the associated GitHub project under the path "../data/raw_data/sem_data" and "../supplementary-materials/ESM_4", respectively.
 
@@
 
-M2_3_map2 <- read_csv('analysis/data/raw_data/sem_data/M2_3_map2.csv')
-M3_1 <- read_csv('analysis/data/raw_data/sem_data/M3_1_map1_map2.csv')
-M3_3_s1 <- read_csv('analysis/data/raw_data/sem_data/M3_3_s1.csv')
-M3_3_s2 <- read_csv('analysis/data/raw_data/sem_data/M3_3_s2.csv')
-M4_2_s5_1 <- read_csv('analysis/data/raw_data/sem_data/M4_2_s5_1.csv')
-M4_2_s5_4 <- read_csv('analysis/data/raw_data/sem_data/M4_2_s5_4.csv')
-M8_2 <- read_csv('analysis/data/raw_data/sem_data/M8_2.csv')
-M10_4 <- read_csv('analysis/data/raw_data/sem_data/M10_4.csv')
-M11_1 <- read_csv('analysis/data/raw_data/sem_data/M11_1.csv')
-M12_1_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M12_1_s2_1.csv')
-M12_1_s3_1 <- read_csv('analysis/data/raw_data/sem_data/M12_1_s3_1.csv')
-M12_3_s3_2 <- read_csv('analysis/data/raw_data/sem_data/M12_3_s3_2.csv')
-M12_5_s1_1 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s1_1.csv')
-M12_5_s1_3 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s1_3.csv')
-M12_5_s3_3 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s3_3.csv')
-M12_5_s6_1 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s6_1.csv')
-M12_5_s6_2 <- read_csv('analysis/data/raw_data/sem_data/M12_5_s6_2.csv')
-M13_1_3_s2 <- read_csv('analysis/data/raw_data/sem_data/M13_1_3_s2.csv')
-M14_3_2_s1_1 <- read_csv('analysis/data/raw_data/sem_data/M14_3_2_s1_1.csv')
-M14_3_2_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M14_3_2_s2_1.csv')
-M14_4_1 <- read_csv('analysis/data/raw_data/sem_data/M14_4_1.csv')
-M15_2_s2_1 <- read_csv('analysis/data/raw_data/sem_data/M15_2_s2_1.csv')
-M15_2_s2_2 <- read_csv('analysis/data/raw_data/sem_data/M15_2_s2_2.csv')
-M16_1_1 <- read_csv('analysis/data/raw_data/sem_data/M16_1_1.csv')
-M18_1_1 <- read_csv('analysis/data/raw_data/sem_data/M18_1_1.csv')
-M18_5_1 <- read_csv('analysis/data/raw_data/sem_data/M18_5_1.csv')
-M20_2_1 <- read_csv('analysis/data/raw_data/sem_data/M20_2_1.csv')
-M22_4.1_1 <- read_csv('analysis/data/raw_data/sem_data/M22_4.1_1.csv')
-M22_5.1_1 <- read_csv('analysis/data/raw_data/sem_data/M22_5.1_1.csv')
-M22_5.1_3 <- read_csv('analysis/data/raw_data/sem_data/M22_5.1_3.csv')
-M23_2_1 <- read_csv('analysis/data/raw_data/sem_data/M23_2_1.csv')
-M23_7_1 <- read_csv('analysis/data/raw_data/sem_data/M23_7_1.csv')
-M24_1_1 <- read_csv('analysis/data/raw_data/sem_data/M24_1_1.csv')
-M24_1_2 <- read_csv('analysis/data/raw_data/sem_data/M24_1_2.csv')
+M2_3_map2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M2_3_map2.csv'))
+M3_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M3_1_map1_map2.csv'))
+M3_3_s1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M3_3_s1.csv'))
+M3_3_s2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M3_3_s2.csv'))
+M4_2_s5_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M4_2_s5_1.csv'))
+M4_2_s5_4 <- read_csv(here::here('analysis/data/raw_data/sem_data/M4_2_s5_4.csv'))
+M8_2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M8_2.csv'))
+M10_4 <- read_csv(here::here('analysis/data/raw_data/sem_data/M10_4.csv'))
+M11_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M11_1.csv'))
+M12_1_s2_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_1_s2_1.csv'))
+M12_1_s3_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_1_s3_1.csv'))
+M12_3_s3_2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_3_s3_2.csv'))
+M12_5_s1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_5_s1_1.csv'))
+M12_5_s1_3 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_5_s1_3.csv'))
+M12_5_s3_3 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_5_s3_3.csv'))
+M12_5_s6_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_5_s6_1.csv'))
+M12_5_s6_2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M12_5_s6_2.csv'))
+M13_1_3_s2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M13_1_3_s2.csv'))
+M14_3_2_s1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M14_3_2_s1_1.csv'))
+M14_3_2_s2_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M14_3_2_s2_1.csv'))
+M14_4_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M14_4_1.csv'))
+M15_2_s2_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M15_2_s2_1.csv'))
+M15_2_s2_2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M15_2_s2_2.csv'))
+M16_1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M16_1_1.csv'))
+M18_1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M18_1_1.csv'))
+M18_5_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M18_5_1.csv'))
+M20_2_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M20_2_1.csv'))
+M22_4.1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M22_4.1_1.csv'))
+M22_5.1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M22_5.1_1.csv'))
+M22_5.1_3 <- read_csv(here::here('analysis/data/raw_data/sem_data/M22_5.1_3.csv'))
+M23_2_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M23_2_1.csv'))
+M23_7_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M23_7_1.csv'))
+M24_1_1 <- read_csv(here::here('analysis/data/raw_data/sem_data/M24_1_1.csv'))
+M24_1_2 <- read_csv(here::here('analysis/data/raw_data/sem_data/M24_1_2.csv'))
 
@@
 #| label: sem-graphs-som
+#| include: false
 
@@
 
-ggsave("M22_4.1_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M22_4.1_1_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-ggsave("M16_1_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M16_1_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-ggsave("M15_2_s2_2_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M15_2_s2_2_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-ggsave("M14_3_2_s2_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M14_3_2_s2_1_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-ggsave("M12_5_s6_2_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M12_5_s6_2_graph.png", width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-ggsave("M8_2_graph.png",width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M8_2_graph.png",width = 5, height = 3.75, units = "in", bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-ggsave("M10_4_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M10_4_graph.png", bg = "white", width = 5, height = 3.75, unit ="in", dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-### Geological samples
-
-#### Lower Jurassic of western Algarve (LJW)
+## Geological samples
+
+### Lower Jurassic of western Algarve (LJW)
 
 The analysed samples from the Lower Jurassic outcrops (Carixian) are homogeneous regarding their structure, inclusions and mineral compositions.
-The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group (Carixian, Lower Jurassic formations of western Algarve, including the inland outcrop of Ferrel).
-
-They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and small peaks of aluminium, with a single sample also showing a high aluminium peak.
+The analysis of their composition and inclusions shows no specific differences between the outcrops, and they will be described as a single group.
+
+They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and small peaks of aluminium, with a single sample also showing a high aluminium peak.
 Small peaks of chlorine and potassium can also be identified in less than 50% of the samples.
@@
 These are all composed of calcium and phosphorus, and frequently show an associated smaller peak of sulphur (\~70% of the samples; @fig-m4-2 and @fig-m3-1).
-These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the fossils (@fig-m4-4, @fig-m2-3-mp1 and @fig-m23-3).
+These fossils may also show some oxidation, due to the associated presence of iron bands or restricted areas within the fossils (@fig-m4-4, @fig-m2-3-mp1 and @fig-m23-3).
 These skeletal remains are frequently irregular and indeterminate; however, a small number of samples show fossils with homogeneous oval shapes, which are currently indeterminate (@fig-m8-3, @fig-m23-2 and @fig-m4-2-m23-2).
 
-Finally, iron oxides are present in all the samples, with an iron (Fe) composition (@fig-m2-1-m4-1, @fig-m23-7 and @fig-m22-4).
+Finally, iron oxides are present in all the samples, with an iron composition (@fig-m2-1-m4-1, @fig-m23-7 and @fig-m22-4).
 Other types of iron oxides are also present in lesser frequencies, as equally large peaks of iron and titanium (40% of the samples; @fig-m4-4) and titanium with a small peak of iron (\<10%).
@@
 Most of these iron or titanium inclusions have irregular shapes and varying sizes.
-Albeit rare (present in 3 samples), pyrite can possibly be identified in the form of spherical framboids or dispersed pyrite nanocrystals (@fig-m22-5-m23-2).
+
+Albeit rare (present in 3 samples), spherical framboids or dispersed nanocrystals were also identified (@fig-m22-5-m23-2).
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M4_2_spectra5_graphs.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M4_2_spectra5_graphs.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M3_1_map1_map2_graphs.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M3_1_map1_map2_graphs.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M4_4_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M4_4_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M2_3_map1_gen.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M2_3_map1_gen.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M23_3_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M23_3_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M8_3_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M8_3_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M23_2_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M23_2_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M4_2_M23_2.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M4_2_M23_2.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M2_1_M4_1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M2_1_M4_1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M23_7_spectra2_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M23_7_spectra2_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M22_4.1_spectra1_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M22_4.1_spectra1_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M2_3_map2_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M2_3_map2_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M22_5.1_M23_2_graph.png"))
-
-```
-
-#### Triassic of western Algarve (TW)
-
-#### Upper Jurassic of western Algarve (UJW)
+knitr::include_graphics(here::here("analysis/figures/SOM/M22_5.1_M23_2_graph.png"))
+
+```
+
+### Triassic of western Algarve (TW)
+
+The analysed sample from the Triassic formations of western Algarve, characterised by thermal origin fibrous quartz veins, is composed of quartz, mainly silica and oxygen and show little to no inclusions (@fig-m13-4).
+
+```{r}
+#| label: fig-m13-4
+#| fig-cap: "General view of the sample (x250 magnification) showing no inclusions or porosity. Geological sample of the Tabual outcrop from the TW formation."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M13_4.png"))
+
+```
+
+### Upper Jurassic of western Algarve (UJW)
 
@@
 
-They are mainly composed of cryptocrystalline quartz (SiO2), and small peaks of aluminium.
+They are mainly composed of cryptocrystalline quartz (SiO~2~), and small peaks of aluminium.
 Small peaks of magnesium were also identified, and calcite was identified in a single sample.
@@
 These are composed of calcium, phosphorus, and also show an associated smaller peak of sulphur.
-These fossils also frequently show some oxidation, due to the associated presence of iron bands or restricted areas within the skeletal remains (@fig-m24-1).
+These fossils also frequently show possible oxidation, due to the associated presence of iron bands or restricted areas within the skeletal remains (@fig-m24-1).
 These skeletal remains are frequently irregular and indeterminate.
 
-Finally, iron oxides are present in 3 of 4 of the samples, with an iron (Fe) composition, but also iron and sulphur or iron and titanium (@fig-m24-4).
+Finally, iron oxides are present in 3 of 4 of the samples, with an iron composition, but also iron and sulphur or iron and titanium (@fig-m24-4).
+
 Although characterized by irregular iron inclusions, iron is also found in framboids.
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M24_1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M24_1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M24_4_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M24_4_map1.png"))
 
@@
 #| label: fig-m24-1-m24-2
-#| fig-cap: "A) detailed view of iron inclusions (x1.8k magnification), possibly originally spherical pyrite framboids (black arrow) and dispersed pyrite nanocrystals (white arrow).
-B) detail of an irregular inclusion (x2.1k magnification) composed mainly of iron (1) and inclusions composed of iron, possibly originally dispersed pyrite nanocrystals (2).
+#| fig-cap: "A) detailed view of iron inclusions (x1.8k magnification), possibly originally spherical pyrite framboids (black arrow) and dispersed nanocrystals (white arrow).
+B) detail of an irregular inclusion (x2.1k magnification) composed mainly of iron (1) and inclusions composed of iron, possibly dispersed nanocrystals (2).
 Dot represents the location of the measured spectrum plotted in graph 1 and 2.
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M24_1_M24_2_graphs.png"))
-
-```
-
-#### Middle Jurassic of eastern Algarve (MJE)
+knitr::include_graphics(here::here("analysis/figures/SOM/M24_1_M24_2_graphs.png"))
+
+```
+
+### Middle Jurassic of eastern Algarve (MJE)
 
@@
 
-They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and small and large peaks of aluminium. Small peaks of chlorine and potassium can also be identified.
+They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and small and large peaks of aluminium. Small peaks of chlorine and potassium can also be identified.
 
@@
 
-Iron oxides are present in more than 90% of the samples, with an iron (Fe) composition (@fig-m12-5-s1).
+Iron oxides are present in more than 90% of the samples, with an iron composition (@fig-m12-5-s1).
 Other types of iron oxides are also present in lesser frequencies, such as equally large peaks of iron and titanium (20% of the samples; @fig-m12-5-s1) and titanium with a small peak of iron (20%; @fig-m7-3).
+
 EDS results also show a small subset of samples with a small peak of calcium frequently associated with iron inclusions (@fig-m12-3-mp1) and inclusions which are mainly composed of large peaks of potassium, aluminium, and magnesium and small peaks of iron or iron with titanium (@fig-m12-1-s2).
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M8_2_map1_spectra1_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M8_2_map1_spectra1_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M12_5_s1_s3_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M12_5_s1_s3_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M7_3_map2.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M7_3_map2.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M12_3_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M12_3_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M12_1_s2_s3.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M12_1_s2_s3.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M12_5_spectra6_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M12_5_spectra6_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M8_1_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M8_1_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M12_3_spectra3.png"))
-
-```
-
-#### Upper Jurassic of eastern Algarve (UJE)
+knitr::include_graphics(here::here("analysis/figures/SOM/M12_3_spectra3.png"))
+
+```
+
+### Upper Jurassic of eastern Algarve (UJE)
 
@@
 
-They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO2), and large peaks of aluminium, with smaller peaks of aluminium being present in a small number of samples (n=3).
+They are mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and large peaks of aluminium, with smaller peaks of aluminium being present in a small number of samples (n=3).
 Small peaks of potassium can also be identified.
@@
 
-Iron oxides are present in all samples, presenting mainly an iron (Fe) composition, and less commonly associated with small titanium (n=2; @fig-m6-1-mp1) and sulphur (n=4) peaks (@fig-m10-3-mp1).
+Iron oxides are present in all samples, presenting mainly an iron composition, and less commonly associated with small titanium (n=2; @fig-m6-1-mp1) and sulphur (n=4) peaks (@fig-m10-3-mp1).
 A single sample shows an inclusion composed of titanium with a small peak of iron (@fig-m10-4-s1).
@@
 
-Spherical pyrite framboids are present in 20% of the samples (n=2) and are composed of iron or iron with sulphur (@fig-m10-1-mp1).
+Spherical framboids are present in 20% of the samples (n=2) and are composed of iron or iron with sulphur (@fig-m10-1-mp1).
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M6_2_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M6_2_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M6_1_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M6_1_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M10_3_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M10_3_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M10_4_spectra1_graph.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M10_4_spectra1_graph.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M6_3_map1_spectra1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M6_3_map1_spectra1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M10_2_map1.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M10_2_map1.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M10_6_map2.png"))
+knitr::include_graphics(here::here("analysis/figures/SOM/M10_6_map2.png"))
 
@@
 
-knitr::include_graphics(here::here("figures/within-chert/SOM/M10_1_map1.png"))
-
-```
-
-### Archaeological samples
-
-#### Local chert types
-
-#### Non-local chert types
+knitr::include_graphics(here::here("analysis/figures/SOM/M10_1_map1.png"))
+
+```
+
+## Archaeological samples
+
+### Local chert types
+
+**Type 1 (T1)** samples are composed of quartz, mainly silica and oxygen and show little to no inclusions (@fig-m13-1-3).
+
+As these correspond to a fibrous quartz of thermal origin, the rare identified inclusions within the chert may be related with naturally occurring fractures and porosity (@fig-m13-1-2).
+
+```{r}
+#| label: fig-m13-1-3
+#| fig-cap: "A) general view of the sample (x250 magnification) without inclusions or visible porosity.
+Archaeological sample from T1 (I21-2394).
+B) general view of the sample (x250 magnification) without inclusions or visible porosity.
+Archaeological sample from T1 (I18-4729)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M13_1_3.png"))
+
+```
+
+```{r}
+#| label: fig-m13-1-2
+#| fig-cap: "A) detail of a porosity or fracture (x1k magnification) filled with iron (A1) and aluminium (A2).
+Archaeological sample from T1 (I21-2394).
+B) general view (x250 magnification) and detailed maps (x450 magnification) of an inclusion or porosity filled with calcium (B1) and sulphur (B2).
+Archaeological sample from T1 (H18-3169)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M13_1_2.png"))
+
+```
+
+**Type 2 (T2)** samples are composed of quartz, mainly silica and oxygen, with small peaks of aluminium. Inclusions vary from uncommon to common, depending on the sample and the area of the sample.
+
+Skeletal remains were identified in a single sample, composed mainly of calcium and phosphorus.
+This fossil is oval, with similarities to those found in other samples from the Lower Jurassic outcrops (@fig-m21-5-mp1).
+
+Other inclusions are irregular or round oxides, composed of iron (@fig-m21-5-mp1).
+In a single sample (sample where the oval fossil was identified), besides iron oxides, titanium with a small iron peak was also identified.
+
+```{r}
+#| label: fig-m21-5-mp1
+#| fig-cap: "A) detail of an oval fossil (x850 magnification) composed of calcium (A1), phosphorus (A2) and small concentrations of iron (A3).
+Iron concentrations can also be seen around the fossil (A3).
+Archaeological sample from T2 (H21-2358)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M21_5_map1.png"))
+
+```
+
+**Type 3 (T3)** samples are homogeneous, showing uncommon inclusions, some porosity and fractures. They are mainly composed of quartz (silica and oxygen), with small peaks of aluminium.
+
+Inclusions include rounded or irregular iron oxides (@fig-m20-2-5), often associated with small peaks of calcium, phosphorus and sulphur, which may be related to the oxidation of skeletal remains.
+
+The presence of iron is also related to the identification of spherical framboids and associated dispersed nanocrystals.
+In both analysed samples, small titanium peaks were also identified in association with iron (@fig-m20-2-5).
+
+```{r}
+#| label: fig-m20-2-5
+#| fig-cap: "A) general view of the sample (x250 magnification) showing uncommon inclusions and porosity.
+Archaeological sample from T3 (H18-4007).
+B) detail of two irregular inclusions (x1k magnification): the inclusion on the right is mainly composed of iron (B1) and the inclusion on the left is mainly composed of iron (B1) and titanium (B2).
+Archaeological sample from T3 (H19-3054)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M20_2_M20_5.png"))
+
+```
+
+**Type 4 (T4)** samples are composed of quartz, mainly silica and oxygen, with small peaks of aluminium.
+Inclusions vary from uncommon to very frequent, depending on the sample and the area of the sample.
+
+Skeletal remains are very frequent, composed mainly of calcium and phosphorus, and with or without the presence of small sulphur peaks (@fig-m15-2-s2).
+Although also irregular-shaped, these fossils are frequently oval and can be found in concentrations of more than 3 fossils (@fig-m15-3-mp1).
+
+Other inclusions are less common and composed mainly of titanium.
+All samples showed the presence of irregular titanium grains, and in two samples grains composed of titanium with small iron peaks were also identified (@fig-m15-2-s2).
+No inclusions composed of large or individual peaks of iron were identified in the samples.
+
+```{r}
+#| label: fig-m15-2-s2
+#| fig-cap: "A) general view of the sample (x270 magnification) showing several inclusions and porosity.
+B) detail of the inclusions (x1.4 magnification): irregular skeletal remains composed of calcium, phosphorus and a small peak of titanium (1); and an irregular inclusion composed mainly of titanium and a small peak of iron (2).
+Dot represents the location of the measured spectra plotted in graphs 1-2.
+Archaeological sample from T4 (H19-3064)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M15_2_spectra2_graph.png"))
+
+```
+
+```{r}
+#| label: fig-m15-3-mp1
+#| fig-cap: "A) general view of a sample (x250 magnification) showing an oval fossil and possibly ghosts of other similar fossils in lighter grey.
+B) detail of the fossil (x1.6k magnification) composed of calcium (B1) and phosphorus (B2).
+Archaeological sample from T4 (H20-2782)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M15_3_map1.png"))
+
+```
+
+**Type 5 (T5)** samples are characterized by high porosity and uncommon inclusions. They are composed of quartz, mainly silica and oxygen, with small peaks of aluminium and magnesium.
+
+Inclusions include rounded or irregular iron oxides, often associated with small peaks of calcium and phosphorus (@fig-m14-3-2 D), which may be related to the oxidation of skeletal remains.
+
+In a single case, a small peak of titanium was also identified in association with iron (@fig-m14-3-2 C).
+
+```{r}
+#| label: fig-m14-3-2
+#| fig-cap: "A) general view of the sample (x250 magnification) showing uncommon inclusions and porosity.
+Archaeological sample from T5 (H20-2573).
+B) general view of the sample (x250 magnification).
+C) detail of an inclusion (x3.7k magnification) composed of iron and a small peak of titanium and aluminium (1).
+D) detail of an inclusion (x1k magnification) composed of iron and small peaks of calcium and phosphorus (2).
+Dot represents the location of the measured spectra plotted in graphs 1-2.
+Archaeological sample from T5 (I20-1942)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M14_3_M14_2_graph.png"))
+
+```
+
+### Non-local chert types
+
+**Type 6 (T6)** samples are cherts composed of cryptocrystalline quartz, characterized by high peaks of silica and oxygen.
+The samples are very homogeneous and with little porosity, and inclusions range from uncommon to rare, and whenever present, are small (@fig-m18-1-s3 C).
+
+These inclusions are always mainly composed of iron (@fig-m18-1-s3), mostly present in the following combinations: 1) single iron peaks; 2) iron peaks with small peaks of calcium and phosphorus, which may be oxidized skeletal remains; 3) iron peaks with associated small peaks of titanium.
+
+```{r}
+#| label: fig-m18-1-s3
+#| fig-cap: "A) detailed view of the sample (x1k magnification).
+B) detail of the inclusion seen in A, composed mostly of iron and small peaks of calcium and phosphorus (1), making this inclusion a possible oxidised fossil.
+Archaeological sample from T6 (I21-2983).
+C) general view of the sample (x250 magnification) showing rare inclusions.
+D) detail of the inclusions (x3.7k magnification) composed mainly of iron (2).
+Archaeological sample from T6 (I21-3673).
+Dot represents the location of the measured spectra plotted in graphs 1-2."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M18_5_spectra2_M18_1_spectra3_graph.png"))
+
+```
+
+**Type 7 (T7)** samples are mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and small to large peaks of aluminium.
+They are homogeneous and with little, observable structures or inclusions (@fig-m17).
+
+Aside from uncommon and very small percentages of skeletal remains (calcium and phosphorus), also associated with sulphur, they show barely any visible skeletal remains or other inclusions, despite their abundant presence observed in thin sections.
+
+Whenever present, inclusions are oxides, either rounded or irregular, and composed of iron (@fig-m17-1-5) and iron and titanium. Pyrite was also identified in a single sample.
+
+```{r}
+#| label: fig-m17
+#| fig-cap: "A-D) General view of samples (x250 magnification) from different facies identified within T7: A) facies 01a (I19-4013); B) facies 02 (I18-3780); C) facies 01b (H21-3663); D) facies 04 (I18-3967)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M17_1_2_3_4.png"))
+
+```
+
+```{r}
+#| label: fig-m17-1-5
+#| fig-cap: "A) detail view of inclusions (x1k magnification) composed mainly of iron (A1).
+Archaeological sample from T7 facies 01a (I19-4013).
+B) A) detail view of inclusions (x1k magnification) composed mainly of iron (A1).
+Archaeological sample from T7 facies 04 (J18-1264)."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M17_1_M17_5.png"))
+
+```
+
+The **Type 8 (T8)** sample shows, like the previous two types, a homogeneous structure, with rare to uncommon inclusions (@fig-m14-4-m16-1 A-B).
+It is mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and large peaks of aluminium.
+
+Similar to the previous type, skeletal remains, despite being abundant in thin sections, are barely visible in the SEM. Whenever present, they show small peaks of calcium, associated with sulphur.
+
+Similarly, oxides are also present in the sample, showing the presence of iron, either in isolated peaks, or combined with large peaks of sulphur (@fig-m14-4-m16-1 B) or titanium.
+
+```{r}
+#| label: fig-m14-4-m16-1
+#| fig-cap: "A) general view the sample (x250 magnification) with rare inclusions.
+B) detail of an inclusion (x3k magnification) composed mainly of sulphur and a smaller peak of iron (1).
+Archaeological sample from T8 (H18-3971).
+B) general view of the sample (x250 magnification) with inclusions and frequent porosity.
+C) detail of the inclusions, composed mainly by iron and a small peak of sulphur (2).
+Archaeological sample from T9 (H18-1982).
+Dot represents the location of the measured spectra plotted in graphs 1-2."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M14_4_M16_1.png"))
+
+```
+
+The **Type 9 (T9)** sample is characterized by high porosity and uncommon, visible inclusions (@fig-m14-4-m16-1 C-D).
+It is mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and small peaks of aluminium and magnesium.
+
+Visible inclusions seem to be small and rounded or irregular.
+They are composed of small peaks of phosphorus and calcium (suggesting poorly preserved skeletal remains), associated with large peaks of iron, which might suggest the oxidation of bioclastic material.
+
+These iron oxides are characterised by either the presence of iron or its combination with sulphur (@fig-m14-4-m16-1 D).
+
+**Type 10 (T10)** is characterised by some porosity and uncommon inclusions (@fig-m16-4-5 A-B).
+It is mainly composed of cryptocrystalline quartz (predominantly silica or SiO~2~), and small peaks of aluminium.
+
+Visible inclusions are small and rounded or irregular. They are composed of small peaks of phosphorus and calcium associated with large peaks of iron.
+
+The iron inclusions also show associated sulphur peaks. Associated with the iron oxides, large copper peaks were also identified, present in irregular-shaped inclusions (@fig-m16-4-5 C-D).
+
+```{r}
+#| label: fig-m16-4-5
+#| fig-cap: "A) general view of the sample (x250 magnification).
+Archaeological sample from T10 (H20-1544).
+B) general view of the sample (x250 magnification).
+C) detail of an inclusion (x1k magnification) and close-up (C1; x3.5k magnification) composed of copper and iron.
+Archaeological sample from T10 (H18-3920).
+Round and oval (although irregular) shapes are visible in the general views (A-B), possibly corresponding to ooids."
+
+knitr::include_graphics(here::here("analysis/figures/SOM/M16_4_M16_5.png"))
+
+```

### 2025-04-24 analysis/paper/paper.qmd
Desktop session. Paper render and improvements.

@@
 title: "Within chert: a multi-technique mineral and geochemical approach to the study of chert of southwestern Iberia"
-author:
-  - Jane Doe:
-      correspondence: "yes"
-      email: janedoe@fosg.org
-      orcid: 0000-0003-1689-0557
-      institute:
-        - fosg
-        - fop
-  - John Q. Doe:
-      institute: fosg
-      orcid: 0000-0003-1689-0558
-  - Peder Ås:
-      institute: fosg
-      orcid: 0000-0003-1689-0559
-  - Juan Pérez:
-      orcid: 0000-0003-1689-0551
-      institute:
-        - name: Acme Corporation
-  - Max Mustermann:
-      orcid: 0000-0003-1689-0552
-institute:
-  - fosg:
-      name: Formatting Open Science Group
-      address: 23 Science Street, Eureka, Mississippi, USA
-  - fop: Federation of Planets
+author: Joana Belmiro
 title-block-published: "Last updated"  
@@
 library(ggpubr)
+library(here)
 
@@
 # Read grouped supporting data
-xrd_perc <- read_csv('analysis/data/raw_data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
-xrd_marks <- read_csv('analysis/data/raw_data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels
+xrd_perc <- read_csv(here::here('analysis/data/raw_data/xrd_data/xrd_goals.csv')) # File with component percentage of all samples for annotations
+xrd_marks <- read_csv(here::here('analysis/data/raw_data/xrd_data/xrd_peaks.csv')) # File with peak data and d spacing for plot labels
 
 # Read individual xrd files
-sp15_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP15_xrd.csv')
-sp58_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP58_xrd.csv')
-sp69_xrd <- read_csv('analysis/data/raw_data/xrd_data/SP69_xrd.csv')
+sp15_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP15_xrd.csv'))
+sp58_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP58_xrd.csv'))
+sp69_xrd <- read_csv(here::here('analysis/data/raw_data/xrd_data/SP69_xrd.csv'))
 
@@
 
-M18_7 <- read_csv('analysis/data/raw_data/sem_data/M18_7.csv')
-M18_8 <- read_csv('analysis/data/raw_data/sem_data/M18_8.csv')
-M4_3 <- read_csv('analysis/data/raw_data/sem_data/M4_3.csv')
+M18_7 <- read_csv(here::here('analysis/data/raw_data/sem_data/M18_7.csv'))
+M18_8 <- read_csv(here::here('analysis/data/raw_data/sem_data/M18_8.csv'))
+M4_3 <- read_csv(here::here('analysis/data/raw_data/sem_data/M4_3.csv'))
 
@@
 
-ggsave("M4_3_graph.png", width = 5, height = 3.75, units = "in", ,bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
+# ggsave("M4_3_graph.png", width = 5, height = 3.75, units = "in", ,bg = "white" , dpi = 300, path = "analysis/figures/SOM/Graphs")
 
@@
 
-pxrf_raw <- read_csv('analysis/data/raw_data/pxrf_data/pxrf_data_dec.csv')
+pxrf_raw <- read_csv(here::here('analysis/data/raw_data/pxrf_data/pxrf_data_dec.csv'))
 
@@
 # Read table
-tbl_samples <- read_csv('analysis/data/raw_data/geochem_samples_n.csv')
+tbl_samples <- read_csv(here::here('analysis/data/raw_data/geochem_samples_n.csv'))
 # Turn NAs into empty strings
@@
 # Read table
-xrd_results_tbl <- read_csv('analysis/data/raw_data/xrd_results.csv')
+xrd_results_tbl <- read_csv(here::here('analysis/data/raw_data/xrd_results.csv'))
 # Turn NAs into empty strings
@@
 Quantifications presented in the supporting table were obtained using Rietveld refinement."
+#| fig-width: 6
+#| fig-asp: 0.8
+#| fig-align: center
 
@@
 #| label: fig-skeletal
-#| fig-cap: "A) Skeletal remains from sample SP16_BLX, from the Lower Jurassic outcrops of western Algarve.
-B) Detail of skeletal remains composed of calcium (B1), phosphorus (B2) and a small peak of sulphur (B3).
-The skeletal remains seem to be oxidized, with small peaks of iron within (B4)."
-
-knitr::include_graphics(here::here("figures/within-chert/M3_3_map1.png"))
+#| fig-cap: "A) general view of the sample (x250 magnification) showing a fracture and a skeletal remain.
+B) detail of skeletal remains (x1.3k magnification) composed of calcium (B1), phosphorus (B2) and a small peak of sulphur (B3).
+The skeletal remains seem to be oxidized, with small peaks of iron within (B4).
+Geological sample SP16_BLX from the LJW formation."
+
+knitr::include_graphics(here::here('analysis/figures/M3_3_map1.png'))
 
@@
 #| label: fig-oval-fossils
-#| fig-cap: "A-C) General view and details from an archaeological sample from local chert type (T4).
-A) general view of a concentration of oval fossils.
-B) general view of a concentration of oval fossils and other inclusions.
-C) detail of an irregular inclusion composed of titanium (C1) and an oval skeletal remains composed mainly of calcium (C2), phosphorus (C3) and a small peak of sulphur."
-
-knitr::include_graphics(here::here("figures/within-chert/M15_4.png"))
+#| fig-cap: "A-B) general view of the sample (x190 and x250 magnification, respectively) showing multiple oval fossils and other inclusions.
+C) detail of an irregular inclusion (x1k magnification) composed of titanium (C1) and an oval skeletal remains composed mainly of calcium (C2), phosphorus (C3) and a small peak of sulphur.
+Archaeological sample from T4 (H20-4359)."
+
+knitr::include_graphics(here::here('analysis/figures/M15_4.png'))
 
@@
 #| label: fig-t6-general
-#| fig-cap: "A-D) General view and details from an archaeological sample from non-local chert type (T6).
-Dots represent the measured spectrum and shown in the corresponding graph.
-A) general view of the sample with uncommon inclusions.
-B) detail of an inclusion seen in A, composed mostly of iron and small peaks of calcium and phosphorus, making this inclusion a possible oxidised skeletal remains of a fossil (1).
-C) general view of the sample with no visible inclusions.
-D) detail of a small and irregular iron inclusion (2)."
-
-knitr::include_graphics(here::here("figures/within-chert/M18_7_M18_8_graphs.png"))
+#| fig-cap: "A) general view of the sample (x250 magnification).
+B) detail of an inclusion seen in A (x3.2k magnification), composed mostly of iron and small peaks of calcium and phosphorus, making this inclusion a possible oxidised skeletal remains of a fossil (1).
+Archaeological sample from T6 (H20-3890).
+C) general view of the sample (x250 magnification) with no visible inclusions.
+D) detail of the sample (x4.2k magnification) showing a small and irregular iron inclusion (2).
+Archaeological sample from T6 (I19-2846).
+Dot represents the location of the measured spectra plotted in graphs 1-2."
+
+knitr::include_graphics(here::here('analysis/figures/M18_7_M18_8_graphs.png'))
 
@@
 #| label: fig-framboids
-#| fig-cap: "A-B) Detail from the geological sample SP59_JOR from the Upper Jurassic outcrops (Jordana formation) of eastern Algarve.
-A) detailed view of a spherical pyrite framboid composed of iron (A1) and sulphur (A2).
-B) detailed view of an inclusion composed of iron (B1) and sulphur (B2)."
-
-knitr::include_graphics(here::here("figures/within-chert/M10_5_map1_map2.png"))
+#| fig-cap: "A) detail of the sample (x6.5k magnification) showing a possibly spherical pyrite framboid composed of iron (A1) and sulphur (A2).
+B) detailed view of an inclusion (x1k magnification) composed of iron (B1) and sulphur (B2).
+Geological sample SP59_JOR_002 from the UJE formation."
+
+knitr::include_graphics(here::here('analysis/figures/M10_5_map1_map2.png'))
 
@@
 #| label: fig-calcite
-#| fig-cap: "General view of calcite inclusions (light grey) from the geological sample SP69_MAR from the Upper Jurassic outcrops of western Algarve, composed mainly of calcium and silica (1).
-Dot represents the measured spectrum and is shown in the corresponding graph."
-
-knitr::include_graphics(here::here("figures/within-chert/M4_3_graphs.png"))
+#| fig-cap: "General view of calcite inclusions (light grey) composed mainly of calcium and silica (1).
+Dot represents the location of the measured spectrum plotted in graph 1.
+Geological sample SP69_MAR from the UJW formation."
+
+knitr::include_graphics(here::here('analysis/figures/M4_3_graphs.png'))
 
@@
 When comparing the regional Algarve cherts (fig.
-++ a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
+@fig-pca a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
 By observing the scatterplots, it is noticeable that the Lower Jurassic cherts (LJW) are highly variable when compared to the other cherts, with samples which display higher positive values in dimension 1 (Dim1) and dimension 2 (Dim2).
@@
 The differences between the local Lower Jurassic cherts and non-local cherts are more marked.
-Fig ++ b represents the correlation between elements and Lower Jurassic cherts, central Portugal Cretaceous cherts and southern Spain Upper Jurassic cherts.
+@fig-pca b represents the correlation between elements and Lower Jurassic cherts, central Portugal Cretaceous cherts and southern Spain Upper Jurassic cherts.
 Dimension 1 explains \~57% of the variability, while dimension 2 explains \~29%, with a combined percentage of `r 56.6+28.6`%.
@@
 A similar pattern emerges when applying a PCA using the same elements but comparing them to archaeological, non-local chert T6, previously attributed to central Portugal (Fig.
-++ c).
+@fig-pca c).
 In this PCA, dimension 1 explains \~52% of the variability and dimension 2 \~30%, with a combined percentage of `r 52.2+30.1`%.
@@
 
-As observed in Fig.
-++ b, Fig.
-++ d shows a concentration of all groups in the centre/bottom of dimension 2 (28.7%) and centre of dimension 1 (58.7%), with the first two dimensions explaining a combined percentage of `r 58.7+28.7`% of the variability.
+As observed in @fig-pca b, @fig-pca d shows a concentration of all groups in the centre/bottom of dimension 2 (28.7%) and centre of dimension 1 (58.7%), with the first two dimensions explaining a combined percentage of `r 58.7+28.7`% of the variability.
 The archaeological local cherts T2 to T5 and non-local type T7 (previously attributed to the Upper Jurassic cherts of southern Spain) are congruent with the corresponding geological sources, although with clear decreased variability.

### 2025-04-25 analysis/paper/paper.qmd
Added discussion and render for second draft review

@@
   keyword 1; keyword 2; keyword 3
-
 ---
@@
   theme_minimal() +
-  coord_cartesian(ylim = c(-5000, 50000)) +
+  coord_cartesian(ylim = c(-5000, 45000)) +
   geom_point(data = sp58_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
@@
   scale_colour_manual(values = c("lightsteelblue3", "salmon", "darkseagreen" ,"orchid","snow4")) +
-  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp58_perc)))
+  annotate(geom = "table", x = 75, y = 40000, label = list(cbind(sp58_perc)))
 
@@
 Quantifications presented in the supporting table were obtained using Rietveld refinement."
-#| fig-width: 6
-#| fig-asp: 0.8
+#| fig-width: 7
+#| fig-height: 5
 #| fig-align: center
@@
 Quantifications presented in the supporting table were obtained using Rietveld refinement."
-
-plot_grid(sp58_plot, sp69_plot, labels = c("a", "b"), rows = 2)
+#| fig-align: center
+#| fig-width: 7
+#| fig-height: 8
+
+plot_grid(sp58_plot, sp69_plot, labels = c("a", "b"), nrow = 2)
 
@@
 Only elements with Success counts above 50% are included."
+#| fig-width: 6
+#| fig-height: 4
+#| fig-align: center
 
@@
 #| fig-cap: "Boxplots of raw element percentages (a: SiO2, b: CaO, c: Fe2O3, d: TiO2) by geological formation or archaeological chert type. Only chert groups with n>5 were included in the boxplots."
-
-ggarrange(si_boxplot, ca_boxplot, fe_boxplot, ti_boxplot, labels = c("a","b","c","d"),
-          ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+#| fig-width: 7
+#| fig-height: 7
+#| fig-align: center
+
+ggarrange(si_boxplot, ca_boxplot, fe_boxplot, ti_boxplot, labels = c("a","b","c","d"), ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
 
@@
 a) PCA using the geological cherts of the regional southern Portugal formations: Lower Jurassic west (LJW), Triassic west (TW), Upper Jurassic west (UPW), Middle Jurassic east (MJE) and Upper Jurassic east (UJE); b) PCA using geological samples of local Lower Jurassic cherts (LJW) and non-local Cretaceous cherts (central Portugal; CPT) and Upper Jurassic cherts (Betic Systems; UCA); c) PCA using geological samples of local Lower Jurassic cherts (LJW), geological chert samples from Cretaceous central Portugal formations (CPT), and archaeological samples from Vale Boi attributed to CPT sources (T6); d) PCA using geological samples of local Lower Jurassic cherts (LJW), geological chert samples from Upper Jurassic southern Spain formations (UCA), and archaeological samples from Vale Boi attributed to local sources (T2-5) and non-local, southern Spain sources (T7)."
+#| fig-width: 8
+#| fig-height: 7
+#| fig-align: center
 
@@
 # Discussion
+
+The results from this study reveal several patterns in the mineral and chemical composition of the analysed chert samples.
+In general, they allowed us to answer and explore the several research questions presented above.
+More importantly, they allow further interpretations about human behaviours related to mobility and raw material procurement and selection.
+All methods used in this chapter allowed a more complete characterisation of the geological cherts from the Algarve region, complementing the petrographic analysis, showing key differences in their mineral and chemical compositions.
+
+For example, the presence of calcite in Upper Jurassic cherts (both eastern and western) seen both in XRD but also SEM-EDS and pXRF provided an added characterisation which may prove useful for future archaeological chert raw material studies.
+This is especially relevant for SEM and pXRF methods, which can be used with little to no impact on the archaeological materials and possibly help distinguish between the visually similarities present among regional cherts (with exception of the Jordana formation cherts in eastern Algarve).
+
+This is especially relevant for the overall goal of providing a detailed lithotheque, both physical and online.
+By expanding the previously used methods to characterise these geological cherts, we have provided a key reference collection of the Algarve region which can, from this point onwards, be consistently used in archaeological studies for prehistoric assemblages across Iberia.
+
+However, despite the added layers of mineral and chemical characterisation to these cherts, one of the main research questions for this paper and regional research [@belmiroCreatingFramesReference2023; @belmiroChertProcurementPatterns2025], continues to be unanswered.
+As mentioned in section 1 “Introduction” one of the goals of this research was to test the potential of several methods to distinguish between the cherts from the Lower Jurassic Carixian (west Algarve basin) and Middle Jurassic Malhão (east Algarve basin) formations.
+
+Perhaps one of the main differences between the cherts from these two formations is related to the presence of dolomite.
+The nodules of LJW show small peaks of dolomite in the XRD results.
+This was expected since a process of dolomitization of the chert nodules, especially on the outer edges, has been identified in previous works [@belmiroCreatingFramesReference2023; @ribeiro_evolucao_2005], and is related to the formation and diagenesis processes of the geological layers and chert nodules [@ribeiro_evolucao_2005; @rochaNoticiaExplicativaFolha1979].
+
+As previously mentioned, however, dolomite is only present in around 60% of the analysed chert samples.
+This may be related to the size and portion of the nodule selected for powdering since it is expected that dolomite will be present at the edges of the nodule, closer to the cortex, and can be absent on the inside of a larger nodule.
+This also may be related to the previously identified heterogeneity in dolomite presence shown in petrographic results or its presence in small percentages within the thin sections [@belmiroCreatingFramesReference2023].
+Similar results were identified in previous mineral and geochemical studies, which showed the presence of different degrees of silicification and dolomitization in nodules in the formations [@ribeiro_evolucao_2005].
+
+In comparison, the XRD results for MJE cherts show a single sample with small dolomite and calcite presence.
+This is probably related to the process of dolomitization occurring in the Carixian (LJW) formation not being present in the Malhão (MJE) formation.
+This is expected given that the chert nodules are found in limestone formations [@manuppellaNoticiaExplicativaFolha2007].
+
+Similarly, elements and compounds like chlorine or calcium oxide are found in found in higher percentages in LJW cherts (Table) but with high standard deviations, meaning that is the presence of outliers.
+Silica is, in all analysed groups, the main component of the cherts.
+However, the LJW group shows a low mean value and high standard deviation (92.4 and 14.7, respectively; Fig), as well as the presence of several outliers, possibly related to an incomplete process of chertification, as identified in previous geochemical works from LJW cherts (Ribeiro, 2005).
+In comparison MJE cherts shows a higher silica mean (96.4; table) and less variability and outliers regarding other elements and compounds, also seen in the PCA results (++ a).
+
+While this variability seems to be specific of LJW cherts, it is not a key element to distinguish between the cherts of the formations, since a great majority of LJW samples show similar values and characteristics to the UJE cherts.
+In other words, if dolomitized or chemically variable with high outliers, the cherts are probably from the Lower Jurassic Carixian formation of western Algarve, but if not exhibiting these characteristics they cannot be ascribed specifically to any of the two.
+
+This means that it is currently still impossible to clearly identify whether chert raw materials are being procured from the eastern Algarve outcrops, and the local LJW cherts are currently still the most probable chert sources being used at Vale Boi during the UP.
+
+Following this interpretation, it is however, possible to use the presence or absence of dolomite and the chemical variability of archaeological cherts to identify other behaviours related to procurement aside from distance and mobility.
+The dolomitization process impacts the knapping characteristics of the nodules.
+This means that there may be the potential for identifying selection behaviours from UP hunter-gatherers focusing on obtaining less dolomitized nodules.
+
+Although the XRD analysis was not applied to archaeological samples, the pXRF results may shed light on this subject.
+In comparison to the results from LJW cherts, the T2-5 cherts show less outliers and variability than their geological counterparts.
+
+This may be related to the process of nodule selection for knapping, choosing mostly less altered chert nodules, and avoiding the highly altered and dolomitized cherts found in the Lower Jurassic formations.
+This is also supported by macroscopic observations of the artefacts knapped in the UP assemblages.
+
+Given the previously referred lack of dolomite and lesser chemical variability from the Middle Jurassic Malhão formation cherts, located east of the site, it is also possible that the T2-5 not only reflects the selection of local cherts, but also the procurement of cherts outside of the local range (30 km; REF).
+However, the lack of clear identifiers between the two geological sources and cherts, and following previous interpretations, it is more likely that the disparities between LJW and T2-5 cherts mostly reflect nodule selection during raw material procurement.
+
+The comparison between other archaeological chert types and the geological samples also provided key results to further explore and ascertain previous source attributions.
+
+Although previously explored, the fibrous quartz chert variety from Vale Boi (T1, known as chalcedony) and the cherts from the Triassic formation of the Algarve also showed clear similarities, especially through the SEM photos and EDS spectra.
+
+Similarly, other local types such as T4 were suggested as not having an exact macroscopic equivalent in the currently known local and regional formations, despite their petrographic similarities.
+However, the SEM-EDS results showed clear similarities in the visible skeletal remain morphologies between T4, T2 and samples from LJW, and the presence of inclusions composed of titanium in T4 and LJW, cementing the interpretation that despite the macroscopic variability, T4 are clearly local, possibly from a limited or currently not available outcrop.
+
+Compared to the similar characteristics of local cherts, both geological and archaeological, non-local types showed clearer differences.
+This allows us to further cement the interpretation that types T6-11 are not local or regional, although their exact sources cannot be pinpointed.
+
+For example, the pXRF results comparing the chemical composion of local cherts, LJW samples, T7 cherts (previously interpreted as coming from the Betic systems) and samples from the UCA lithotheque showed little differences.
+In this case, methods such as macroscopy, petrography [@belmiroChertProcurementPatterns2025] and SEM-EDS proved more adequate to show similarities between T7 and cherts of southern Spain.
+
+In comparison, cases like T11 showed clear differences in their chemical compositions, especially in the presence of copper and zirconium while macroscopically they show no visible fossils or inclusions.
+These black cherts have been identified in south Portugal, in the Alentejo region and central Spain [@araujoRegioesCentralSul2013; @oliveiraTracosGeraisEvolucao1991].
+Further, studies have discussed the presence of copper mineralization present in cherts from the region [@oliveiraComplexoVulcanoSedimentarToca2013].
+While further studies need to explore this, given the lack of a reference collection for cherts and jaspers from southeastern Portugal and central western Spain, or the inclusion of jasper in this study, we suggest that alike TL01 [@belmiroChertProcurementPatterns2025], T11 may have a source in the Zona Ossa Morena in central Iberia.
+
+The T6 cherts also showed clear differences from the local cherts.
+While the results obtained through SEM-EDS are like those surmised from petrographic analysis, T6 cherts show a clear absence of inclusions, fossils or porosity, as well as lower to near absent peaks of aluminium, hinting at different rates of sedimentation (and thus different formation environment) than LJW cherts.
+Unlike T7 cherts, T6 cherts when analysed through a PCA showed a clearer association with geological samples from central Portugal Cretaceous cherts.
+
+These results show that a multi-layered approach to raw material studies, specifically chert, seems to be key for sourcing studies.
+While XRD may not be an ideal technique for archaeological artefacts, given its destructive nature and the general homogeneity of chert, other methods provided their own contributions to source attribution.
+
+For example, coupled with petrography, SEM-EDS provides a good characterisation for cherts and preliminary chemical composition.
+The method showed that with little impact on archaeological artefacts, it was possible to further characterise them and establish parallels or differences between archaeological samples and geological.
+However, it is important to note that for SEM-EDS the only geological samples used were from the regional Jurassic and Triassic chert bearing formations of the Algarve.
+Similarly, this technique was not tested with unpolished samples, possibly placing a difficulty on the identification of some fossils or other inclusions.
+
+A future approach should include added geological samples from cherts of other regions of Iberia which can be directly compared to the archaeological samples and the use of non-polished samples.
+
+Similarly, pXRF provides a quick and non-destructive technique for the chemical characterisation of cherts, being ideal to complement other methods and results.
+However, our results showed some limitations of this methods.
+While the PCAs using pXRF results were able to discern between cherts of different ages (Cretaceous and Jurassic), it showed little differences between cherts of the same age but different formations (Jurassic from the Algarve basin and Jurassic from the Betic systems).
+
+As such, pXRF seems to be an ideal method for preliminary studies and characterisations, that can then be complemented with other geochemical techniques, such as LA-ICP-MS or REEs.
+In this case, a future approach for the geochemical study from Vale Boi would be to apply these methods and compare their potential to the current results.
 

### 2025-04-25 analysis/paper/paper.qmd
Mac work session. Improved graphs and colours

@@
   ggplot(aes(x = `Formation/Type`, y = SiO2, fill = `Formation/Type`)) +
-  geom_boxplot(color = "#5C5C5C") +
+  geom_boxplot(aes(color = `Formation/Type`), alpha = 0.4) +
   stat_summary(fun = mean, geom = "point", shape = 18) +
@@
   theme_minimal() +
-  scale_fill_manual(values = c("LJW" = "#EDA92B", "UJW" = "#F1B6DA", "UJE" = "#8E95C8", "MJE" = "#BA68C8", "CPT" = "#4EB3D3", "UCA" = "#93C54E", "T2-5" = "#FEE391","T6" = "#9CCFD1", "T7" = "#B1C7AB", "T8" = "#D7617A")) +
+  scale_fill_manual(values = c("LJW" = "#EDA92B", "UJW" = "#D7617A", "UJE" = "#5C6BC0", "MJE" = "#BA68C8", "CPT" = "#114D94", "UCA" = "#93C54E", "T2-5" = "#EF6C00","T6" = "#4EB3D3", "T7" = "#00796B", "T8" = "#777")) +
+  scale_color_manual(values = c("LJW" = "#EDA92B", "UJW" = "#D7617A", "UJE" = "#5C6BC0", "MJE" = "#BA68C8", "CPT" = "#114D94", "UCA" = "#93C54E", "T2-5" = "#EF6C00","T6" = "#4EB3D3", "T7" = "#00796B", "T8" = "#777"), guide = "none") +
   theme(axis.title.x = element_blank(),
@@
   ggplot(aes(x = `Formation/Type`, y = CaO, fill = `Formation/Type`)) +
-  geom_boxplot(color = "#5C5C5C") +
+  geom_boxplot(aes(color = `Formation/Type`), alpha = 0.4) +
   stat_summary(fun = mean, geom = "point", shape = 18) +
@@
   ggplot(aes(x = `Formation/Type`, y = TiO2, fill = `Formation/Type`)) +
-  geom_boxplot(color = "#5C5C5C") +
+  geom_boxplot(aes(color = `Formation/Type`), alpha = 0.4) +
   stat_summary(fun = mean, geom = "point", shape = 18) +
@@
   ggplot(aes(x = `Formation/Type`, y = Fe2O3, fill = `Formation/Type`)) +
-  geom_boxplot(color = "#5C5C5C") +
+  geom_boxplot(aes(color = `Formation/Type`), alpha = 0.4) +
   stat_summary(fun = mean, geom = "point", shape = 18) +
@@
                 ellipse.level = 0.80,
-                palette = c("LJW" = "#EDA92B", "CPT" = "#4EB3D3", "UCA" = "#93C54E"),
+                palette = c("LJW" = "#EDA92B", "CPT" = "#114D94", "UCA" = "#93C54E"),
                 col.var = "#757575") +
@@
                 ellipse.level = 0.80,
-                palette = c("LJW" = "#EDA92B", "T6" = "#7FCDBB", "CPT" = "#4EB3D3"),
+                palette = c("LJW" = "#EDA92B", "T6" = "#4EB3D3", "CPT" = "#114D94"),
                 col.var = "#757575") +

### 2025-04-26 analysis/paper/paper.qmd
Improvements to intro

--- 2025-04-25 analysis/paper/paper.qmd
@@
 
-In the Iberian Peninsula - a key region for studying human adaptations and cultural transitions - these approaches have also been applied to UP archaeological assemblages.
-These multi-method strategies have successfully characterized cherts from diverse geological formations and identified mobility patterns throughout the Upper Paleolithic, including evidence of long-distance contacts between hunter-gatherer groups.
-
-In recent years, a wide range of analytical techniques - such as X-Ray Diffraction (XRD), X-Ray Fluorescence (XRF), portable XRF, Laser Ablation Inductively Coupled Plasma Mass Spectrometry (LA-ICP-MS), Raman Spectroscopy, and Scanning Electron Microscopy with Energy Dispersive Spectroscopy (SEM-EDS) - has been employed to enhance chert characterisation in this region.
-Each method offers distinct advantages in data resolution, sample preparation, and cost-effectiveness.
-However, no single technique is universally applicable, and a multi-analytical approach is often necessary to effectively address specific research questions.
-
-Southern Portugal, while having benefited from previous geochemical studies of one of the four chert-bearing formations, as well as macroscopic and petrographic studies of chert-bearing formations (Belmiro et al., 2023) and Upper Palaeolithic chert assemblages from the site of Vale Boi (Belmiro et al., 2025), remains largely unexplored through other analytical techniques.
-These may be key to address open research topics and questions by recent studies.
-
-For example, local and regional Jurassic chert sources are abundant in the region and in relative proximity to the site, yet the differentiation between specific chert formations remains unclear, which is critical for understanding regional mobility patterns (Belmiro et al., 2023).
+As seen in the previous chapters, we have applied macroscopic and petrographic approaches to the study of cherts from the Algarve region (Belmiro et al., 2023, 2025).
+However, the chert-bearing formations from the Algarve region remain largely unexplored through other analytical techniques.
+
+An exception would be the Lower Jurassic (Carixian) and Upper Jurassic (Malm) limestone and dolomitic limestone formations from western Algarve [@rochaNoticiaExplicativaFolha1979], which show previous mineral and geochemical characterisations [@ribeiro_evolucao_2005].
+However, given the geological scope of the study, only 3 chert samples from the Carixian and 3 samples of the Malm were analysed.
+While providing a basis for mineral and chemical composition of these cherts, it is key to expand the analysis to a more robust sample size.
+
+Similarly, the Upper Palaeolithic chert assemblages from the site of Vale Boi have yet to be further characterised through geochemical methods.
+These may be key to address open research topics and questions by recent studies focusing on chert provisioning and selection during the UP at Vale Boi.
+
+For example, the Lower Jurassic (Carixian) cherts, located near Vale Boi (\<20 km) show macroscopic and petrographic similarities to the Middle Jurassic (Malhão formation) cherts located eastwards (\>70 km).
+This makes the differentiation between specific chert formations unclear.
+Making this distinction is critical for understanding regional mobility patterns and whether local cherts at Vale Boi are a mix of local (\<30 km) and regional (xkm) resources (Belmiro et al., 2023).
+
+Questions regarding nodule selection may also be further explored to further characterise close-range chert procurement.
+This is particularly relevant for the Lower Jurassic (Carixian) cherts which show different degrees of dolomitization, with a clear impact on the knapping quality.
+By further characterising the mineral and chemical composition of geological cherts from all outcrops and macroscopic variability, and comparing with the archaeological samples, we may explore behaviours associated to selection associated with nodule quality (sensu higher silicification).
 
@@
 However, these findings were based on macroscopic and petrographic studies, which leaves room for further characterization to strengthen source attributions and test current hypotheses.
-
-Following the results and questions from previous studies by our team (Belmiro et al., 2023; Belmiro et al., 2025), this study employs an integrated analytical approach to geological chert samples from local, regional, and non-local formations, as well as archaeological assemblages from the UP levels of Vale Boi to accomplish three main goals:
+This is especially relevant for archaeological chert types for which we did not have a direct comparison regarding petrography, as for example T7 possibly originating from the Betic systems (south Spain) or other types for which a source could not be identified (T8-11).
+
+To further explore these questions and topics, this study employs an integrated analytical approach to geological chert samples from local, regional, and non-local formations, as well as archaeological assemblages from the UP levels of Vale Boi to accomplish the following goals:
 
@@
 
-By refining the characterization of chert in the region and improving source attribution at the site, this research contributes to broader discussions on lithic procurement, human mobility, and exchange networks in southwestern Iberia during the UP.
+By refining the characterization of chert in the region, improving source attribution at the site, and identifying the strengths and weaknesses of different analytical methods, this research contributes to broader discussions on lithic procurement, human mobility, and exchange networks in southwestern Iberia during the UP.
 
@@
 The standard databases available in the equipment from Bruker were employed: Geochem application and the Dual Mining method.
-A detailed description of the preparation and analysis process is available in our online repository (SOM_1).
+A detailed description of the preparation and analysis process is available in our online repository (Online resource 1).
 
 Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 3).
-After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size (Newlander et al., 2015; SOM ++).
+After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size (Newlander et al., 2015; Online resource 2).
 
@@
 The results presented in this section focus on the main observed patterns of the analysis between outcrops, formations and chert types.
-The detailed analysis/graphs regarding individual samples for all methods can be found in our online repository (SOM_2).
+The detailed analysis/graphs regarding individual samples for all methods can be found in our online repository (Online resource 4).
 
@@
 ++).
-This list and a summarised description of the elements and their presence in sedimentary rocks/chert can be found in the supplementary materials (SOM ++).
+This list and a summarised description of the elements and their presence in sedimentary rocks/chert can be found in the supplementary materials (Online resource 3).
 

### 2025-04-26 analysis/paper/paper.qmd
Improvements to paper and render

@@
 
--   Complete the characterization of chert samples from the Algarve region, including data on their mineral and chemical composition.
-This will deepen our understanding of the characteristics of outcrops across the region and make this information accessible to the scientific community through an online database.
-Additionally, we aim to assess the potential of mineral and chemical analyses to distinguish between macroscopic and petrographically similar chert samples from the Lower Jurassic Carixian (west Algarve basin) and Middle Jurassic Malhão (east Algarve basin) formations.
-This will enhance our understanding of which regional sources were exploited by hunter-gatherer groups at Vale Boi—whether the majority of “pink/yellow” cherts were sourced from local outcrops (within 20 km of Vale Boi) or if some portion was procured from more distant sources.
+-   Expand the characterization of chert samples from the Algarve region, including data on their mineral and chemical composition.
+This will deepen our understanding of the characteristics of outcrops across the region and make this information accessible to the scientific community through its integration in our online lithotheque (https://lusolit.icarehb.com).
+
+-	  Assess the potential of mineral and chemical analyses to distinguish between macroscopic and petrographically similar chert samples from the Lower Jurassic Carixian (west Algarve) and Middle Jurassic Malhão (east Algarve formations.
 
@@
 
-## X-Ray Diffraction (XRD)
-
-X-ray diffraction (XRD) is a widely used laboratory technique for examining the atomic structure of crystalline materials (Adams, 2005).
-The minerals are crystalline, meaning their atoms are arranged in a three-dimensional repeating pattern.
-The crystal structure is defined by the arrangement of atoms within a single unit cell, as well as the cell’s size and shape.
-
-XRD works by directing a beam of X-rays onto a crystal, causing the rays to diffract and produce a characteristic diffraction pattern.
-The d-spacings—the distances between parallel planes of atoms in the crystal lattice—are derived from the angular positions and intensities of the diffracted beams.
-These d-spacings serve as a unique fingerprint for crystalline compounds, allowing for mineral identification (Marshall & Fairbridge, 1998).
-
-Although powder-XRD is a destructive technique, it has been used alongside petrographic characterisation (thin-section analysis) to gain a comprehensive understanding of the petrographic and mineral composition of lithic raw materials, such as chert, in both geological and archaeological contexts.
-This combined approach enables comparisons with potential archaeological implications, particularly in raw material sourcing studies in the Iberian Peninsula (e.g., Columbu et al., 2023; Gómez De Soler et al., 2023; Ortega et al., 2018).
-
-This study utilised X-ray diffraction (XRD) to determine the mineral composition of 23 geological powdered samples of chert from the Algarve region (Table 1).
-These samples, previously characterised and published by our team (Belmiro et al., 2023), represent at least one specimen from each outcrop, with a focus on primary and sub-primary locations across all chert-bearing formations in the region.
-A detailed description of the sample preparation process is available in our online research repository.
+Given the particularities of each analytical technique used in this study, availability of samples and level of destructiveness on the artefacts, and research questions, different samples were used in each method.
+These methods, they application in chert characterization studies, especially in Iberia, and summarized methodology are further discussed below.
+
+Table 1 shows the samples used for each method.
+These included chert geological samples from different regions of Iberia, with a focus on the Algarve region.
+The Algarve materials include geological samples from western and eastern Algarve basin collected from 2021 and 2022 prospection works.
+The western groups include the Triassic formation cherts (sensu Luedke ++), also identified as chalcedony (Triassic west, hereon TW), the Lower Jurassic (Carixian) formation (Lower Jurassic west, hereon LJW) and Upper Jurassic (Malm) formations (Upper Jurassic west, hereon UJW).
 
@@
 
-The samples were analysed using a Bruker AXS D8 Discover XRD with the Da Vinci design, equipped with a Cu Kα source operating at 40 kV and 40 mA, and a Lynxeye 1D detector.
-Scans were conducted from 3° to 75° 2θ, with a 0.05° 2θ step, at the HERCULES Laboratory, University of Évora.
-
-Preliminary phase identification was performed using Diffrac.Suite™ (Bruker) and further refined with Profex (version 5.2.9) after converting raw files using PowDll Converter.
-All data processed in Profex was organised and plotted in R Studio, where visual representations of the XRD data were also generated.
-The code and original files used in this analysis are available in our online repository.
-
-Three additional samples (SP21, SP40, and a geological sample from the Triassic outcrop of Western Algarve) had previously been sent for XRD analysis and interpretation at Geociencias Barcelona-CSIC (Table 1).
-The analysis was conducted using a Bruker D8-A25 XRD, equipped with a Cu source (wavelength λ = 1.5405 Å) and an ultra-fast PSD detector (Lynxeye).
-The diffraction spectrum data were interpreted using the Diffrac.Suite software by Bruker by a technician.
-
-## Scanning Electron Microscopy-X-Ray Microscopy (SEM-EDS)
-
-Scanning electron microscopy coupled with energy-dispersive X-ray spectroscopy (SEM-EDS or SEM-EDX) provides various types of data by combining SEM imaging with elemental analysis.
-SEM is an electron microscopy technique that generates high-resolution images by directing a beam of electrons onto a specimen’s surface.
-As the electron beam interacts with the sample, it scatters as backscattered electrons and emits X-rays.
-As different atomic elements release characteristic X-ray at specific energies, it is possible to identify the elements present in the sample (Kirkbride, 2000) using an Energy dispersive detector (EDS).
-
-Although SEM-EDS is a versatile and efficient method, EDS has a relatively high chemical analysis limit of detection (Kirkbride, 2000).
-
-Despite this limitation, EDS coupled with SEM has been widely used alongside other techniques, such as petrography, XRD, and XRF, to characterise geological and archaeological lithic resources, including chert raw materials.
-This combined approach enhances the understanding of a formation’s characteristics and aids in identifying sources in provenance studies in the Iberian Peninsula (e.g., Abrunhosa et al., 2020; Columbu et al., 2023; Costa et al., 2022; Ortega et al., 2018).
-
-This study employed scanning electron microscopy and energy-dispersive X-ray spectroscopy (SEM-EDS) to analyse the surface structure and elemental composition of the samples.
-The analysis included 39 geological samples from all five chert-bearing formations (Belmiro et al., 2023; Table 2a).
-Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Belmiro et al., 2025; Table 2b).
-
-The samples consisted of lithic chips mounted in resin and sanded until minimal to no scratches were visible under the microscope.
-A detailed description of the sample preparation process and analysis is available in our online repository (SOM_1).
-
-The chip analysis was conducted at the HERCULES Laboratory (Évora) using a variable pressure Hitachi S3700N SEM, coupled with a Quantax EDS microanalysis system.
-Specifically, the Quantax system is equipped with a Bruker AXS 5010 XFlash Silicon Drift Detector, which offers a spectral resolution of 129 eV at FWHM for MnKα.
-The EDS elemental analysis and quantification was carried out with the ESPRIT software by Bruker.
-
-The analysis focused on identifying representative traits of the samples for characterization and comparison.
-Images were captured at 250x and 1000x magnification for each analysed sample.
-To determine the elemental composition, elemental maps and spectra were generated, providing both visual and semi-quantitative data on the selected elements and sample regions.
-At least one map or spectrum was obtained for each distinct type of inclusion or structure identified in the samples.
-
-## Portable X-Ray Fluorescence (pXRF)
-
-X-ray Fluorescence (XRF) works by irradiating the specimen with X-rays, which causes electrons to attain an excited state.
-The specimen emits X-rays (fluorescence) when the atom returns to ground state; this energy is characteristic of the atoms present in the sample.
-As such, XRF spectra contain information about the chemical elements present in the analysed sample (Kirkbride, 2000).
-
-However, laboratory-based XRF has the limitation of being non-portable and requiring sample preparation, which is often incompatible with archaeological studies where assemblages cannot be transported or destroyed.
-In this context, portable XRF (pXRF) instruments offer a practical alternative.
-These devices are easily transportable, non-destructive, and have been successfully used in source discrimination studies, where they provide comparable results to their laboratory-based counterparts (Craig et al., 2007).
-
-pXRF has been used to discriminate between sources of various lithic raw materials, such as obsidian (e.g., Craig et al., 2007; Gemici et al., 2022; Nazaroff et al., 2010), as well as other materials like volcanic and metamorphic rocks (Egeland et al., 2019; Grave et al., 2012).
-
-Studies on chert archaeological assemblages in the Iberian Peninsula have also employed pXRF to distinguish between sources.
-When combined with other methods, pXRF has provided valuable insights (Costa et al., 2022; Gómez De Soler et al., 2023; Pereira et al., 2016, 2021).
-
-For this experiment, a Bruker portable XRF Titan S1 was used in a laboratory benchtop setup, powered by battery (with up to 25% battery charge before being replaced by a fully charged battery).
-The standard databases available in the equipment from Bruker were employed: Geochem application and the Dual Mining method.
-A detailed description of the preparation and analysis process is available in our online repository (Online resource 1).
-
-Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 3).
-After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size (Newlander et al., 2015; Online resource 2).
-
-For geological samples, fresh, flat surfaces were scanned, avoiding altered faces or cortex.
-When necessary, the samples were prepared by breaking the nodules.
-The samples selected represented all varieties of chert identified in the Algarve region, as well as in the archaeological record of Vale Boi.
-Additional samples were included from other regions, such as central Portugal and southern Spain, to facilitate comparisons and test hypotheses based on macroscopic and petrographic data.
-For archaeological samples, artefacts were chosen from previously identified types (Belmiro et al., 2025), prioritising larger, flatter morphologies with minimal surface alterations.
-
-The analysis and reporting protocol were established based on previous studies, emphasising both the accuracy of the obtained data and the transparency and reproducibility of the results (Johnson et al., 2024; Newlander et al., 2015).
-The analysis focused on descriptive statistics and the generation of ANOVAs to compare different groups.
+The eastern groups include the Middle Jurassic (Malhão) formation (Middle Jurassic east, MJE hereon) and the Jordana formation (Upper Jurassic east, hereon UJW).
+
+Other geological samples included in the study were the central Portugal (Rio Maior region) cherts from the Cretaceous formation (hereon CPT), previously prospected and collected by previous researchers (Pereira et al., 2016 Lusolit), and the southern Portugal cherts from the Betic systems (hereon UCA) from the UCA lithotheque.
+
+The archaeological samples used for the study are from the archaeological site of Vale Boi, mainly the Terrace and Shelter areas.
+The samples were chosen based on the groups identified in our previous works, with T1-5 corresponding to the archaeological chert types corresponding to local types, T6 possibly originating from central Portugal Cretaceous formation, T7 and T9 with sources possibly from southern Portugal Betic systems and T8 and T10-11 without a current identified source.
 
@@
 
+## X-Ray Diffraction (XRD)
+
+X-ray diffraction (XRD) is a widely used laboratory technique for examining the atomic structure of crystalline materials (Adams, 2005).
+The minerals are crystalline, meaning their atoms are arranged in a three-dimensional repeating pattern.
+The crystal structure is defined by the arrangement of atoms within a single unit cell, as well as the cell’s size and shape.
+
+XRD works by directing a beam of X-rays onto a crystal, causing the rays to diffract and produce a characteristic diffraction pattern.
+The d-spacings—the distances between parallel planes of atoms in the crystal lattice—are derived from the angular positions and intensities of the diffracted beams.
+These d-spacings serve as a unique fingerprint for crystalline compounds, allowing for mineral identification (Marshall & Fairbridge, 1998).
+
+Although powder-XRD is a destructive technique, it has been used alongside petrographic characterisation (thin-section analysis) to gain a comprehensive understanding of the petrographic and mineral composition of lithic raw materials, such as chert, in both geological and archaeological contexts.
+This combined approach enables comparisons with potential archaeological implications, particularly in raw material sourcing studies in the Iberian Peninsula (e.g., Columbu et al., 2023; Gómez De Soler et al., 2023; Ortega et al., 2018).
+
+This study utilised X-ray diffraction (XRD) to determine the mineral composition of 23 geological powdered samples of chert from the Algarve region (Table 1).
+These samples, previously characterised and published by our team (Belmiro et al., 2023), represent at least one specimen from each outcrop, with a focus on primary and sub-primary locations across all chert-bearing formations in the region.
+A detailed description of the sample preparation process is available in our online research repository (Online resource 1).
+
+The samples were analysed using a Bruker AXS D8 Discover XRD with the Da Vinci design, equipped with a Cu Kα source operating at 40 kV and 40 mA, and a Lynxeye 1D detector.
+Scans were conducted from 3° to 75° 2θ, with a 0.05° 2θ step, at the HERCULES Laboratory, University of Évora.
+
+Preliminary phase identification was performed using Diffrac.Suite™ (Bruker) and further refined with Profex (version 5.2.9) after converting raw files using PowDll Converter.
+All data processed in Profex was organised and plotted in R Studio, where visual representations of the XRD data were also generated.
+
+Three additional samples (SP21, SP40, and a geological sample from the Triassic outcrop of Western Algarve) had previously been sent for XRD analysis and interpretation at Geociencias Barcelona-CSIC (Table 1).
+The analysis was conducted using a Bruker D8-A25 XRD, equipped with a Cu source (wavelength λ = 1.5405 Å) and an ultra-fast PSD detector (Lynxeye).
+The diffraction spectrum data were interpreted using the Diffrac.Suite software by Bruker by a technician.
+
+## Scanning Electron Microscopy-X-Ray Microscopy (SEM-EDS)
+
+Scanning electron microscopy coupled with energy-dispersive X-ray spectroscopy (SEM-EDS or SEM-EDX) provides various types of data by combining SEM imaging with elemental analysis.
+SEM is an electron microscopy technique that generates high-resolution images by directing a beam of electrons onto a specimen’s surface.
+As the electron beam interacts with the sample, it scatters as backscattered electrons and emits X-rays.
+As different atomic elements release characteristic X-ray at specific energies, it is possible to identify the elements present in the sample (Kirkbride, 2000) using an Energy dispersive detector (EDS).
+
+Although SEM-EDS is a versatile and efficient method, EDS has a relatively high chemical analysis limit of detection (Kirkbride, 2000).
+
+Despite this limitation, EDS coupled with SEM has been widely used alongside other techniques, such as petrography, XRD, and XRF, to characterise geological and archaeological lithic resources, including chert raw materials.
+This combined approach enhances the understanding of a formation’s characteristics and aids in identifying sources in provenance studies in the Iberian Peninsula (e.g., Abrunhosa et al., 2020; Columbu et al., 2023; Costa et al., 2022; Ortega et al., 2018).
+
+This study employed scanning electron microscopy and energy-dispersive X-ray spectroscopy (SEM-EDS) to analyse the surface structure and elemental composition of the samples.
+The analysis included 39 geological samples from all five chert-bearing formations (Belmiro et al., 2023; Table 2a).
+Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Belmiro et al., 2025; Table 2b).
+
+The samples consisted of lithic chips mounted in resin and sanded until minimal to no scratches were visible under the microscope.
+A detailed description of the sample preparation process and analysis is available in our online repository (Online resource 1).
+
+The chip analysis was conducted at the HERCULES Laboratory (Évora) using a variable pressure Hitachi S3700N SEM, coupled with a Quantax EDS microanalysis system.
+Specifically, the Quantax system is equipped with a Bruker AXS 5010 XFlash Silicon Drift Detector, which offers a spectral resolution of 129 eV at FWHM for MnKα.
+The EDS elemental analysis and quantification was carried out with the ESPRIT software by Bruker.
+
+The analysis focused on identifying representative traits of the samples for characterization and comparison.
+Images were captured at 250x and 1000x magnification for each analysed sample.
+To determine the elemental composition, elemental maps and spectra were generated, providing both visual and semi-quantitative data on the selected elements and sample regions.
+At least one map or spectrum was obtained for each distinct type of inclusion or structure identified in the samples.
+
+## Portable X-Ray Fluorescence (pXRF)
+
+X-ray Fluorescence (XRF) works by irradiating the specimen with X-rays, which causes electrons to attain an excited state.
+The specimen emits X-rays (fluorescence) when the atom returns to ground state; this energy is characteristic of the atoms present in the sample.
+As such, XRF spectra contain information about the chemical elements present in the analysed sample (Kirkbride, 2000).
+
+However, laboratory-based XRF has the limitation of being non-portable and requiring sample preparation, which is often incompatible with archaeological studies where assemblages cannot be transported or destroyed.
+In this context, portable XRF (pXRF) instruments offer a practical alternative.
+These devices are easily transportable, non-destructive, and have been successfully used in source discrimination studies, where they provide comparable results to their laboratory-based counterparts (Craig et al., 2007).
+
+pXRF has been used to discriminate between sources of various lithic raw materials, such as obsidian (e.g., Craig et al., 2007; Gemici et al., 2022; Nazaroff et al., 2010), as well as other materials like volcanic and metamorphic rocks (Egeland et al., 2019; Grave et al., 2012).
+
+Studies on chert archaeological assemblages in the Iberian Peninsula have also employed pXRF to distinguish between sources.
+When combined with other methods, pXRF has provided valuable insights (Costa et al., 2022; Gómez De Soler et al., 2023; Pereira et al., 2016, 2021).
+
+For this experiment, a Bruker portable XRF Titan S1 was used in a laboratory benchtop setup, powered by battery (with up to 25% battery charge before being replaced by a fully charged battery).
+The standard databases available in the equipment from Bruker were employed: Geochem application and the Dual Mining method.
+A detailed description of the preparation and analysis process is available in our online repository (Online resource 1).
+
+Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 3).
+After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size (Newlander et al., 2015; Online resource 2).
+
+For geological samples, fresh, flat surfaces were scanned, avoiding altered faces or cortex.
+When necessary, the samples were prepared by breaking the nodules.
+The samples selected represented all varieties of chert identified in the Algarve region, as well as in the archaeological record of Vale Boi.
+Additional samples were included from other regions, such as central Portugal and southern Spain, to facilitate comparisons and test hypotheses based on macroscopic and petrographic data.
+For archaeological samples, artefacts were chosen from previously identified types (Belmiro et al., 2025), prioritising larger, flatter morphologies with minimal surface alterations.
+
+The analysis and reporting protocol were established based on previous studies, emphasising both the accuracy of the obtained data and the transparency and reproducibility of the results (Johnson et al., 2024; Newlander et al., 2015).
+The analysis focused on descriptive statistics and the generation of ANOVAs to compare different groups.
+
 # Results
@@
 #| tbl-cap: "Identified crystalline phases of chert geological samples from the Algarve.
-LJW: Lower Jurassic West; MJE: Middle Jurassic East; TW: Triassic West; UPE: Upper Jurassic East; UPW: Upper Jurassic West.
+LJW: Lower Jurassic west (Carixian formation); MJE: Middle Jurassic west (Malm formation); TW: Triassic west; UPE: Upper Jurassic west (Malhão formation); UPW: Upper Jurassic West (Jordana formation).
 (*) refers to results obtained from samples analysed and interpreted at the Geociencias Barcelona-CSIC.
@@
 #| label: fig-xrd-15
-#| fig-cap: "Plot of minerals identified through XRD in the Lower Jurassic (Carixian) chert sample from the outcrop Praia do Belixe (SP15_PBX).
+#| fig-cap: "Plot of minerals identified through XRD in the LJW chert sample from the outcrop Praia do Belixe (SP15_PBX).
 The rhombus shape signifies the main peaks key to identify each mineral.
@@
 In opposition, calcite was mainly identified in the Upper Jurassic samples and the Triassic samples.
-The primary and sub-primary cherts from the Jordana formation (Eastern Algarve) and the Upper Jurassic of Praia da Mareta (Western Algarve) all show the presence of calcite peaks.
-In comparison, only 1 sample out of 12 of the Lower Jurassic outcrops and 1 out of 5 of the Middle Jurassic outcrops show the presence of calcite.
+The primary and sub-primary cherts from the UJE and the UJW all show the presence of calcite peaks.
+In comparison, only 1 sample out of 12 of the LJW and 1 out of 5 of the MJE show the presence of calcite.
 
@@
 #| label: fig-xrd-58-69
-#| fig-cap: "Plot of minerals identified through XRD in the a) eastern Upper Jurassic (Jordana formation) chert sample from the outcrop Jordana (SP58_JOR) and b) western Upper Jurassic (Malm) chert sample from the outcrop Praia da Mareta (SP69_MAR).
+#| fig-cap: "Plot of minerals identified through XRD in the a) UJE chert sample from the outcrop Jordana (SP58_JOR) and b) UJW chert sample from the outcrop Praia da Mareta (SP69_MAR).
 The rhombus shape signifies the main peaks key to identify each mineral.
@@
 Iron oxides were found both as inclusions scattered in the samples, but also frequently associated with skeletal remains, which may be related to their oxidisation.
+
 Pyrite was also identified in the shape of spherical pyrite framboids, as well as dispersed pyrite nanocrystals.
 These are characterised not only by large peaks of iron and associated sulphur, but also by round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids.
-These were present mainly in geological samples from the Upper Jurassic (both west and eastern formations), but barely identified in archaeological samples.
+These were present mainly in geological samples from the Upper Jurassic (UJW and UJE), but barely identified in archaeological samples.
 
@@
 
-In a single sample from Lower Jurassic and samples from local T4 (both characterised by the previously mentioned oval fossils) inclusions composed of titanium (without associated iron) were identified, which may be related to the presence of rutile (TiO~2~).
+In a single sample from LJW and samples from local T4 (both characterised by the previously mentioned oval fossils) inclusions composed of titanium (without associated iron) were identified, which may be related to the presence of rutile (TiO~2~).
 Small peaks of titanium were also identified in 3 out of 4 non-local T6 samples.
@@
 
-Whenever isolated, calcium can be found in high peaks in cherts from Upper Jurassic formations and is probably related to the presence of calcite inclusions.
+Whenever isolated, calcium can be found in high peaks in cherts from Upper Jurassic formations (UJW and UJE) and is probably related to the presence of calcite inclusions.
 
@@
 An exception is the cherts from the Eastern basin of the Algarve (MJE and UJE), samples from non-local archaeological T7 and T8 which show large peaks.
-The exception to the presence of aluminium is the chert of both geological samples from the Triassic formations and archaeological group T1 and all archaeological samples from the non-local group T6, with sources previously attributed to the Cretaceous cherts from Central Portugal.
+The exception to the presence of aluminium is the chert of both geological samples from the Triassic formations and archaeological group T1 and all archaeological samples from the non-local group T6, with sources previously attributed to the CPT formation.
 
@@
 However, there is some variability and outliers in both silica and other elements.
-From these, LJW cherts (Lower Jurassic West) seem to show higher variability and more outliers.
+From these, LJW cherts seem to show higher variability and more outliers.
 This may be related to the higher dolomitization process occurring in LJW cherts.
@@
 
-No clear patterns seem to occur between geological groups and archaeological groups, with the exception of the samples from central Portugal (CPT) and non-local archaeological types T6 and T7.
+No clear patterns seem to occur between geological groups and archaeological groups, with the exception of the CPT samples and non-local archaeological types T6 and T7.
 This is more marked in Fe~2~O~3~ (Fig.
@@
 @fig-pca a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
-By observing the scatterplots, it is noticeable that the Lower Jurassic cherts (LJW) are highly variable when compared to the other cherts, with samples which display higher positive values in dimension 1 (Dim1) and dimension 2 (Dim2).
-However, when focusing on the concentration of LJW results, they seem to fit within the range shown by Middle Jurassic and Upper Jurassic eastern cherts.
-In comparison, Upper Jurassic cherts from the western basin show higher positive values in dimension 2, which is explained by the higher values of CaO.
-
-The differences between the local Lower Jurassic cherts and non-local cherts are more marked.
-@fig-pca b represents the correlation between elements and Lower Jurassic cherts, central Portugal Cretaceous cherts and southern Spain Upper Jurassic cherts.
+By observing the scatterplots, it is noticeable that the LJW cherts are highly variable when compared to the other cherts, with samples which display higher positive values in dimension 1 (Dim1) and dimension 2 (Dim2).
+However, when focusing on the concentration of LJW results, they seem to fit within the range shown by MJE and UJE cherts.
+In comparison, UJW cherts show higher positive values in dimension 2, which is explained by the higher values of CaO.
+
+The differences between the LJW cherts and non-local cherts are more marked.
+@fig-pca b represents the correlation between elements and LJW, CPT and UCA cherts.
 Dimension 1 explains \~57% of the variability, while dimension 2 explains \~29%, with a combined percentage of `r 56.6+28.6`%.
 Once again, it is clear that Jurassic cherts, both from southwestern Portugal and southern Spain, show variability, although they are mostly located in the center/bottom of dimension 2 and center of dimension 1.
-When compared to these values, central Portugal Cretaceous cherts are located in the negative axis of dimension 1 and show little variability.
+When compared to these values, CPT cherts are located in the negative axis of dimension 1 and show little variability.
 
@@
 In this PCA, dimension 1 explains \~52% of the variability and dimension 2 \~30%, with a combined percentage of `r 52.2+30.1`%.
-Similarly to geological Cretaceous samples from central Portugal, T6 samples are also concentrated in the negative side of dimension 1, with a clear overlap between the two.
+Similarly to geological Cretaceous samples from CPT, T6 samples are also concentrated in the negative side of dimension 1, with a clear overlap between the two.
 
@@
 
-However, despite the added layers of mineral and chemical characterisation to these cherts, one of the main research questions for this paper and regional research [@belmiroCreatingFramesReference2023; @belmiroChertProcurementPatterns2025], continues to be unanswered.
-As mentioned in section 1 “Introduction” one of the goals of this research was to test the potential of several methods to distinguish between the cherts from the Lower Jurassic Carixian (west Algarve basin) and Middle Jurassic Malhão (east Algarve basin) formations.
+However, despite the added layers of mineral and chemical characterisation to these cherts, one of the main research questions for this paper and regional research (Belmiro et al., 2025, 2023), continues to be unanswered.
+As previously mentioned, one of the goals of this research was to test the potential of several methods to distinguish between the cherts from the LJW and MJE formations.
 
@@
 While the results obtained through SEM-EDS are like those surmised from petrographic analysis, T6 cherts show a clear absence of inclusions, fossils or porosity, as well as lower to near absent peaks of aluminium, hinting at different rates of sedimentation (and thus different formation environment) than LJW cherts.
-Unlike T7 cherts, T6 cherts when analysed through a PCA showed a clearer association with geological samples from central Portugal Cretaceous cherts.
+Unlike T7 cherts, T6 cherts when analysed through a PCA showed a clearer association with CPT geological samples.
 

### 2025-04-26 analysis/paper/paper.qmd
Refs added. Final draft for review

@@
 
-Understanding the movement of raw materials in prehistory is essential for reconstructing the available technology, human mobility, and trade networks of past hunter-gatherers.
-Among these materials, chert played a crucial role in lithic tool production, making its characterisation and sourcing a key issue in archaeological research, particularly in Upper Palaeolithic (UP) assemblages.
-Determining the origins of chert artefacts reveals patterns of raw material availability and provides valuable insights into prehistoric procurement strategies, territorial behaviours, and long-distance exchange networks over time.
-
-Chert sourcing studies have traditionally relied on macroscopic and petrographic techniques, providing valuable insights into raw material formations and procurement by hunter-gatherer groups.
+The study of lithic raw materials has been a key topic in archaeology, as stone tools are ubiquitous in the archaeological record and identified as key element in hunter-gatherer cultural, economical and social organisation [@odellLithicAnalysis2006; @gamblePalaeolithicSocietiesEurope1999; @tostevinSeeingLithicsMiddlerange2012; @andrefskyRawMaterialAvailabilityOrganization1994; @andrefskyLithics1998; @shottTechnologicalOrganizationSettlement1986].
+Among these materials, chert played a crucial role in lithic tool production, making its study, characterisation and sourcing a key issue in archaeological research, particularly in Upper Palaeolithic (UP) assemblages [e.g., @larickPerigordChertsAnalytical2011; @soto_chert_2018; @tomassoUrgonianChertProvence2019; @eixeaGeochemicalPetrographicAnalyses2023; @jensenFunctionalAnalysisPrehistoric1988; @brandlGenesisProvenanceClassification2013; @brandlMultiLayeredChert2016; @luedtkeIdentificationSourcesChert1979].
+Determining the origins of chert artefacts reveals patterns of raw material availability and provides valuable insights into prehistoric procurement strategies, territorial behaviours, and long-distance exchange networks over time [@ambroseSocialEcologicalModels1990; @kuhnFormalApproachDesign1994; @surovellBehavioralEcologyLithic2009; @whallonSocialNetworksInformation2006; @torrenceProductionExchangeStone1986; @binfordWillowSmokeDogs1980].
+
+Chert sourcing studies have frequently relied on macroscopic and petrographic techniques, providing valuable insights into raw material formations and procurement by hunter-gatherer groups [e.g., @garcia-simonMonegrostypeChertPetrographic2016; @matiasRawMaterialSourcing2016; @tomassoUrgonianChertProvence2019; @gomezdesolerPanadellaChertMontmaneu2020].
 While these methods remain essential in lithic sourcing, they also have limitations.
 Their reliance on physical appearance and petrographic homogeneity can make it difficult to distinguish geologically similar cherts, especially in cases of long-distance procurement.
-Moreover, although macroscopic analysis is non-destructive, thin-section petrography requires destructive sampling, raising concerns about the preservation of archaeological materials.
+Moreover, although macroscopic analysis is non-destructive, thin-section petrography requires destructive sampling, raising concerns about the preservation of archaeological materials [@brandlMultiLayeredChert2016; @luedtkeArchaeologistGuideChert1992].
 
@@
 For example, the Lower Jurassic (Carixian) cherts, located near Vale Boi (\<20 km) show macroscopic and petrographic similarities to the Middle Jurassic (Malhão formation) cherts located eastwards (\>70 km).
-This makes the differentiation between specific chert formations unclear.
-Making this distinction is critical for understanding regional mobility patterns and whether local cherts at Vale Boi are a mix of local (\<30 km) and regional (xkm) resources (Belmiro et al., 2023).
+This makes the differentiation between specific chert formations unclear [@belmiroCreatingFramesReference2023].
+Making this distinction is critical for understanding regional mobility patterns and whether local cherts at Vale Boi are a mix of local (\<30 km) and regional (30-120 km) resources [@tarrinoFlintRawMaterial2015; @herrero-alonsoManagementLithicRaw2020].
 
@@
 This will deepen our understanding of the characteristics of outcrops across the region and make this information accessible to the scientific community through its integration in our online lithotheque (https://lusolit.icarehb.com).
-
 -   Assess the potential of mineral and chemical analyses to distinguish between macroscopic and petrographically similar chert samples from the Lower Jurassic Carixian (west Algarve) and Middle Jurassic Malhão (east Algarve formations.
-
 -   Use geochemical methods to further characterize the archaeological samples, as well as geological samples from other regions of interest (e.g., Cretaceous cherts from central Portugal and cherts from the Betic Systems of southern Spain).
@@
 Additionally, comparing local geological and archaeological data may provide valuable insights into the selection of nodules for knapping, especially considering the known process of dolomitization in the Lower Jurassic outcrops.
-
 -   Test the potential of various techniques and approaches to enhance the interpretation of archaeological chert raw materials and human mobility patterns.
@@
 These included chert geological samples from different regions of Iberia, with a focus on the Algarve region.
-The Algarve materials include geological samples from western and eastern Algarve basin collected from 2021 and 2022 prospection works.
-The western groups include the Triassic formation cherts (sensu Luedke ++), also identified as chalcedony (Triassic west, hereon TW), the Lower Jurassic (Carixian) formation (Lower Jurassic west, hereon LJW) and Upper Jurassic (Malm) formations (Upper Jurassic west, hereon UJW).
+The Algarve materials include geological samples from western and eastern Algarve basin collected from 2021 and 2022 prospection works [@belmiroCreatingFramesReference2023].
+The western groups include the Triassic formation cherts (fibrous chalcedony), also identified as chalcedony (Triassic west, hereon TW), the Lower Jurassic (Carixian) formation (Lower Jurassic west, hereon LJW) and Upper Jurassic (Malm) formations (Upper Jurassic west, hereon UJW) [@rochaNoticiaExplicativaFolha1979; @terrinhaBaciaAlgarveEstratigrafia2006; @ribeiro_evolucao_2005].
 
@@
 
-The eastern groups include the Middle Jurassic (Malhão) formation (Middle Jurassic east, MJE hereon) and the Jordana formation (Upper Jurassic east, hereon UJW).
-
-Other geological samples included in the study were the central Portugal (Rio Maior region) cherts from the Cretaceous formation (hereon CPT), previously prospected and collected by previous researchers (Pereira et al., 2016 Lusolit), and the southern Portugal cherts from the Betic systems (hereon UCA) from the UCA lithotheque.
+The eastern groups include the Middle Jurassic (Malhão) formation (Middle Jurassic east, MJE hereon) and the Jordana formation (Upper Jurassic east, hereon UJW) [@manuppellaNoticiaExplicativaFolha2007; @terrinhaBaciaAlgarveEstratigrafia2006].
+
+Other geological samples included in the study were the central Portugal (Rio Maior region) cherts from the Cretaceous formation (hereon CPT; @matiasRawMaterialSourcing2016), previously prospected and collected by previous researchers [@pereirab_presenting_2016], and the southern Portugal cherts from the Upper Jurassic of the Betic systems [@rodriguezPrehistoricFlintExploitations2011] (hereon UCA), obtained from the lithotheque of the Unit of Geoarchaeology and Archaeometry Applied to Historic Artistic and Monumental Heritage (University of Cadiz).
 
 The archaeological samples used for the study are from the archaeological site of Vale Boi, mainly the Terrace and Shelter areas.
-The samples were chosen based on the groups identified in our previous works, with T1-5 corresponding to the archaeological chert types corresponding to local types, T6 possibly originating from central Portugal Cretaceous formation, T7 and T9 with sources possibly from southern Portugal Betic systems and T8 and T10-11 without a current identified source.
-
-For further reproducibility and replicability, and working towards the goal of open science (Johnson et al., 2024; Marwick, 2017), all analyses were performed through R Studio.
-The R code used for the analysis, datasets, and visual representations presented in this paper are available through our online repository (++).
+The samples were chosen based on the groups identified in our previous works [@belmiroChertProcurementPatterns2025], with T1-5 corresponding to the archaeological chert types corresponding to local types, T6 possibly originating from central Portugal Cretaceous formation, T7 and T9 with sources possibly from southern Portugal Betic systems and T8 and T10-11 without a current identified source.
+
+For further reproducibility and replicability, and working towards the goal of open science [@Marwick2017], all analyses were performed through R Studio.
+The R code used for the analysis, datasets, and visual representations presented in this paper are available through our online repository (https://doi.org/10.17605/OSF.IO/DBFT2).
 We utilised the rrtools package by Marwick et al.
-(2018) to create a research compendium and produce a reproducible journal article.
-The repository includes the complete set of raw data used in the analysis, along with a custom R project (Wickham, 2015) that contains the necessary code to generate all tables and figures.
+[-@marwickPackagingDataAnalytical2018] to create a research compendium and produce a reproducible journal article.
+The repository includes the complete set of raw data used in the analysis, along with a custom R project [@wickhamPackagesOrganizeTest2015] that contains the necessary code to generate all tables and figures.
 To facilitate broad reuse, the code is provided under the MIT license, the data under CC‐0, and the figures under CC‐BY.
-For more details, see Marwick (2017).
 
@@
 
-X-ray diffraction (XRD) is a widely used laboratory technique for examining the atomic structure of crystalline materials (Adams, 2005).
+X-ray diffraction (XRD) is a widely used laboratory technique for examining the atomic structure of crystalline materials [@adamsXRAYABSORPTIONDIFFRACTION2005].
 The minerals are crystalline, meaning their atoms are arranged in a three-dimensional repeating pattern.
@@
 The d-spacings—the distances between parallel planes of atoms in the crystal lattice—are derived from the angular positions and intensities of the diffracted beams.
-These d-spacings serve as a unique fingerprint for crystalline compounds, allowing for mineral identification (Marshall & Fairbridge, 1998).
+These d-spacings serve as a unique fingerprint for crystalline compounds, allowing for mineral identification [@marshallEncyclopediaGeochemistry1998].
 
 Although powder-XRD is a destructive technique, it has been used alongside petrographic characterisation (thin-section analysis) to gain a comprehensive understanding of the petrographic and mineral composition of lithic raw materials, such as chert, in both geological and archaeological contexts.
-This combined approach enables comparisons with potential archaeological implications, particularly in raw material sourcing studies in the Iberian Peninsula (e.g., Columbu et al., 2023; Gómez De Soler et al., 2023; Ortega et al., 2018).
+This combined approach enables comparisons with potential archaeological implications, particularly in raw material sourcing studies in the Iberian Peninsula [e.g., @columbuMultiAnalyticalTechniquesDefine2023; @gomezdesolerMultitechniqueApproachCharacterization2023; @ortegaChertCastelltallatFormation2018].
 
 This study utilised X-ray diffraction (XRD) to determine the mineral composition of 23 geological powdered samples of chert from the Algarve region (Table 1).
-These samples, previously characterised and published by our team (Belmiro et al., 2023), represent at least one specimen from each outcrop, with a focus on primary and sub-primary locations across all chert-bearing formations in the region.
+These samples represent at least one specimen from each outcrop, with a focus on primary and sub-primary locations across all chert-bearing formations in the region.
 A detailed description of the sample preparation process is available in our online research repository (Online resource 1).
@@
 As the electron beam interacts with the sample, it scatters as backscattered electrons and emits X-rays.
-As different atomic elements release characteristic X-ray at specific energies, it is possible to identify the elements present in the sample (Kirkbride, 2000) using an Energy dispersive detector (EDS).
-
-Although SEM-EDS is a versatile and efficient method, EDS has a relatively high chemical analysis limit of detection (Kirkbride, 2000).
+As different atomic elements release characteristic X-ray at specific energies, it is possible to identify the elements present in the sample [@kirkbrideANALYTICALTECHNIQUESSpectroscopic2000] using an Energy dispersive detector (EDS).
+
+Although SEM-EDS is a versatile and efficient method, EDS has a relatively high chemical analysis limit of detection [@kirkbrideANALYTICALTECHNIQUESSpectroscopic2000].
 
 Despite this limitation, EDS coupled with SEM has been widely used alongside other techniques, such as petrography, XRD, and XRF, to characterise geological and archaeological lithic resources, including chert raw materials.
-This combined approach enhances the understanding of a formation’s characteristics and aids in identifying sources in provenance studies in the Iberian Peninsula (e.g., Abrunhosa et al., 2020; Columbu et al., 2023; Costa et al., 2022; Ortega et al., 2018).
+This combined approach enhances the understanding of a formation’s characteristics and aids in identifying sources in provenance studies in the Iberian Peninsula [e.g., @abrunhosaPetrographicSEMEDXCharacterization2020; @columbuMultiAnalyticalTechniquesDefine2023; @costaLithicArrowheadsSiliceous2022; @ortegaChertCastelltallatFormation2018].
 
 This study employed scanning electron microscopy and energy-dispersive X-ray spectroscopy (SEM-EDS) to analyse the surface structure and elemental composition of the samples.
-The analysis included 39 geological samples from all five chert-bearing formations (Belmiro et al., 2023; Table 2a).
-Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Belmiro et al., 2025; Table 2b).
+The analysis included 39 geological samples from all five chert-bearing formations (Table 2a).
+Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Table 2b).
 
@@
 The specimen emits X-rays (fluorescence) when the atom returns to ground state; this energy is characteristic of the atoms present in the sample.
-As such, XRF spectra contain information about the chemical elements present in the analysed sample (Kirkbride, 2000).
+As such, XRF spectra contain information about the chemical elements present in the analysed sample [@kirkbrideANALYTICALTECHNIQUESSpectroscopic2000].
 
@@
 In this context, portable XRF (pXRF) instruments offer a practical alternative.
-These devices are easily transportable, non-destructive, and have been successfully used in source discrimination studies, where they provide comparable results to their laboratory-based counterparts (Craig et al., 2007).
-
-pXRF has been used to discriminate between sources of various lithic raw materials, such as obsidian (e.g., Craig et al., 2007; Gemici et al., 2022; Nazaroff et al., 2010), as well as other materials like volcanic and metamorphic rocks (Egeland et al., 2019; Grave et al., 2012).
+These devices are easily transportable, non-destructive, and have been successfully used in source discrimination studies, where they provide comparable results to their laboratory-based counterparts [@craigComparisonXRFPXRF2007].
+
+pXRF has been used to discriminate between sources of various lithic raw materials, such as obsidian [e.g., @craigComparisonXRFPXRF2007; @gemiciNewInsightsMesolithic2022; @nazaroffAssessingApplicabilityPortable2010], as well as other materials like volcanic and metamorphic rocks [@egelandGeochemicalPhysicalCharacterization2019; @graveNondestructivePXRFMafic2012].
 
 Studies on chert archaeological assemblages in the Iberian Peninsula have also employed pXRF to distinguish between sources.
-When combined with other methods, pXRF has provided valuable insights (Costa et al., 2022; Gómez De Soler et al., 2023; Pereira et al., 2016, 2021).
+When combined with other methods, pXRF has provided valuable insights [@costaLithicArrowheadsSiliceous2022; @gomezdesolerMultitechniqueApproachCharacterization2023; @pereiraRawMaterialProcurement2021; @pereiraLithicEconomyTerritory2016].
 
@@
 Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 3).
-After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size (Newlander et al., 2015; Online resource 2).
+After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size [@newlanderEmpiricalStudyEffect2015; Online resource 2].
 
@@
 Additional samples were included from other regions, such as central Portugal and southern Spain, to facilitate comparisons and test hypotheses based on macroscopic and petrographic data.
-For archaeological samples, artefacts were chosen from previously identified types (Belmiro et al., 2025), prioritising larger, flatter morphologies with minimal surface alterations.
-
-The analysis and reporting protocol were established based on previous studies, emphasising both the accuracy of the obtained data and the transparency and reproducibility of the results (Johnson et al., 2024; Newlander et al., 2015).
+For archaeological samples, artefacts were chosen from previously identified types, prioritising larger, flatter morphologies with minimal surface alterations.
+
+The analysis and reporting protocol were established based on previous studies, emphasising both the accuracy of the obtained data and the transparency and reproducibility of the results [@johnsonBestPracticesPublishing2024; @newlanderEmpiricalStudyEffect2015].
 The analysis focused on descriptive statistics and the generation of ANOVAs to compare different groups.
@@
 
-However, despite the added layers of mineral and chemical characterisation to these cherts, one of the main research questions for this paper and regional research (Belmiro et al., 2025, 2023), continues to be unanswered.
+However, despite the added layers of mineral and chemical characterisation to these cherts, one of the main research questions for this paper and regional research, continues to be unanswered.
 As previously mentioned, one of the goals of this research was to test the potential of several methods to distinguish between the cherts from the LJW and MJE formations.
@@
 
-Given the previously referred lack of dolomite and lesser chemical variability from the Middle Jurassic Malhão formation cherts, located east of the site, it is also possible that the T2-5 not only reflects the selection of local cherts, but also the procurement of cherts outside of the local range (30 km; REF).
+Given the previously referred lack of dolomite and lesser chemical variability from the Middle Jurassic Malhão formation cherts, located east of the site, it is also possible that the T2-5 not only reflects the selection of local cherts, but also the procurement of cherts outside of the local range.
 However, the lack of clear identifiers between the two geological sources and cherts, and following previous interpretations, it is more likely that the disparities between LJW and T2-5 cherts mostly reflect nodule selection during raw material procurement.

### 2025-04-26 analysis/paper/paper.qmd
Correction of small mistakes and render

@@
 
-When comparing the regional Algarve cherts (fig.
-@fig-pca a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
+When comparing the regional Algarve cherts (@fig-pca a), we see that the two first dimensions explain a combined `r 34.6+26.3`% variability in the results, with dimension 1 explaining 34.6% and dimension 2 explaining 26.3%.
 By observing the scatterplots, it is noticeable that the LJW cherts are highly variable when compared to the other cherts, with samples which display higher positive values in dimension 1 (Dim1) and dimension 2 (Dim2).
@@
 
-A similar pattern emerges when applying a PCA using the same elements but comparing them to archaeological, non-local chert T6, previously attributed to central Portugal (Fig.
-@fig-pca c).
+A similar pattern emerges when applying a PCA using the same elements but comparing them to archaeological, non-local chert T6, previously attributed to central Portugal (@fig-pca c).
 In this PCA, dimension 1 explains \~52% of the variability and dimension 2 \~30%, with a combined percentage of `r 52.2+30.1`%.
@@
 
-As observed in @fig-pca b, @fig-pca d shows a concentration of all groups in the centre/bottom of dimension 2 (28.7%) and centre of dimension 1 (58.7%), with the first two dimensions explaining a combined percentage of `r 58.7+28.7`% of the variability.
+@fig-pca d shows a concentration of all groups in the centre/bottom of dimension 2 (28.7%) and centre of dimension 1 (58.7%), with the first two dimensions explaining a combined percentage of `r 58.7+28.7`% of the variability.
 The archaeological local cherts T2 to T5 and non-local type T7 (previously attributed to the Upper Jurassic cherts of southern Spain) are congruent with the corresponding geological sources, although with clear decreased variability.

### 2025-04-26 analysis/paper/paper.qmd
Git stash issue, solved. Render

@@
 This will deepen our understanding of the characteristics of outcrops across the region and make this information accessible to the scientific community through its integration in our online lithotheque (https://lusolit.icarehb.com).
+
 -   Assess the potential of mineral and chemical analyses to distinguish between macroscopic and petrographically similar chert samples from the Lower Jurassic Carixian (west Algarve) and Middle Jurassic Malhão (east Algarve formations.
+
 -   Use geochemical methods to further characterize the archaeological samples, as well as geological samples from other regions of interest (e.g., Cretaceous cherts from central Portugal and cherts from the Betic Systems of southern Spain).
@@
 Additionally, comparing local geological and archaeological data may provide valuable insights into the selection of nodules for knapping, especially considering the known process of dolomitization in the Lower Jurassic outcrops.
+
 -   Test the potential of various techniques and approaches to enhance the interpretation of archaeological chert raw materials and human mobility patterns.
@@
 The standard databases available in the equipment from Bruker were employed: Geochem application and the Dual Mining method.
-A detailed description of the preparation and analysis process is available in our online repository (Online resource 1).
-
-Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 3).
+A detailed description of the preparation and analysis process and a description of the elements and compounds used in the statistical analysis are available in our online repository (Online resource 1 and 3, respectively).
+
+Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 1).
 After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size [@newlanderEmpiricalStudyEffect2015; Online resource 2].
@@
 
-The SEM-EDS results show that all analysed samples are mainly composed of SiO~2~.
+The SEM-EDS results show that all analysed samples are mainly composed of SiO~2~. A complete table with all identified elements per sample is available in our online repository (Online resource 5).
 
@@
 ++).
-This list and a summarised description of the elements and their presence in sedimentary rocks/chert can be found in the supplementary materials (Online resource 3).
+This list and a summarised description of the elements and their presence in sedimentary rocks/chert can be found in the supplementary materials (Online resource 3), as well as a table with means, counts and standard deviations of each of these elements and compounds per group (geological formation or archaeological chert type; Online resource 6).
 
@@
 This may be related to the higher dolomitization process occurring in LJW cherts.
-Similarly, the lower mean values of silica for UJW (Upper Jurassic West) samples may be related to high values of calcium oxide (Fig.
-++).
+Similarly, the lower mean values of silica for UJW (Upper Jurassic West) samples may be related to high values of calcium oxide (@fig-pxrf-boxplots).
 
@@
 No clear patterns seem to occur between geological groups and archaeological groups, with the exception of the CPT samples and non-local archaeological types T6 and T7.
-This is more marked in Fe~2~O~3~ (Fig.
-++ b) and, to a lesser degree, in TiO~2~ (Fig.
-++ c), which show smaller means and values for these groups.
-The similarity between these groups can also be seen in the near absence of Sr; where other groups show smaller values but are well represented, the numbers of CPT, T6 and T8 samples are reduced to \<50%, meaning Sr (Table s++) presence is limited in these cherts.
-Similarly, CPT and T6 cherts also show less than 50% samples with Zr (Table s++), making this element also limited in these chert groups.
+This is more marked in Fe~2~O~3~ (@fig-pxrf-boxplots b) and, to a lesser degree, in TiO~2~ (@fig-pxrf-boxplots c), which show smaller means and values for these groups.
+The similarity between these groups can also be seen in the near absence of Sr; where other groups show smaller values but are well represented, the numbers of CPT, T6 and T8 samples are reduced to \<50%, meaning Sr (Online resource 6) presence is limited in these cherts.
+Similarly, CPT and T6 cherts also show less than 50% samples with Zr (Online resource 6), making this element also limited in these chert groups.
 
@@
 As expected, other elements show much higher percentages, especially Fe~2~O~3~, TiO~2~, Al~2O~3 and K~2~O.
-T11 shows the lower percentages of silica (\~90%; Table s++), and, similarly to T9-10 high peaks of Fe~2~O~3~, TiO~2~, K~2~O.
+T11 shows the lower percentages of silica (\~90%; Online resource 6), and, similarly to T9-10 high peaks of Fe~2~O~3~, TiO~2~, K~2~O.
 Compared to all the other samples analysed in this study, T11 shows the highest values for Cu and Zr.
@@
 
-Fig ++ shows the PCAs for several combinations of elements and comparisons between different groups to understand the various relations.
+@fig-pca shows the PCAs for several combinations of elements and comparisons between different groups to understand the various relations.
 The elements used for the PCAs were S, Cl, P~2~O~5~, Fe~2~O~3~, TiO~2~, K~2~O and CaO.

### 2025-04-26 analysis/paper/paper.qmd
Mac session. Figure and table refs + render

@@
 
-Table 1 shows the samples used for each method.
+@tbl-samples shows the samples used for each method.
 These included chert geological samples from different regions of Iberia, with a focus on the Algarve region.
@@
 
-This study utilised X-ray diffraction (XRD) to determine the mineral composition of 23 geological powdered samples of chert from the Algarve region (Table 1).
+This study utilised X-ray diffraction (XRD) to determine the mineral composition of 23 geological powdered samples of chert from the Algarve region (@tbl-samples).
 These samples represent at least one specimen from each outcrop, with a focus on primary and sub-primary locations across all chert-bearing formations in the region.
@@
 
-Three additional samples (SP21, SP40, and a geological sample from the Triassic outcrop of Western Algarve) had previously been sent for XRD analysis and interpretation at Geociencias Barcelona-CSIC (Table 1).
+Three additional samples (SP21, SP40, and a geological sample from the Triassic outcrop of Western Algarve) had previously been sent for XRD analysis and interpretation at Geociencias Barcelona-CSIC (@tbl-samples).
 The analysis was conducted using a Bruker D8-A25 XRD, equipped with a Cu source (wavelength λ = 1.5405 Å) and an ultra-fast PSD detector (Lynxeye).
@@
 This study employed scanning electron microscopy and energy-dispersive X-ray spectroscopy (SEM-EDS) to analyse the surface structure and elemental composition of the samples.
-The analysis included 39 geological samples from all five chert-bearing formations (Table 2a).
-Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (Table 2b).
+The analysis included 39 geological samples from all five chert-bearing formations (@tbl-samples).
+Additionally, 27 samples were examined from archaeological artefacts, representing all previously identified chert types (@tbl-samples).
 
@@
 
-Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (Table 1).
+Chert samples from various sources and types, including both geological and archaeological specimens, were scanned (@tbl-samples).
 After scanning, the thickness and diameter of the scanned face were measured to ensure that each sample met the minimum required thickness and size [@newlanderEmpiricalStudyEffect2015; Online resource 2].
@@
 
-The XRD results are homogeneous, showing little differences between samples.
+The XRD results are homogeneous, showing little differences between samples (@tbl-xrd-results).
 The samples are predominantly composed of quartz.
@@
 
-Dolomite was mainly identified in the samples from the Lower Jurassic and Triassic outcrops of the western Algarve.
-Small dolomite peaks are present in more than \~60% of the Lower Jurassic cherts and the single sample from the Triassic.
+Dolomite was mainly identified in the samples from LJW and TW outcrops.
+Small dolomite peaks are present in more than \~60% of the LJW cherts and the single sample from the TW.
 In comparison, only 1 of the 11 samples from the Middle and Upper Jurassic outcrops showed the presence of dolomite.
@@
 
-In opposition, calcite was mainly identified in the Upper Jurassic samples and the Triassic samples.
+In opposition, calcite was mainly identified in the Upper Jurassic samples and the TW samples.
 The primary and sub-primary cherts from the UJE and the UJW all show the presence of calcite peaks.
@@
 
-Similarly, the presence of skeletal remains, represented by peaks of calcium, phosphorus and frequently accompanied by small peaks of sulphur, occurs in samples from all formations and chert types.
+Similarly, the presence of skeletal remains, represented by peaks of calcium, phosphorus and frequently accompanied by small peaks of sulphur, occurs in samples from all formations and chert types (@fig-skeletal).
 Although frequently without a discernible shape, oval fossils (although currently unidentified) were observed in LJW geological samples and archaeological samples from local cherts (T2 and T4).
-This type of large, oval fossil, also composed of calcium, phosphorus and sulphur, seems to be limited to local cherts and to specific areas of nodules, characterised by high amounts of these skeletal remains.
+
+This type of large, oval fossil, also composed of calcium, phosphorus and sulphur, seems to be limited to local cherts and to specific areas of nodules, characterised by high amounts of these skeletal remains (@fig-oval-fossils).
 
@@
 By observing the surfaces of samples and the presence of skeletal remains, some differences could be noted between local cherts and non-local archaeological samples.
-In comparison to the frequent to common skeletal remains in the Jurassic cherts from the Algarve, and archaeological cherts attributed to local sources (T2-5), cherts from non-local sources such as T6-8 show little porosity and rare presence of inclusions.
+In comparison to the frequent to common skeletal remains in the Jurassic cherts from the Algarve, and archaeological cherts attributed to local sources (T2-5), cherts from non-local sources such as T6-8 show little porosity and rare presence of inclusions (@fig-t6-general) .
 Whenever present, skeletal remains are small and show small peaks.
@@
 Pyrite was also identified in the shape of spherical pyrite framboids, as well as dispersed pyrite nanocrystals.
-These are characterised not only by large peaks of iron and associated sulphur, but also by round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids.
+These are characterised not only by large peaks of iron and associated sulphur, but also by round/angular small nanocrystals organised in spherical shapes, semi-spherical or simply dispersed crystals, frequently in association with spherical framboids (@fig-framboids).
 These were present mainly in geological samples from the Upper Jurassic (UJW and UJE), but barely identified in archaeological samples.
@@
 
-Whenever isolated, calcium can be found in high peaks in cherts from Upper Jurassic formations (UJW and UJE) and is probably related to the presence of calcite inclusions.
+Whenever isolated, calcium can be found in high peaks in cherts from Upper Jurassic formations (UJW and UJE) and is probably related to the presence of calcite inclusions (@fig-calcite).
 
@@
 The pXRF measured 44 major and minor elements, although, for a large majority, these were mostly composed of values below the detection level (\< LOD) or with values equal to 0.
-Only elements with more than 50% (n\>80) of successful values were maintained in the analysis to allow the successful interpretation of the data (Fig.
-++).
+Only elements with more than 50% (n\>80) of successful values were maintained in the analysis to allow the successful interpretation of the data (@fig-pxrf-barplot).
 This list and a summarised description of the elements and their presence in sedimentary rocks/chert can be found in the supplementary materials (Online resource 3), as well as a table with means, counts and standard deviations of each of these elements and compounds per group (geological formation or archaeological chert type; Online resource 6).
@@
 
-Finally, archaeological types T9 to T11 seem to show the most differences in element percentages of all analysed samples; due to the small sample, instead of boxplots (with a minimum in the sample of 5), the results are shown in Table s++ in the SOM.
+Finally, archaeological types T9 to T11 seem to show the most differences in element percentages of all analysed samples; due to the small sample, instead of boxplots (with a minimum in the sample of 5), the results are shown in Online Resource 6.
 Albeit with some variation, T9 and T10 to show lower percentages of silica in comparison to other samples.
@@
 
-Similarly, elements and compounds like chlorine or calcium oxide are found in found in higher percentages in LJW cherts (Table) but with high standard deviations, meaning that is the presence of outliers.
+Similarly, elements and compounds like chlorine or calcium oxide are found in higher percentages in LJW cherts (Online resource 6) but with high standard deviations, meaning that is the presence of outliers.
 Silica is, in all analysed groups, the main component of the cherts.
-However, the LJW group shows a low mean value and high standard deviation (92.4 and 14.7, respectively; Fig), as well as the presence of several outliers, possibly related to an incomplete process of chertification, as identified in previous geochemical works from LJW cherts (Ribeiro, 2005).
-In comparison MJE cherts shows a higher silica mean (96.4; table) and less variability and outliers regarding other elements and compounds, also seen in the PCA results (++ a).
+However, the LJW group shows a low mean value and high standard deviation (92.4 and 14.7, respectively; Online resource 6), as well as the presence of several outliers (@fig-pxrf-boxplots a), possibly related to an incomplete process of chertification, as identified in previous geochemical works from LJW cherts (Ribeiro, 2005).
+In comparison MJE cherts shows a higher silica mean (96.4; Online resource 6) and less variability and outliers regarding other elements and compounds, also seen in the PCA results ( @fig-pca a).
 
@@
 While the results obtained through SEM-EDS are like those surmised from petrographic analysis, T6 cherts show a clear absence of inclusions, fossils or porosity, as well as lower to near absent peaks of aluminium, hinting at different rates of sedimentation (and thus different formation environment) than LJW cherts.
-Unlike T7 cherts, T6 cherts when analysed through a PCA showed a clearer association with CPT geological samples.
+Unlike T7 cherts, T6 cherts when analysed through the PCA showed a clearer association with CPT geological samples (@fig-pca c).
 

### 2025-04-27 analysis/paper/paper.qmd
Changed table. Added abstract, keywords, conclusion and SOM list

@@
 abstract: |
-  Text of abstract
+  Recent advances in the study of chert in the Algarve have applied macroscopic and petrographic methods to geological samples from regional chert-bearing formations and archaeological materials from the Upper Palaeolithic assemblages of Vale Boi to explore hunter-gatherer mobility and technological organisation.
+However, key questions regarding human behaviour, mobility, and raw material selection remain unanswered, largely due to the limitations of macroscopic and petrographic methods in fully characterising the cherts.
+This study expands the characterisation of both geological and archaeological samples through the application of additional analytical methods, aiming to address critical research questions and assess the potential of these methods for chert characterisation in the Algarve and at Vale Boi.
+The results demonstrate that a multi-layered analytical approach is most effective for chert characterisation.
+While the applied geochemical methods yield strong results in distinguishing between cherts of different geological periods, they struggle to differentiate between cherts from the same period but different epochs.
+By building on previous analyses, this research contributes to broader discussions on lithic procurement strategies, human mobility, and exchange networks in southwestern Iberia during the Upper Palaeolithic.
 keywords: |
-  keyword 1; keyword 2; keyword 3
+  Geoarchaeology; XRD; SEM-EDS; pXRF
 ---
@@
 
+This study highlights the importance of applying a multi-method analytical framework to the characterisation of cherts from the Algarve region.
+By combining mineralogical and geochemical techniques like XRD, SEM-EDS and pXRF, we were able to refine the characterisation of geological and archaeological cherts and address key questions related to raw material procurement and human mobility during the Upper Palaeolithic at Vale Boi.
+
+Although some challenges remain—particularly the difficulty of consistently distinguishing between cherts from the Lower Jurassic West (LJW) and Middle Jurassic East (MJE) formations—the integration of mineralogical indicators, such as the presence of dolomite and chemical variability, has opened new avenues for understanding raw material selection strategies.
+
+The results suggest that prehistoric groups engaged in selective procurement behaviours, favouring cherts with better knapping qualities, such as those showing less dolomitization.
+Furthermore, the use of geochemical methods of both local and non-local cherts within the Vale Boi assemblages and their comparison to samples from other regions strengthens the previous source attributions.
+This is especially relevant for clearly defining central Portugal as a source for a portion of non-local cherts present in Vale Boi.
+
+The study also underscores the value and limitations of different analytical methods.
+Techniques such as SEM-EDS and pXRF proved highly effective for an (almost) non-destructive analysis.
+However, pXRF alone showed limited ability to distinguish between closely related chert formations, highlighting the need for complementary high-precision geochemical techniques in future research.
+
+Finally, this work has contributed to the development of a robust regional lithotheque for the Algarve, providing a critical reference point for future studies of prehistoric lithic assemblages in Iberia.
+
+Future research should expand the geological reference collections beyond the Algarve and incorporate higher-resolution geochemical analyses to further refine raw material sourcing.
+In doing so, it will be possible to deepen our understanding of human behavioural adaptations and regional interaction networks during the Upper Palaeolithic.
+
 # Acknowledgements
@@
 I would like to thank Dr.
-José Mirão for the training in the XRD and SEM-EDS methods and aid in the interpretation of the results, and the Hércules laboratory, for facilitating the equipment necessary for preparation and analyses (++).
+José Mirão for the training in the XRD and SEM-EDS methods and aid in the interpretation of the results, and the HERCULES laboratory, for facilitating the equipment necessary—the analyses were developed at the HERCULES Laboratory of the University of Évora, funded by FCT through the projects UIDB/04449/2020 and UIDP/04449/2020.
 Still from the Hércules team, I thank the technician Sandra Velez for the training, help and bright mood during the preparation of SEM-EDS samples.
-In a similar fashion, I thank Dr.
+Similarly, I thank Dr.
 Javier Sanchéz-Martinéz for the help in obtaining chips for analysis through meticulous knapping whenever necessary.
@@
 In particular, I would like to thank Dr.
-Hussein Kanbar for the support during the analysis and key insights for interpretation and visual presentation of geochemical data; and technician Pedro ++ and Dr.
+Hussein Kanbar for the support during the analysis and key insights for interpretation and visual presentation of geochemical data; and technician Pedro Coxito and Dr.
 Vera Aldeias for the continuous support and advice.
+
+# Supplementary Online Materials
+
+**Online resource 1:** Sample preparation and detailed method description.
+
+**Online resource 2:** pXRF sample measurements (thickness and min. diameter).
+
+**Online resource 3:** Chemical elements/compounds and geological implications.
+
+**Online resource 4:** Supplementary graphs and detailed description of XRD and SEM-EDS results.
+
+**Online resource 5:** SEM-EDS results by sample. Presence of chemical elements (EDS) and inclusions/fossils (SEM).
+
+**Online resource 6:** pXRF descriptive statistics table.
 

### 2025-05-03 analysis/paper/paper.qmd
Small updates + ggsaves for the main thesis

@@
 
+#ggsave("xrd_15.png", bg = "white", width = 7, height = 5, unit ="in", dpi = 300, path = "analysis")
+
 ```
@@
 
-plot_grid(sp58_plot, sp69_plot, labels = c("a", "b"), nrow = 2)
+grid_plot <- plot_grid(sp58_plot, sp69_plot, labels = c("a", "b"), nrow = 2)
+
+#ggsave("grid_plot.png", bg = "white", width = 7, height = 8, unit ="in", dpi = 300, path = "analysis")
 
@@
 
+#ggsave("succ_barplot.png", bg = "white", width = 6, height = 4, unit ="in", dpi = 300, path = "analysis")
+
 ```
@@
 #| label: fig-pxrf-boxplots
-#| fig-cap: "Boxplots of raw element percentages (a: SiO2, b: CaO, c: Fe2O3, d: TiO2) by geological formation or archaeological chert type. Only chert groups with n>5 were included in the boxplots."
+#| fig-cap: "Boxplots of raw element percentages (a: SiO2, b: CaO, c: Fe2O3, d: TiO2) by geological formation or archaeological chert type.
+Only chert groups with n>5 were included in the boxplots.
+Diamond symbol represents mean values within each group."
 #| fig-width: 7
@@
 
-ggarrange(si_boxplot, ca_boxplot, fe_boxplot, ti_boxplot, labels = c("a","b","c","d"), ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+pxrf_boxplots <- ggarrange(si_boxplot, ca_boxplot, fe_boxplot, ti_boxplot, labels = c("a","b","c","d"), ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
+
+#ggsave("pxrf_boxplots.png", bg = "white", width = 7, height = 7, unit ="in", dpi = 300, path = "analysis")
 
@@
 
-plot_grid(biplot_alg, biplot_nl, biplot_comp, biplot_comp2, labels = c("a", "b", "c", "d"))
+pca_plots <- plot_grid(biplot_alg, biplot_nl, biplot_comp, biplot_comp2, labels = c("a", "b", "c", "d"))
+
+#ggsave("pca_plots.png", bg = "white", width = 9, height = 7, unit ="in", dpi = 300, path = "analysis")
 
